---
title: "ch2_00d_BLAST"
output: html_document
date: "2023-10-10"
---

Changes Sept 2024:
We're greatly reducing our usage of ASVs moving forward so I've eliminated some of the unnecessary steps to add in asv conversions.

Blast modified from Evan's script to use system2

```{r}
library(tidyverse)
library(taxize)
library(here)

# here filepaths
pecofastq <- here('rawdata', 'peco', 'fastq', 'miseq_nextseq', 'combined_next_mi') 
datefield <- '20260331_'
datefield_new <- '20260331_'
# if different
datefield_swarm <- '20260331_'
runfield <- 'peco_next-mi-combined_'

# Destination folders
processededna <- here('processeddata', 'peco', 'nextseq_miseq')
```

# blast

Evan's call:

mkdir blast_96_sim_LCA_besthit
blastn 
-task megablast 
-num_threads 38 
-evalue 1e-5 
-max_target_seqs 10 
-perc_identity 96 
-qcov_hsp_perc 50 
-db /data/taxonomyDBs/NCBI_NT/2023-06-08/nt 
-outfmt '6 qseqid stitle sacc staxid pident qcovs evalue bitscore' 
-query 12S.PECO_combined.ASV_sequences.fasta  
-out blast_96_sim_LCA_besthit/12S_ASV_sequences.length_var.blast.out


Note that you need to have a blast database set up locally; this is not an api call

```{r run-blast}
#path.blast <- file.path(here(calvertfastq, "blast_96sim"))
#if(!dir.exists(path.blast)) dir.create(path.blast)

mifishdb <- file.path("/Users/katesheridan/Documents/GitHub/peco/database12s/crabs/blastdb_make/crabs-mifish")


edna_to_blast <- file.path(pecofastq,
                           '20260331_peco_next-mi-combined_asv-sequences.fasta')
blast_out <- file.path(pecofastq, paste0(datefield_new, runfield, 'swarm_blastresults_asv.txt'))

# check its working
system2("blastn", args = "-h")

system2("blastn", args=c("-task", "'megablast'", # which kind of blast
                         #"-num_threads", 38, # for parallelizing
                         "-evalue", 1e-5,
                         "-max_target_seqs", 50,
                         "-perc_identity", 85,
                         "-qcov_hsp_perc", 50,
                         "-db", mifishdb, # directory of db to blast against
                         "-outfmt", "'6 qseqid pident evalue sacc stitle bitscore qcovs mismatch'",
                         "-query", edna_to_blast,
                         "-out", blast_out
                         ))



```



# cleanup

```{r cleanup-motus}
# top 10 list
asv_50blast <- read_delim(here(pecofastq, paste0(datefield_new, runfield, 'swarm_blastresults_asv.txt')),
                           delim = '\t', col_names = c("asv", 
                                                       "identity_percentage", "evalue", "accession", "hit_name",
                                                       "bitscore", "coverage", "n_mismatches")) %>%
  mutate(hit_name = trimws(hit_name)) %>%
  mutate(hit_name = str_remove_all(hit_name, "(root;)")) %>%
  separate_wider_delim(hit_name, delim = ";", names = c('kingdom', 'phylum', 'class', 'order',
                                                        'family', 'genus', 'species')) %>%
  mutate(species = str_replace_all(species, "_", " ")) %>%
  # fill in NAs
  mutate(across(kingdom:genus, ~ na_if(., "")))

# motu numbers to join with the blast results
motu_key <- read_csv(here(pecofastq, '20260331_peco_next-mi-combined_lulu_motu2asv.csv')) %>%
  rename(motu = motu_lulu)

# load-in keys
taxa_key <- read_csv(here('processeddata', 'splists',
                     '20260130_12s_crabs_taxonomy-key.csv'))[-1]
```


 previous notes

 5 assign_tax
 taxonomy assignment by accession number
 --missing for custom names, same format as lineage input
 note content from db_import may need to be updated first
open a github issue later
crabs assign_tax --input pgaout.fasta --output taxassigned.tsv --acc2tax nucl_gb.accession2taxid --taxid nodes.dmp --name names.dmp


## top 50 taxonomy
```{r top hits}
# strip and replace higher taxonomy
# remove out of region and hybrids
blast50_updatetaxa <- asv_50blast %>%
    # remove columns that will be duplicated
  select(!c(kingdom, phylum, class, order, family, genus, species)) %>%
  left_join(motu_key) %>%
  left_join(taxa_key) %>%
  # remove known incorrect hits when good hits exist (from manual check)
  # species out of region
  filter(!(c(found_taxa %in% c('Amphisticus argenteus', 'Aspiolucius esocinus',
                            'Alcichthys elongatus', 'Alburnus alburnus',
                            'Albula pacifica',
                            'Acipenser mikadoi', 'Acipenser schrenckii',
                            'Acipenser dabryanus', 'Acantholabrus palloni',
                            'Acanthogobius stigmothonus', "Anisotremus virginicus",
                            'Anisotremus scapularis', 'Anisotremus surinamensis', 
                            'Acanthogobius hasta', 'Alosa alosa', 'Alosa agone', 'Alosa fallax',
                            'Ammodytes japonicus', 'Ammodytes marinus', 'Ammodytes tobianus',
                            'Ammodytes dubius', 'Ammodytes americanus', 
                            'Argyrosomus amoyensis', 'Argyrosomus japonicus',
                            'Anchoa mitchilli', 'Atheresthes evermanni', 
                            'Aplodinotus grunniens', 'Askoldia variegata',
                            #'Barbatula barbatula', 'Barbatula hispanica', 
                            'Brachygenys chrysargyreum', 'Ballerus sapa',
                            'Bairdiella chrysoura', 'Bero elegans', 'Bryx dunckeri',
                            'Bahaba taipingensis', 'Boesemania microlepis',
                            'Batrachocottus talievi', 'Batrachocottus nikolskii', 
                            'Batrachocottus multiradiatus', 'Brachyopsis segaliensis',
                            'Chondrostoma nasus', "Clariger papillosus",
                            'Chirolophis saitone', 'Chirolophis ascanii', 
                            'Chirolophis wui', 'Chirolophis japonicus',
                            'Cleisthenes pinetorum', 'Cleisthenes herzensteini',
                            'Collichthys lucidus',
                            'Cottus gobio', 'Comephorus dybowskii', 'Cottus bairdii',
                            'Cottus pollux', 'Cottus szanaga', 'Cottus poecilopus', 
                            'Cottus czerskii', 'Cottus perifretum', 'Cottus hangiongensis',
                            'Clupea harengus', 'Conodon nobilis', 
                            'Cottocomephorus grewingki', 'Cottocomephorus grewingkii',
                            'Cottocomephorus inermis', 'Comephorus baikalensis',
                            'Cosmocampus elucens', 'Cryptacanthodes bergi',
                            'Ctenogobius boleosoma', 'Ctenogobius saepepallens',
                            'Ctenogobius shufeldti', 'Ctenolabrus rupestris',
                            "Cryptacanthodes maculatus", 'Cynoscion nothus',
                            'Cynoscion arenarius', 'Cynoscion nebulosus',
                            'Cynoscion regalis',
                            'Ditrema jordani', 'Dicentrarchus labrax',
                            'Dicentrarchus punctatus',
                            'Ernogrammus hexagrammus', 'Eques lanceolatus',
                            'Enchelyopus cimbrius',
                            'Equetus punctatus', 'Evorthodus lyricus',
                            'Gadus morhua','Gadus finnmarchicus',
                            'Gasterosteus nipponicus', 'Gasterosteus wheatlandi',
                            'Glyptocephalus stelleri', 'Gymnogobius macrognathos',
                            'Glyptocephalus cynoglossus', 'Gasterosteus nipponicus',
                            'Hyperoplus immaculatus', 'Hippoglossus hippoglossus',
                            'Hexagrammos agrammus', 'Hexagrammos otakii', 
                            'Haemulon carbonarium', 'Haemulon flavolineatum',
                            'Haemulon macrostomum', 'Haemulon plumierii',
                            'Hypomesus japonicus', 'Haemulon sciurus',
                            'Huso dauricus',
                            'Hyperoplus lanceolatus', 'Hozukius emblemarius', 
                            'Haemulon aurolineatum', 'Hypsoblennius ionthas',
                            'Hypleurochilus geminatus', 'Hypleurochilus springeri',
                            'Kasatkia memorabilis', 'Leuciscus baicalensis',
                            'Leuciscus merzbacheri', 'Leiostomus xanthurus',
                            'Liparis fabricii',
                            'Liparis miostomus', 'Liparis tessellatus', 
                            'Liparis tanakae', 'Liparis inquilinus',
                            'Leuciscus leuciscus', 'Limanda limanda', 'Limanda sakhalinensis',
                            'Lepidopsetta mochigarei', 'Liopsetta pinnifasciata', 
                            'Leuciscus idus', 'Leuciscus oxyrrhis', 'Lutjanus griseus',
                            'Leuciscus burdigalensis', 'Lutjanus argentimaculatus',
                            'Myzopsetta ferruginea', 'Microstomus achne', 
                            # introduced on the west coast
                            #'Morone saxatilis',
                            'Myoxocephalus scorpius', 'Myoxocephalus aenaeus',
                            'Myoxocephalus brandtii', 
                            'Malacocottus gibber',
                            'Melanogrammus aeglefinus', 'Menticirrhus littoralis',
                            'Miichthys miiuy', 'Netuma thalassina',
                            # there is a gbif record in alaskan panhandle
                            #'Myoxocephalus jaok',
                            # good point nearby?
                            #'Occella dodecaedron'
                            'Oncorhynchus kawamurae', 'Opisthocentrus tenuis',
                            'Oncorhynchus gilae', 'Opisthocentrus ocellatus',
                            'Opisthocentrus zonope',
                            'Oreochromis niloticus', 'Occella iburia',
                            'Orthopristis chrysoptera', 'Oxyurichthys cornutus',
                            'Parahucho perryi', 'Procottus major', 
                            'Pareques umbrosus', 'Parablennius thysanius',
                            'Parablennius yatabei', 'Pholidapus dybowskii',
                            'Paralichthys olivaceus', 'Paralichthys adspersus',
                            'Platichthys flesus', 'Pholis gunnellus', 
                            'Pholis nebulosa', 'Pholis crassispina', 
                            'Pholis picta', 'Peprilus medius', 
                            'Pogonias cromis', 'Pungitius laevis',
                            'Pomadasys stridens', 'Pomadasys argenteus',
                            'Pomadasys kaakan', 'Peprilus triacanthus',
                            'Peprilus burti', 'Peprilus paru',
                            'Psenopsis anomala', 'Platichthys bicoloratus',
                            'Rutilus rutilus', 'Scardinius erythrophthalmus',
                            'Salvelinus elgyticus', 'Salvethymus svetovidovi',
                            'Salvelinus boganidae', 'Salvelinus kuznetzovi',
                            'Salvelinus taranetzi', 'Salvlinus albus',
                            'Salvelinus schmidti', 'Salvelinus levanidovi',
                            'Salvelinus kronocius', 'Salvelinus drjagini',
                            'Salvelinus drjagini', 'Salvelinus albus',
                            'Salvelinus leucomaenis', 'Salvelinus curilus',
                            'Scomber scombrus',
                            'Scartella emarginata', 'Scartella cristata',
                            'Serranus scriba', 'Serranus cabrilla',
                            'Sebastes mentella', 'Sebastes hubbsi',
                            'Sebastes inermis', 'Sebastes wakiyai',
                            'Sebastes cheni', 'Sebastes pachycephalus',
                            'Sebastes joyneri', 'Sebastes baramenuke',
                            'Sebastes viviparus', 'Sebastes norvegicus',
                            'Sebastes mentella', 'Sebastes koreanus',
                            'Sprattus sprattus', #'Squalius cephalus', ?maybe?
                            'Squalus acanthias', 'Squalus japonicus',
                            'Squalus mitsukurii', 'Sciaenops ocellatus',
                            'Spirinchus lanceolatus',
                            'Squalius carolitertii', 'Squalius pyrenaicus',
                            'Squalius alburnoides', 'Strongylura marina',
                            'Syngnathus fuscus', 'Stellifer lanceolatus',
                            'Syngnathus louisianae', 'Syngnathus floridae',
                            'Stichaeopsis nana', 'Sebastiscus marmoratus',
                            'Scomber colias', 'Stichaeopsis nevelskoi',
                            'Strongylura timucu', 'Stromateus brasiliensis',
                            'Taurocottus bergii', 'Thunnus thynnus', 
                            'Thunnus maccoyii', 'Thunnus tonggol',
                            'Thunnus atlanticus', 'Trachurus murphyi',
                            'Trachurus mediterraneus', 'Trachurus japonicus',
                            'Trachurus trachurus', 'Trachurus lathami',
                            'Trachurus picturatus', 'Tilesina gibbosa',
                            # biological control with atlatic salmon?
                            # could this be in runoff?
                            #'Tautogolabrus adspersus', 
                            'Tautoga onitis',
                            # invasive established in California
                            #'Tridentiger trigonocephalus',
                            'Ulvaria subbifurcata', 'Xenolumpenus longipterus',
                            # extinct
                            'Salmo pallayri',
                            # deep sea
                            'Embassichthys bathybius', # synonimized to microstomus
                            # arctic canada, gulf of alaska
                            'Arctogadus glacialis', 'Boreogadus saida',
                            'Myzopsetta proboscidea', 'Bryozoichthys lysimus',
                            'Salvelinus alpinus',
                            # aleutian islands
                            'Hemilepidotus zapus',
                            # Gulf of California/Baja peninsula; may be seen sampling if expanded south
                            'Gillichthys seta', 'Scomber australasicus',
                            'Liparis tunicatus', 'Merluccius angustimanus',
                            'Totoaba macdonaldi', 'Paralabrax auroguttatus',
                            'Haemulon flaviguttatum',
                            # freshwater?
                            'Morone chrysops', 'Morone americana'
                            #'Prosopium williamsoni', 'Rhinichthys cataractae'
                            )))) %>%
  # drop level on outgroups that dont matter
  mutate(found_taxa = case_when(genus %in% c('Bos', 'Anas', 'Pseudo-nitzschia') ~ genus,
                                TRUE ~ found_taxa)) %>%
    # filter out odd sequences
  filter(!(accession %in% c( # hybrids
                            'NC_082285', 'OQ656297', 'OQ656298', 
                            'OQ656299','LC145951', 'NC_027522',
                            'NC_034784', 'KP218514', 'KR059905',
                            'KY132098', 'MF958972', 'JX119017',
                            'KM401549', 'KC346973', 'KC351186',
                            # genus level IDs
                            'MT025508', 'MW181768', 'LC578888',
                            'MW181767', 'MW181766', 'MZ597980',
                            'PQ423886', 'PQ423885', 'OP931726',
                            'OP931727', 'OP931728', 'OP931732',
                            'OP931733', 'OP931734', 'OP931735',
                            'OP931736',
                            #'LC277738', # incorrectly on the list?
                            'LC741215', 'LC772864',
                            'KF668288',
                            # family level IDs
                            'NC_026537', 'OQ846303',
                            'OQ789493', 'LC036776', 'LC036775'))) %>%
   # set it up exactly as needed
  select(asv, motu, identity_percentage, evalue, found_taxa, common_name, rank, 
         genus, family, order, class, phylum, kingdom,
         freshwater_only, accession, coverage, bitscore, 
         status)



# the original pipeline assumed we were working from a matrix with taxonomy already assigned.
# Here we're going to assign to MOTUs from scratch
# First; assign all MOTUs with 100% one species
blast_assign <- blast50_updatetaxa %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, identity_percentage, .keep_all = TRUE) %>%
  group_by(asv) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
    # remove unknowns when sp % is < 100
  filter(!(sp_percent < 1 & found_taxa %in% c('Unknown bacteria', 'Unknown organism',
                                              'Unknown eukaryote', 'Unknown algae',
                                              'Unknown protist'))) %>%
  # how many blast hits
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  # remove duplicate rows by percent
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, sp_percent, .keep_all = TRUE) %>%
  # indicate keep rows with 100% same species
  mutate(one_hit = ifelse(sp_percent == 1, 'yes', 'no')) %>%
  group_by(asv, found_taxa) %>%
  ## hits with relatively rare frequency and low identity percentage
  mutate(unlikely = case_when((any(identity_percentage == 100) & any(identity_percentage <= 99.5)) |
                                (any(sp_percent <= .25) & any(identity_percentage <= 99.5)) | # check the 20% cutoff in top 10
                               (any(identity_percentage >= 99) & any(identity_percentage <= 97.5)) |
                                identity_percentage <= 96.5 ~ 'x')) %>%
  ungroup() %>%
  # remove unlikely hits
  filter(!(one_hit == 'no' & unlikely %in% c('x') & identity_percentage <= 98.5|
             identity_percentage <= 97)) %>%
  # how many hits are left?
  add_count(asv, name = 'hit_remain') %>%
  # filter for anything 100%/98% that passed the original filter
  group_by(asv) %>%
  mutate(unlikely = case_when((hit_remain > 1 & any(identity_percentage == 100) & any(identity_percentage <= 99.5)) ~ 'x')) %>%
  ungroup() %>%
  # remove unlikely hits
  filter(!(unlikely %in% c('x') & identity_percentage <= 99.5)) %>%
  # assign asvs with one hit remaining after unlikely and < 100% frequency
  # # how many hits are left?
  add_count(asv, name = 'hit_remain') %>%
  filter(hit_remain == 1) %>%
  select(!(c(hit_remain, unlikely, one_hit, sp_percent, num_blast, sp_freq)))
  
```

## Conflicts

If there isn't a single identification, use stricter criteria or drop to higher taxonomy
```{r conflicts}
# identify asvs that have conflicts in their top 10
# This removes asvs without conflicts, leaving only the potential problems.
blast_conflict <- blast50_updatetaxa %>%
  filter(!(asv %in% blast_assign$asv)) %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, identity_percentage, .keep_all = TRUE) %>%
  group_by(asv, found_taxa) %>%
  mutate(max_sp_id = max(identity_percentage)) %>%
  ungroup() %>%
  mutate(max_keep = ifelse(max_sp_id == identity_percentage, 'keep', 'toss')) %>%
  filter(max_keep == 'keep') %>%
  group_by(asv) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
      # remove unknowns when sp % is < 100
  filter(!(sp_percent < 1 & found_taxa %in% c('Unknown bacteria', 'Unknown organism',
                                              'Unknown eukaryote', 'Unknown algae',
                                              'Unknown protist'))) %>%
  # how many blast hits
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  # remove duplicate rows by percent
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, sp_percent, .keep_all = TRUE) %>%
    group_by(asv, found_taxa) %>%
  # remove hits with less than 98.5%% when an ID in the list has 100%,
  ## hits with relatively rare frequency and low identity percentage
  mutate(unlikely = case_when((any(identity_percentage == 100) & any(identity_percentage <= 99.5)) |
                                (any(sp_percent <= .2) & any(identity_percentage <= 99.5)) | # check the 20% cutoff in top 10
                               (any(identity_percentage >= 99) & any(identity_percentage <= 97.5)) |
                                 identity_percentage <= 96.5 ~ 'x')) %>%
  ungroup() %>%
  filter(!(unlikely %in% c('x') & identity_percentage <= 98 |
             identity_percentage <= 97)) %>%
  select(!(c(unlikely))) %>%
    #remove anything that's more than .5 lower
  group_by(asv) %>%
  # what is the highest identity percentage on that asv
  mutate(max_id = max(identity_percentage)) %>%
  ungroup() %>%
  # if the species max is .5% lower than the max
  # this seems to be about as far away as makes sense for this data
  # Note that maybe for other data it might be different.
  # May also need more solid empirical justification than 
  # my past experience handling NCBI data
  # Filter to the max
  mutate(point_diff = max_id - max_sp_id) %>%
  filter(point_diff < .5) %>%
  select(!(c(max_id, max_sp_id, point_diff))) %>%
    # recalculate species percent
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  group_by(asv, found_taxa) %>%
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(asv) %>%
    # na handling
  mutate(na_gen = ifelse(any(is.na(genus)), 'yes', 'no'),
         na_fam = ifelse(any(is.na(family)), 'yes', 'no'),
         na_ord = ifelse(any(is.na(order)), 'yes', 'no'),
         na_class = ifelse(any(is.na(class)), 'yes', 'no'),
         na_phy = ifelse(any(is.na(phylum)), 'yes', 'no'),
         na_king = ifelse(any(is.na(kingdom)), 'yes', 'no')) %>%
  mutate(updated_taxon = case_when(length(unique(found_taxa)) == 1 ~ found_taxa,
                                   length(unique(genus)) == 1 & na_gen == 'no' ~ genus,
                                   length(unique(family)) == 1 & na_fam == 'no' ~ family,
                                   length(unique(order)) == 1 & na_ord == 'no' ~ order,
                                   length(unique(class)) == 1 & na_class == 'no' ~ class,
                                   length(unique(phylum)) == 1 & na_phy == 'no' ~ phylum,
                                   length(unique(kingdom)) == 1  & na_king == 'no'~ kingdom,
                                   na_king == 'yes' | length(unique(kingdom)) > 1 ~ 'Unknown organism')) %>%
  select(!c(na_gen, na_fam, na_ord, na_class, na_phy, na_king, 
            num_blast, sp_percent, sp_freq)) 

blast_lowmatch <- blast50_updatetaxa %>%
  filter(!(asv %in% blast_assign$asv | asv %in% blast_conflict$asv)) %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  ## what is the highest identity percentage per species, to eliminate duplicates
  group_by(asv, found_taxa) %>%
  mutate(max_sp_id = max(identity_percentage)) %>%
  ungroup() %>%
  distinct(asv, found_taxa, max_sp_id, .keep_all = TRUE) %>%
  group_by(asv) %>%
  # what is the highest identity percentage on that asv
  mutate(max_id = max(identity_percentage)) %>%
  ungroup() %>%
  # if the species max is .5 points or about 99.5% lower than the max for the asv
  # eliminate that hit
  # Filter to the max
  mutate(point_diff = max_id - max_sp_id) %>%
  filter(point_diff < .5) %>%
  # recalculate species percent and remove 100s
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  group_by(asv, found_taxa) %>%
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(asv) %>%
  mutate(low_match = ifelse(max_id < 96, 'yes', 'no')) %>%
  mutate(very_low_match = ifelse(max_id < 93, 'yes', 'no')) %>%
  mutate(extremely_low_match = ifelse(max_id < 90.99, 'yes', 'no')) %>%
    # if the number of flagged records always has the same genus, update. 
  # if the number of flaggd records always has the same family, update.
  # if the species percent is 100% and the species was actually not local, update
    # na handling
  mutate(na_gen = ifelse(any(is.na(genus)), 'yes', 'no'),
         na_fam = ifelse(any(is.na(family)), 'yes', 'no'),
         na_ord = ifelse(any(is.na(order)), 'yes', 'no'),
         na_class = ifelse(any(is.na(class)), 'yes', 'no'),
         na_phy = ifelse(any(is.na(phylum)), 'yes', 'no'),
         na_king = ifelse(any(is.na(kingdom)), 'yes', 'no')) %>%
  mutate(updated_taxon = case_when(length(unique(found_taxa)) == 1 & low_match == 'no' ~ found_taxa,
                                   length(unique(genus)) == 1 & na_gen == 'no' & 
                                     very_low_match == 'no' ~ genus,
                                   length(unique(family)) == 1 & na_fam == 'no' & 
                                     extremely_low_match == 'no' ~ family,
                                   length(unique(order)) == 1 & na_ord == 'no' ~ order,
                                   length(unique(class)) == 1 & na_class == 'no' ~ class,
                                   length(unique(phylum)) == 1 & na_phy == 'no' ~ phylum,
                                   length(unique(kingdom)) == 1  & na_king == 'no'~ kingdom,
                                   na_king == 'yes' | length(unique(kingdom)) > 1 ~ 'Unknown organism')) %>%
  ungroup() %>%
  select(!(c(max_id, max_sp_id, point_diff, 
             accession, na_gen, na_fam, na_ord, na_class, na_phy, na_king, 
             num_blast, sp_percent, sp_freq)))

```

## all nonnative hits
this shouldbe 0

```{r}
hit_nohit <- asv_50blast %>%
  filter(!(asv %in% blast_conflict$asv | asv %in% blast_lowmatch$asv | asv %in% blast_assign$asv)) %>%
  group_by(asv) %>%
    mutate(updated_taxon = case_when(length(unique(genus)) == 1 ~ genus,
                                   length(unique(family)) == 1 ~ family)) %>%
  ungroup() %>%
  mutate(genus = case_when(family == updated_taxon ~ NA_character_,
                           TRUE ~ genus)) %>% 
  select(!c(accession, identity_percentage, evalue, 
            species, bitscore, coverage, n_mismatches)) %>%
  distinct()
```


## Update

Now we'll use updated taxa from the conflicts, and add a column for MOTU level ids based on seed asv
```{r update}
# drop uncertanties to next taxonomic level
blast_updates <- blast_conflict %>%
      # modify dataframe to be more useful
  select(asv, motu, found_taxa, updated_taxon, common_name,
         identity_percentage, coverage, evalue, 
         bitscore, rank, kingdom, phylum, class, order, 
         family, genus, freshwater_only
         ) %>%
  rbind(blast_lowmatch) %>%
  rbind(hit_nohit) %>%
  distinct() %>%
  mutate(rank = case_when(rank == 'offtarget' ~ rank,
                          str_detect(updated_taxon, ' ') == TRUE ~ 'species',
                          updated_taxon == genus ~ 'genus',
                          updated_taxon == family ~ 'family',
                          updated_taxon == order ~ 'order',
                          updated_taxon == class ~ 'class',
                          updated_taxon == phylum ~ 'phylum')) %>%
  mutate(genus = case_when(rank %in% c('family', 'order')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                           TRUE ~ genus)) %>%
  mutate(family = case_when(rank %in% c('order')| 
                             str_detect(updated_taxon, 'Unknown')== TRUE ~ NA_character_,
                            TRUE ~ family)) %>%
  mutate(order = case_when(rank %in% c('class')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ order)) %>%
  mutate(class = case_when(rank %in% c('phylum')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ class)) %>%
  mutate(phylum = case_when(rank %in% c('kingdom')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ phylum)) %>%
  mutate(kingdom = case_when(str_detect(updated_taxon, 'Unknown organism') == TRUE ~ NA_character_,
                            TRUE ~ kingdom)) %>%
  distinct(asv, updated_taxon, .keep_all = TRUE) %>%
  rbind(blast_assign) %>%
  mutate(found_taxa = coalesce(updated_taxon,found_taxa)) %>%
  select(!c(updated_taxon, coverage, bitscore,
            accession, status)) %>%
  relocate(asv, motu, found_taxa, common_name, rank) %>%
  arrange(motu, asv) %>%
  # open group somewhere?
  ungroup()
```

Now we'll use updated taxa from the conflicts, and add a column for MOTU level ids based on seed asv
```{r update}
taxonomy_assigned_asv <- motu_key %>%
  full_join(blast_updates) %>%
  # update based on research
  #mutate(found_taxa = case_when(found_taxa == 'Merluccius angustimanus' ~ 'Merluccius',
  #                              TRUE ~ found_taxa)) %>%
  mutate(rank = case_when(found_taxa %in% c('Merluccius') ~ 'genus',
                          TRUE ~ rank)) %>%
 # left_join(common_name) %>%
  #mutate(common_name = coalesce(common_name_key, common_name)) %>%
  #relocate(common_name, .before = rank) %>%
  mutate(blast = ifelse(is.na(found_taxa), 'No BLAST hit', 'BLAST')) %>%
  mutate(found_taxa = coalesce(found_taxa, blast)) %>%
  mutate(common_name = case_when(found_taxa == 'Unknown organism' ~ 'Unknown organism',
                                 ## FISH ##
                                 found_taxa == 'Teleostei' ~ 'Unknown fish',
                                 # order level
                                 found_taxa == 'Acanthuriformes' ~ 'Acanthuriform order',
                                 found_taxa == 'Gobiiformes' ~ 'Goby order',
                                 found_taxa == 'Perciformes' ~ 'Perciformes order',
                                 # family level
                                 found_taxa == 'Agonidae' ~ 'Poacher family',
                                 found_taxa == 'Cottidae' ~ 'Cottid sculpin',
                                 found_taxa == 'Embiotocidae' ~ 'Surfperch family',
                                 found_taxa == 'Gadidae' ~ 'Cod family',
                                 found_taxa == 'Liparidae' ~ 'Snailfish family',
                                 found_taxa == 'Pleuronectidae' ~ 'Flatfish family',
                                 found_taxa == 'Serranidae' ~ 'Sea bass family',
                                 found_taxa == 'Stichaeidae' ~ 'Prickleback family',
                                 # genus level
                                 found_taxa == 'Albula' ~ 'Bonefish species',
                                 found_taxa == 'Ammodytes' ~ 'Sandlance species',
                                 found_taxa == 'Clinocottus' ~ 'Clinocottus sculpin species',
                                 found_taxa == 'Cottus' ~ 'Cottus sculpin species',
                                 found_taxa == 'Enophrys' ~ 'Enophrys sculpin species',
                                 found_taxa == 'Gadus' ~ 'Gadus cod species',
                                 found_taxa == 'Gibbonsia' ~ 'Kelpfish species',
                                 found_taxa == 'Halichoeres' ~ 'Halichoeres wrasse',
                                 found_taxa == 'Hemilepidotus' ~ 'Hemilepidotus Irish lord species',
                                 found_taxa == 'Hexagrammos' ~ 'Hexagrammos kelp species',
                                 found_taxa == 'Lepidopsetta' ~ 'Lepidopsetta sole species',
                                 found_taxa == 'Leuroglossus' ~ 'Leuroglossus smoothtongue species',
                                 found_taxa == 'Liparis' ~ 'Snailfish species',
                                 found_taxa == 'Malacocottus' ~ 'Malacocottus sculpin species',
                                 found_taxa == 'Merluccius' ~ 'Hake species',
                                 found_taxa == 'Myoxocephalus' ~ 'Myoxocephalus sculpin species',
                                 found_taxa == 'Oligocottus' ~ 'Oligocottus sculpin species',
                                 found_taxa == 'Oncorhynchus' ~ 'Salmon species',
                                 found_taxa == 'Paralabrax' ~ 'Paralabrax sea bass species',
                                 found_taxa == 'Phanerodon' ~ 'Phanerodon seaperch species',
                                 found_taxa == 'Pholis' ~ 'Gunnel species',
                                 found_taxa == 'Rimicola' ~ 'Rimicola kelp clingfish species',
                                 found_taxa == 'Ruscarius' ~ 'Roughcheek sculpin species',
                                 found_taxa == 'Salvelinus' ~ 'Char species',
                                 found_taxa == 'Sebastes' ~ 'Rockfish species',
                                 found_taxa == 'Syngnathus' ~ 'Pipefish species',
                                 found_taxa == 'Thunnus' ~ 'Tuna species',
                                 found_taxa == 'Xiphister' ~ 'Prickleback species',
                                 # species level
                                 # invasive?
                                 found_taxa == 'Barbatula barbatula' ~ 'Stone loach invasive',
                                 found_taxa == 'Alosa pseudoharengus' ~ 'Alewife invasive',
                                 found_taxa == 'Esox lucius' ~ 'Northern pike invasive',
                                 found_taxa == 'Ctenopharyngodon idella' ~ 'Grass carp invasive',
                                 found_taxa == 'Gambusia affinis' ~ 'Mosquitofish invasive',
                                 found_taxa == 'Morone saxatilis' ~ 'Striped bass invasive',
                                 found_taxa == 'Squalius cephalus' ~ 'European chub no records',
                                 found_taxa == 'Tautogolabrus adspersus' ~ 'Cunner no records',
                                 found_taxa == 'Tridentiger trigonocephalus' ~ 'Chameleon goby invasive',
                                 TRUE ~ common_name)) %>%
  # some silliness
  mutate(found_taxa = case_when(found_taxa == 'Homo' ~ 'Homo sapiens',
                                TRUE ~ found_taxa)) %>%
  mutate(rank = case_when(class %in% c('Mammalia', 'Aves') ~ 'outgroup',
                          common_name %in% c('Bacteria', 'Diatom', 'Foraminifera',
                                             'Unknown organism') ~ 'outgroup',
                          TRUE ~ rank)) %>%
  #filter(is.na(common_name)) %>%
  #arrange(common_name)
  arrange(found_taxa) %>%
  distinct()
```

# consensus for MOTUs

add a column for motu level ids, this way we can select columns based on the analysis we want to do

```{r motus}
taxonomy_assigned_motu <- taxonomy_assigned_asv %>%
  filter(blast == 'BLAST') %>%
  group_by(motu) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(motu,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  group_by(motu, common_name) %>%
  # how often does that species come up
  add_count(name = 'com_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  mutate(com_percent = com_freq/num_blast) %>%
    # indicate keep rows with 100% same species
  mutate(one_hit = ifelse(sp_percent == 1, 'yes', 'no')) %>%
  mutate(one_com = ifelse(com_percent == 1 & rank == 'offtarget', 'yes', 'no')) %>%
  mutate(hit_100_higher = ifelse(identity_percentage == 100 & !(rank %in% c('species', 'offtarget')), 1, NA_character_)) %>%
  group_by(motu) %>%
  # if any species has 100% and the rest are other ranks without 100%
  # go with highest ranking species
  mutate(sp_100 = ifelse((sum(hit_100_higher < 1, na.rm = TRUE) | is.na(sum(hit_100_higher < 1))) & any(one_hit == 'no' & rank != 'species' & rank != 'offtarget'),
         found_taxa[which.max(identity_percentage) & rank == 'species'], NA_character_)) %>%
  # all hits genus or higher, just collapse to highest match
  mutate(higher_taxa = ifelse(all(rank != 'species'), found_taxa[which.max(sp_percent)], NA_character_)) %>%
  # if family level has 100%
 # mutate(fam_100 = case_when(any(one_hit == 'no' & rank == 'family' & hit_100_higher == 1) ~ found_taxa[which.max(identity_percentage) & rank == 'family'],
#                             any(blast == 'No BLAST hit') ~ found_taxa[which.max(identity_percentage) & rank == 'family'])) %>%
  mutate(fam_100 = ifelse(any(one_hit == 'no' & rank == 'family' & hit_100_higher == 1),
         found_taxa[which.max(identity_percentage) & rank == 'family'], NA_character_)) %>%
  # genus level has 100%
  mutate(gen_100 = ifelse(any(one_hit == 'no' & rank == 'genus' & hit_100_higher == 1),
         found_taxa[which.max(identity_percentage) & rank == 'genus'], NA_character_)) %>%
  # species level conflict, expand later to be one family etc.
  mutate(sp_conflict = case_when(length(unique(genus)) == 1 & one_hit == 'no' ~ 'one genus',
                                 length(unique(family))== 1 & one_hit == 'no' ~ 'one family')) %>%
  #mutate(sp_conflict = ifelse(length(unique(genus)) == 1 & one_hit == 'no', 'one genus', NA_character_)) %>%
  ungroup() %>%
  mutate(found_taxa_motu = case_when(one_hit == 'yes' ~ found_taxa,
                                     one_com == 'yes' ~ common_name,
                                     one_com == 'no' & rank == 'offtarget' ~ 'Unknown organism',
                                     !is.na(fam_100) ~ fam_100,
                                     !is.na(gen_100) ~ gen_100,
                                     !is.na(sp_100) ~ sp_100,
                                     !is.na(higher_taxa) ~ higher_taxa,
                                     sp_conflict == 'one genus' ~ genus,
                                     sp_conflict == 'one family' ~ family)) %>%
    select(!c(one_hit, one_com, sp_100, higher_taxa, fam_100, gen_100, sp_conflict,
            low_match, very_low_match, extremely_low_match, blast,
            sp_percent, num_blast, sp_freq, com_freq, hit_100_higher, com_percent)) %>%
  relocate(found_taxa_motu) %>%
  rename(common_name_asv = common_name,
         found_taxa_asv = found_taxa,
         rank_asv = rank) %>%
  group_by(motu) %>%
  mutate(common_name = ifelse(any(found_taxa_asv == found_taxa_motu), common_name_asv[which(found_taxa_asv == found_taxa_motu)], NA_character_)) %>%
    mutate(common_name = ifelse(any(common_name_asv == found_taxa_motu), common_name_asv[which(common_name_asv == found_taxa_motu)], common_name)) %>%
  ungroup() %>%
  mutate(common_name = case_when(!is.na(common_name) ~ common_name,
                                 found_taxa_motu == 'Unknown organism' ~ 'Unknown organism',
                                 # ugh just manual for now
                                 found_taxa_motu == 'Oncorhynchus' ~ 'Oncorhynchus salmon species')) %>%
  # rank now
  group_by(motu) %>%
    mutate(rank = ifelse(any(found_taxa_asv == found_taxa_motu), rank_asv[which(found_taxa_asv == found_taxa_motu)], NA_character_)) %>%
      mutate(rank = ifelse(any(common_name_asv == found_taxa_motu), rank_asv[which(common_name_asv == found_taxa_motu)], rank)) %>%
  ungroup() %>%
  mutate(rank = case_when(!is.na(rank) ~ rank,
                          found_taxa_motu == 'Unknown organism' ~ 'offtarget',
                          # for expansion later
                          found_taxa_motu == genus ~ 'genus')) %>%
  relocate(common_name, rank, .before = motu) %>%
  rename(found_taxa = found_taxa_motu)


motu_test <- taxonomy_assigned_motu %>%
  select(motu, found_taxa, common_name) %>%
  distinct()


write_csv(taxonomy_assigned_motu, here(processededna, 
                                       paste0(datefield_new, runfield,'taxonomy.csv')))
```



```{r not-in}
motu_reblast <- taxonomy_assigned_asv %>%
  mutate(to_reblast = case_when(found_taxa %in% c('No BLAST hit') ~ 'x',
                                extremely_low_match == 'yes' & rank != 'offtarget' ~ 'x',
                                very_low_match == 'yes' & rank != 'offtarget' ~ 'x',
                                rank %in% c('order', 'family') ~ 'x',
                                rank == 'genus' & is.na(common_name) ~ 'x')) %>%
  filter(!is.na(to_reblast)) %>%
  select(motu) %>%
  distinct()

write_delim(motu_reblast, here(processededna, paste0(datefield_new, runfield, 'to-reblast2.txt')),
          col_names = FALSE)
```

