---
title: "ch2_00f_lulu"
author: "Kate Sheridan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)

library(lulu)

fastq <- here('rawdata', 'peco', 'fastq', 'miseq_nextseq', 'combined_next_mi')
edna_out <- here('processeddata', 'peco', 'nextseq_miseq')

datefield <- '20260331_'
runfield <- 'peco_next-mi-combined_'
```

# load in 
```{r loadin}
motu2asv <- read_csv(here(fastq, '20260331_peco_next-mi-combined_motu2asv.csv'))%>%
  distinct()

asv_seq_table <- read_delim(here(fastq, 
                                 "20260331_peco_next-mi-combined_sequence-table_asv-names.txt"))
```


# set up dataframes for lulu
```{r prep}
# join data to filter into the required dfs
motuxsample <-  asv_seq_table %>%
  # pivot to long so sequence can map to ASV
  pivot_longer(cols = !row_names,
               names_to = 'asv') %>%
  filter(value >= 1) %>% 
  left_join(motu2asv) %>%
  rename(reads_raw = value,
         sample_id = row_names) %>%
  mutate(sample_id = tolower(sample_id)) %>%
  select(sample_id, reads_raw, motu) %>%
  pivot_wider(names_from = sample_id,
              values_from = (reads_raw),
              values_fn = sum) %>%
    # NA to 0
  mutate_all(~replace(., is.na(.), 0)) %>%
  column_to_rownames('motu')
```


We'll want to use the already-swarmed stuff, but we need our cluster representatives to match the motu names

From github:
First produce a blastdatabase with the OTUs
makeblastdb -in OTU_sequences.fasta -parse_seqids -dbtype nucl
 Then blast the OTUs against the database
blastn -db OTU_sequences.fasta -outfmt '6 qseqid sseqid pident' -out match_list.txt -qcov_hsp_perc 80 -perc_identity 84 -query OTU_sequences.fasta

```{r}
otu_sequences <- file.path(fastq,
                           '20260331_peco_next-mi-combined_swarm_clusterreps.fasta')
blast_out <- file.path(fastq, paste0(datefield, runfield, 'lulu_matchlist.txt'))

# check its working
system2("blastn", args = "-h")
system2("makeblastdb", args = "-h")

# first produce a blast database with the otus
system2("makeblastdb", args=c("-in", otu_sequences,
                         "-parse_seqids",
                         "-dbtype", 'nucl'
                         ))

system2("blastn", args=c("-db", otu_sequences, # directory of db to blast against
                         "-outfmt", "'6 qseqid sseqid pident'",
                         "-qcov_hsp_perc", 80,
                         "-perc_identity", 84,
                         "-query", otu_sequences,
                         "-out", blast_out
                         ))

motu_matchlist <- read_delim(here(fastq, paste0(datefield, runfield, 'lulu_matchlist.txt')),
                           delim = '\t', col_names = c("motu_query", "motu_hit",  "identity_percentage"))
```

Now that we have both parts we can try lulu

We can look at the settings more in depth at some point!

```{r}
curated_result <- lulu(motuxsample, motu_matchlist)

# make sure the motus are in a column to merge in later steps
curated_table <- curated_result$curated_table %>%
  rownames_to_column('motu')
motu_map <- curated_result$otu_map %>%
  rownames_to_column('motu')

discard <- data.frame(curated_result$discarded_otus) %>%
  rename(discarded = `curated_result.discarded_otus`)

curated_result

curated_result$minimum_relative_cooccurence #.95
curated_result$minimum_match # 84

motu2asv2 <- motu2asv %>%
  left_join(motu_map) %>%
  rename(motu_swarm = motu,
         motu_lulu = parent_id,
         lulu_rank = rank) %>%
  relocate(motu_lulu, motu_swarm) %>%
  select(!c(seed_asv, total, spread, asv_reads))
```

# Export
```{r}
# I don't think we need to write this out since we're going to be processing year by year
#write_csv(curated_table, here(edna_out, paste0(datefield, runfield, 'motuxsample_lulu.csv')))
#write_csv(motu_map, here(fastq, paste0(datefield, runfield, 'lulu_motu-map.csv')))

# key for QC
write_csv(motu2asv2, here(fastq, paste0(datefield, runfield, 'lulu_motu2asv.csv')))

# TODO later make into a txt with commas
# Used to remove from fasta
write_csv(discard, here(fastq, paste0(datefield, runfield, 'removed-motus.csv')))
```

Now we need to remove the discarded motus from the fasta