---
title: "ch2_00b_run-combine-2024"
output: html_document
date: "2024-06-04"
---


Combines sequence tables for 2022 nextseq, 2022 miseq, and 2023 nextseq both runs. 

Sequence tables
Input: sequence tables from each run
Output: sequence table per year, master sequence table from all years

Read retention
Input: read retention log from each run
Output: read retention by year and total


NOTE ON FILEPATHS:
datefield by year is the date the run was last processed -- not today's date! That's regular `datefield`!

```{r setup, include=FALSE}
####Libraries####
library(tidyverse)
library(dada2)
library(seqinr)

library(here)
# here filepaths
# combined
fastq_combined <- here('rawdata', 'peco', 'fastq', 'miseq_nextseq', 'combined_next_mi')
datefield <- '20260404_'
runfield <- 'peco_next-mi-combined_'


# miseq
pecofastq2022_mi <- here('rawdata', 'peco', 'fastq', 
                  'miseq_nextseq', 'miseq_2022_run2',
                  'fixed', 'processed') 
runfield2022_mi <- 'peco2022_miseq-library2_'

# nextseq
pecofastq2022_next <- here('rawdata', 'peco', 'fastq', 
                  'miseq_nextseq', 'nextseq_2022_run2',
                  'fixed', 'processed') 
runfield2022 <- 'peco2022_nextseq-library2_'

# nextseq
pecofastq2022_udi <- here('rawdata', 'peco', 'fastq', 
                  '2022') 
runfield2022_udi <- 'peco2022_nextseq-udi_'


# 2023 next-first
pecofastq2023 <- here('rawdata', 'peco', 'fastq', 
                  'miseq_nextseq', '2023', 'processed') 
datefield2023 <- '20240603_'
runfield2023 <- 'peco2023_nextseq-firstrun_'

# 2023 next-redo
pecofastq2023_re <- here('rawdata', 'peco', 'fastq', 
                  'miseq_nextseq', '2023', 'processed') 
datefield2023_re <- '20250208_'
runfield2023_re <- 'peco2023_nextseq-rerun_'
```

# eDNA data

## load in each year


```{r each year}
####read in sequence tables####
# miseq 2022
seqtab_2022_mi <- read_delim(here(pecofastq2022_mi,
                                 "20260331_peco2022_miseq-library2_sequence-table_12S-merged.txt")) %>%
  mutate(row_names = paste0(row_names, '-miseq')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()



# nextseq 2022
seqtab_2022_next <- read_delim(here(pecofastq2022_next, 
                                 '20260331_peco2022_nextseq-library2_sequence-table_12S-merged.txt')) %>%
  mutate(row_names = paste0(row_names, '-nextseq')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

# nextseq 2022 udi
seqtab_2022_udi <- read_delim(here(pecofastq2022_udi, 
                                 '20240210_peco2022_12s_nextseq_sequence-table.txt')) %>%
  mutate(row_names = paste0(row_names, '-nextseq-udi')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()



# nextseq-first 2023
seqtab_2023 <- read_delim(here(pecofastq2023,
                               "20240603_peco2023_12s_20240328_sequence-table_12S-merged.txt")) %>%
  mutate(row_names = paste0(row_names, '-first')) %>%
 column_to_rownames('row_names') %>%
 as.matrix()

#peco 2024
seqtab_2023_re <- read_delim(here(pecofastq2023_re,
                               "20250428_peco2023_12s_20240907_sequence-table_12S-merged.txt")) %>%
  mutate(row_names = paste0(row_names, '-redone')) %>%
 column_to_rownames('row_names') %>%
 as.matrix()
```

# Combine eDNA

All sequence tables should be merged into one master table. This means that every identical sequence in every year will be given the same ASV number. This master combined table should be exported and run through seqtable2swarm.py before moving on to clustering and taxonomy assignment.

```{r merge-all-tables}
####merge sequence tables####
#ASVs are combined on sequence alone. exact matches are simply merged
seqtab_all <- mergeSequenceTables(seqtab_2022_mi, seqtab_2022_next, seqtab_2022_udi,
                                  seqtab_2023, seqtab_2023_re) 


####save sequences as fasta, give sequences human-readable names, save modified sequence table####


#saving the new names and sequences as a .fasta file in your project working directory, and save a table that shows the mapping of sequences to new ASV names
asvtab_all <- t(as.data.frame(seqtab_all)) 
#transposed (OTUs are rows) data frame. 
# unclassing the otu_table() output avoids type/class errors later on

#re-save sequence table w updated names
write.table(data.frame("row_names"=rownames(seqtab_all), seqtab_all),
            here(fastq_combined,
                 paste0(datefield, runfield, "sequence-table.txt")), row.names=FALSE, quote=F, sep="\t")

ASV.seq <- as.character(unclass(row.names(asvtab_all))) #store sequences in character vector
ASV.num <- paste0("ASV", seq(ASV.seq), sep='') #create new names

asv_map <- cbind(asv = ASV.num, seq = ASV.seq) %>%
  as.data.frame()

write.table(asv_map, here(fastq_combined, 
                          paste0(datefield, runfield,
                                 "asv-sequences_name-mapping.txt")),
            sep="\t", quote=F, row.names=F, col.names=F)



write.fasta(sequences=as.list(ASV.seq), names=ASV.num, 
            here(fastq_combined, 
                 paste0(datefield, runfield, "asv-sequences.fasta"))) #save sequences with new names in fasta format



#IMPORTANT: sanity checks
colnames(seqtab_all) == ASV.seq #only proceed if this tests as true for all elements

#rename your ASVs in the taxonomy table and sequence table objects
colnames(seqtab_all) <- ASV.num

#re-save sequence table w updated names
write.table(data.frame("row_names"=rownames(seqtab_all),seqtab_all),
            here(fastq_combined,
                 paste0(datefield, runfield, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")

```

## ASV tables by year

These will be used for eDNA QC after taxonomy assignment.

```{r yearly sequence tables}
asvtab_2022_mi <- seqtab_2022_mi %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2022_mi,
            here(pecofastq2022_mi,
                 paste0(datefield, runfield, "sequence-table_asv-names_2022-miseq.txt")), 
            row.names=FALSE, quote=F, sep="\t")

# 2022
asvtab_2022_next <- seqtab_2022_next %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2022 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2022_next,
            here(pecofastq2022_next,
                 paste0(datefield, runfield, "sequence-table_asv-names_2022-nextseq.txt")), 
            row.names=FALSE, quote=F, sep="\t")


# 2022
asvtab_2022_udi <- seqtab_2022_udi %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2022 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2022_udi,
            here(pecofastq2022_next,
                 paste0(datefield, runfield, "sequence-table_asv-names_2022-nextseq-udi.txt")), 
            row.names=FALSE, quote=F, sep="\t")


# 2023
asvtab_2023 <- seqtab_2023 %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2023 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2023,
            here(pecofastq2023,
                 paste0(datefield, runfield, "sequence-table_asv-names_2023-first.txt")), 
            row.names=FALSE, quote=F, sep="\t")


# 2024
asvtab_2023_re <- seqtab_2023_re %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2024 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2023_re,
            here(pecofastq2023_re,
                 paste0(datefield, runfield, "sequence-table_asv-names_2023-redone.txt")), 
            row.names=FALSE, quote=F, sep="\t")

```



