---
title: "ch2-by-yearqc"
author: "Kate Sheridan"
date: "`r Sys.Date()`"
output: html_document
---

Right now this primarily identifies what would be removed with read subtraction.

It also generates files for plotting from the metadata, and semiquantitive metrics.


# Needed for all
```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(PNWColors)
library(conflicted)

# for phyloseq / SQ
library(phyloseq)
library(DESeq2)
library(pheatmap)
library(vsn)
library(ape)

# conflicts from packages
conflicts_prefer(dplyr::rename())
conflicts_prefer(dplyr::filter)

# here filepaths
# In
fastq <- here('rawdata', 'peco', 'fastq', 'miseq_nextseq') 
# Out
processededna <- here('processeddata', 'peco', 'nextseq_miseq')
# fields
datefield <- '20260331_'
runfield <- 'peco_next-mi-combined_'

# Palette
# manual with 2 oregon regions
pal_peco <- c("Juneau Region, Alaska" = "#00496F",
             "Southeast Alaska" = "#03587B",
             "Haida Gwaii, British Columbia" = "#076787",
             "Central Coast British Columbia" = "#0B7693",
             "North Vancouver Island, British Columbia" = "#0F85A0",
             "Inner Coast, British Columbia" = "#469989",
             "Pacific Rim Park, British Columbia" = "#7EAE73",
             "Puget Sound & Padilla Bay, Washington" = "#B5C25C",
             "Willapa Bay, Washington" = "#EDD746",
             "Yaquina Bay, Oregon" = "#EDC334",
             "Coos Bay, Oregon" = "#EDB123",
             "Humboldt Bay, California" = "#ED9E11",
             "Bodega Bay, California" = "#ED8B00",
             "San Francisco Bay, California" = "#E97809",
             "Morro Bay, California" = "#E56512",
             "Los Angeles, California" = "#E1531B",
             "San Diego region, California" = "#DD4124")
```
In: taxonomy from BLAST, asv table, metadata
```{r loadin}
metadata <- read_csv(here(processededna, '20260319_peco-2021-24_metadata.csv')) %>%
  mutate(extraction_blank = str_replace_all(extraction_blank, '_', '-'))


taxonomy <- read_csv(here(processededna, '20260331_peco_next-mi-combined_taxonomy.csv'))

motu2asvs <- read_csv(here(fastq, 'combined_next_mi',
                           '20260331_peco_next-mi-combined_lulu_motu2asv.csv'))
```

# load-in

We'll prep these to merge, then QC the merged files
```{r}
# 2022
asvtable_2022_mi <- read_delim(here(fastq, 'miseq_2022_run2',
                                 'fixed', 'processed',
                            '20260331_peco_next-mi-combined_sequence-table_asv-names_2022-miseq.txt'))

motumatrix_metadata_2022_mi <- asvtable_2022_mi %>%
  # pivot to long so sequence can map to ASV
  pivot_longer(cols = !sample,
               names_to = 'asv') %>%
  filter(value >= 1) %>% 
  mutate(sample = tolower(sample)) %>%
  mutate(sample_id = str_remove(sample, '-miseq')) %>%
  left_join(taxonomy) %>%
  mutate(motu_pa = 1) %>%
  rename(reads_raw = value) %>%
  left_join(metadata) %>%
  # remove columns that don't matter this year
  janitor::remove_empty() %>%
  # check before removing
  select(!c(sample)) %>%
  mutate(library = '2022miseq') %>%
  mutate(year = 2022) %>%
  mutate(seqmethod = 'miseq')

# 2022
asvtable_2022_next <- read_delim(here(fastq, 'nextseq_2022_run2',
                                 'fixed', 'processed',
                            '20260331_peco_next-mi-combined_sequence-table_asv-names_2022-nextseq.txt'))

motumatrix_metadata_2022 <- asvtable_2022_next %>%
  # pivot to long so sequence can map to ASV
  pivot_longer(cols = !sample,
               names_to = 'asv') %>%
  filter(value >= 1) %>% 
  mutate(sample = tolower(sample)) %>%
  mutate(sample_id = str_remove(sample, '-miseq')) %>%
  left_join(taxonomy) %>%
  mutate(motu_pa = 1) %>%
  rename(reads_raw = value) %>%
  left_join(metadata) %>%
  # remove columns that don't matter this year
  janitor::remove_empty() %>%
  # check before removing
  select(!c(sample)) %>%
  mutate(library = '2022nextseq') %>%
  mutate(year = 2022) %>%
  mutate(seqmethod = 'nextseq')

# 2023
asvtable_2023 <- read_delim(here(fastq, '2023', 'processed',
                            '20260331_peco_next-mi-combined_sequence-table_asv-names_2023-first.txt'))

motumatrix_metadata_2023 <- asvtable_2023 %>%
  # pivot to long so sequence can map to ASV
  pivot_longer(cols = !sample,
               names_to = 'asv') %>%
  filter(value >= 1) %>% 
  mutate(sample = tolower(sample)) %>%
  mutate(sample_id = str_remove(sample, '-first')) %>%
      mutate(sample_id = case_when(str_detect(sample_id, 'pecoe_8') ~ 
                                 str_replace_all(sample_id, 'pecoe_8', 'pecoe-08'),
                                 str_detect(sample_id, 'pecoe_9') ~ 
                                 str_replace_all(sample_id, 'pecoe_9', 'pecoe-09'),
                                 str_detect(sample_id, 'pecoe_1') ~ 
                                 str_replace_all(sample_id, 'pecoe_1', 'pecoe-1'),
                                 str_detect(sample_id, 'neg_pcr_1_2023') ~ 
                                 str_replace_all(sample_id, 'neg_pcr_1_2023', 'negpcr-2023-1'),
                                 str_detect(sample_id, 'neg_pcr_3_2023') ~ 
                                 str_replace_all(sample_id, 'neg_pcr_3_2023', 'negpcr-2023-3'),
                                 str_detect(sample_id, 'neg_pcr_4_2023') ~ 
                                 str_replace_all(sample_id, 'neg_pcr_4_2023', 'negpcr-2023-4'),
                                 str_detect(sample_id, 'pecoe_xn') ~ 
                                 str_replace_all(sample_id, 'pecoe_xn', 'exblank-2023-'),
                               TRUE ~ sample_id)) %>%
  select(!sample) %>%
  filter(sample_id %in% metadata$sample_id) %>%
  left_join(taxonomy) %>%
  mutate(motu_pa = 1) %>%
  rename(reads_raw = value) %>%
  left_join(metadata) %>%
    # remove columns that don't matter this year
  janitor::remove_empty() %>%
  mutate(library = '2023first') %>%
  mutate(year = 2023)


# 2023
asvtable_2023_re <- read_delim(here(fastq, '2023', 'processed',
                            '20260331_peco_next-mi-combined_sequence-table_asv-names_2023-redone.txt'))

motumatrix_metadata_2023_re <- asvtable_2023_re %>%
  # pivot to long so sequence can map to ASV
  pivot_longer(cols = !sample,
               names_to = 'asv') %>%
  filter(value >= 1) %>% 
  mutate(sample = tolower(sample)) %>%
  mutate(sample_id = str_remove(sample, '-redone')) %>%
      mutate(sample_id = case_when(str_detect(sample_id, 'pecoe-8') ~ 
                                 str_replace_all(sample_id, 'pecoe-8', 'pecoe-08'),
                                 str_detect(sample_id, 'pecoe-9') ~ 
                                 str_replace_all(sample_id, 'pecoe-9', 'pecoe-09'),
                               TRUE ~ sample_id)) %>%
  select(!sample) %>%
  filter(sample_id %in% metadata$sample_id) %>%
  left_join(taxonomy) %>%
  mutate(motu_pa = 1) %>%
  rename(reads_raw = value) %>%
  left_join(metadata) %>%
    # remove columns that don't matter this year
  janitor::remove_empty() %>%
  mutate(library = '2023redone') %>%
  mutate(year = 2023) %>%
  mutate(seqmethod = 'nextseq')

```

## merge


```{r}
motumatrix_metadata_merged <- motumatrix_metadata_2022_mi %>%
  bind_rows(motumatrix_metadata_2022) %>%
  bind_rows(motumatrix_metadata_2023) %>%
  bind_rows(motumatrix_metadata_2023_re) %>%
  mutate(treatment = case_when(str_detect(sample_id, 'exblank') ~ 'Extraction blank',
                               str_detect(sample_id, 'pcrblank') ~ 'PCR blank',
                               TRUE ~ treatment)) %>%
  filter(!is.na(found_taxa))
```



## Lab controls

Subtracting reads from pcr / extraction controls

labs; 
extraction negative -- group = extraction date
calibration -- group == library. samples that were run twice, could look at library level differences in read number etc.
pcr negative -- group == library
Field blank -- group = site
libraries: by run

Need to make sure I have info about which blanks are for which samples.
```{r control-subtract}
pcr_ctrls <- motumatrix_metadata_merged %>%
  filter(treatment %in% c('PCR blank')) %>%
  select(asv, reads_raw, treatment, library, found_taxa, rank) %>%
  # make single code that combines site and asv
  mutate(pcrasv = paste0(library, '_', asv)) %>%
  mutate(is_fish_lab = ifelse(!(rank == 'outgroup'), 'yes', NA_character_)) %>%
  select(pcrasv, reads_raw, found_taxa, is_fish_lab) %>%
  rename(reads_control = reads_raw)


extract_ctrls <- motumatrix_metadata_merged %>%
  filter(treatment %in% c('Extraction blank')) %>%
  select(asv, reads_raw, treatment, sample_id, found_taxa, rank) %>%
  mutate(extraction_blank = sample_id) %>%
  # make single code that combines site and motu
  mutate(extractasv = paste0(extraction_blank, '_', asv)) %>%
  mutate(is_fish_lab = ifelse(!(rank == 'outgroup'), 'yes', NA_character_)) %>%
  select(extractasv, reads_raw, found_taxa, is_fish_lab) %>%
  rename(reads_control = reads_raw)

motumatrix_metadata_nolab <- motumatrix_metadata_merged %>%
  # no controls
  filter(treatment %in% c('Sample', 'Field blank')) %>%
  # insert minus part here
  # subtract extraction controls
  # single code to combine site and motu
  mutate(extractasv = paste0(extraction_blank, '_', asv)) %>%
  # merge to have field control values for each site-motu combo
  left_join(extract_ctrls) %>%
  mutate(in_ctrl_extract = ifelse(reads_control > 0, 'yes', NA_character_)) %>%
  #subtract control reads from regular reads
  mutate(reads_adjust = reads_raw - reads_control) %>%
  mutate(reads_adjust = coalesce(reads_adjust, reads_raw)) %>%
  mutate(reads_adjust = ifelse(reads_adjust <= 0, 0, reads_adjust)) %>%
  # only samples that still have reads
  #filter(reads_adjust > 0) %>%
  select(!(c(reads_control, extractasv))) %>%
  # subtract pcr controls by library
  # single code to combine library and motu
  mutate(pcrasv = paste0(library, '_', asv)) %>%
  # merge to have field control values for each site-motu combo
  left_join(pcr_ctrls) %>%
  mutate(in_ctrl_pcr = ifelse(reads_control > 0, 'yes', NA_character_)) %>%
  #subtract control reads from regular reads
  mutate(reads_adjust_lab = reads_adjust - reads_control) %>%
  mutate(reads_adjust_lab = coalesce(reads_adjust_lab, reads_adjust)) %>%
  mutate(reads_adjust_lab= ifelse(reads_adjust_lab <= 0, 0, reads_adjust_lab)) %>%
  # only samples that still have reads
  #filter(reads_adjust_pcr > 0) %>%
  select(!(c(reads_control, reads_adjust, pcrasv))) %>%
  #rename(reads_raw = reads) %>%
  # n site code = lab control
  filter(!(is.na(site_code))) %>%
  mutate(ctrl_lab0 = ifelse(reads_adjust_lab == 0, 'yes', 'no')) %>%
  mutate(in_ctrl_lab = ifelse(in_ctrl_extract == 'yes' | in_ctrl_pcr == 'yes', 'yes', NA_character_)) %>%
  relocate(in_ctrl_lab, in_ctrl_pcr, in_ctrl_extract, reads_adjust_lab, ctrl_lab0, .before = kingdom) %>%
  distinct()

# NOTE we have some info about which field controls are also in extraction blanks;
# do we want to take this into consideration?

field_ctrls <- motumatrix_metadata_nolab %>%
  filter(treatment == 'Field blank') %>%
  select(site_code, year, asv, reads_raw, found_taxa, ctrl_lab0, rank) %>%
  # make single code that combines site and motu
  mutate(siteasv = paste0(as.character(year), '_', site_code, '_', asv)) %>%
  mutate(is_fish_field = ifelse(!(rank == 'outgroup'), 'yes', NA_character_)) %>%
  select(siteasv, reads_raw, found_taxa, is_fish_field#, ctrl_lab0
         ) %>%
  rename(reads_control = reads_raw)

motumatrix_metadata_nofield <- motumatrix_metadata_nolab %>%
  filter(treatment == 'Sample') %>%
  # single code to combine site and motu
  mutate(siteasv = paste0(as.character(year), '_', site_code, '_', asv)) %>%
  # merge to have field control values for each site-motu combo
  left_join(field_ctrls) %>%
  #subtract control reads from regular reads
  mutate(reads_adjust_fieldlab = reads_adjust_lab - reads_control) %>%
  mutate(reads_adjust_fieldlab = coalesce(reads_adjust_fieldlab, reads_adjust_lab)) %>%
  mutate(reads_adjust_fieldlab = ifelse(reads_adjust_fieldlab <= 0, 0, reads_adjust_fieldlab)) %>%
  mutate(ctrl_fieldlab0 = ifelse(reads_adjust_fieldlab == 0, 'yes', 'no')) %>%
  mutate(in_ctrl_field = ifelse(reads_control > 0, 'yes', 'no')) %>%
  # only samples that still have reads
  #filter(reads_adjust > 0) %>%
  #subtract control reads from raw reads
  mutate(reads_adjust_field = reads_raw - reads_control) %>%
  mutate(reads_adjust_field = coalesce(reads_adjust_field, reads_raw)) %>%
  mutate(reads_adjust_field = ifelse(reads_adjust_field <= 0, 0, reads_adjust_field)) %>%
  select(!(c(reads_control, siteasv))) %>%
  #relocate(reads_adjust, .before = reads_raw) %>%
  mutate(ctrl_field0 = ifelse(reads_adjust_field == 0, 'yes', 'no')) %>%
  group_by(sample_id) %>%
  mutate(ctrl_lab_fish = ifelse(any(is_fish_lab == 'yes'), 'yes', NA_character_)) %>%
  mutate(ctrl_field_fish = ifelse(any(is_fish_field == 'yes'), 'yes', NA_character_)) %>%
  ungroup() %>%
  relocate(in_ctrl_field, reads_adjust_field, ctrl_field0, 
           ctrl_fieldlab0, reads_adjust_fieldlab, 
           ctrl_lab_fish, ctrl_field_fish, .before = reads_adjust_lab) %>%
  relocate(sample_id, motu, found_taxa, common_name, rank, year) %>%
  select(!c(is_fish_lab, is_fish_field))
```

# fish and ranges

```{r export}
# filter to fish, remove redone samples
motumatrix_fish_qcd <- motumatrix_metadata_nofield %>%
  # we only assign fish ranks, everything else is 'outgroup'
  #filter(!(rank == 'outgroup')) %>%
  # remove duplicates from redo (follow up on how we want to do about it)
  # right now picked from which one has more unique species
  mutate(remove = case_when(sample_id %in% c('pecoe-1672', 'pecoe-0985',
                                             'pecoe-1395', 'pecoe-1485') & library == 'redo_2025oct' ~ 'x',
                            sample_id %in% c('pecoe-1420') & library == '2025may' ~ 'x')) %>%
  filter(is.na(remove)) %>%
  janitor::remove_empty() %>%
  # remove columns we don't need later
  select(!c(evalue, identity_percentage, kingdom, phylum,
            freshwater_only)) %>%
  # in any control
  mutate(asv_year = paste0(asv, '-', year)) %>%
  # identifying which contain high amounts of fish seen in any control that year
  # not filtering by this yet
  #mutate(any_ctrl = ifelse(asv_year %in% c(all_ctrls$asv_year), 'yes', NA_character_)) %>%
  mutate(any_ctrl = coalesce(in_ctrl_field, in_ctrl_lab)) %>%
  # from jenn's filtering
  group_by(site_code, year) %>%
  mutate(field_blank_richness_site= sum(in_ctrl_field == 'yes', na.rm = TRUE)) %>%
  mutate(lab_richness_site = sum(in_ctrl_lab =='yes', na.rm = TRUE)) %>%
  mutate(richness_site_avs_anyctrl=sum(any_ctrl == 'yes', na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(sample_id, year) %>%
  mutate(sample_richness = n_distinct(asv)) %>%
  mutate(richness_sample_avs_in_labblank = sum(in_ctrl_lab =='yes', na.rm = TRUE)) %>%
  mutate(richness_sample_avs_anyctrl=sum(any_ctrl == 'yes', na.rm = TRUE)) %>%
  mutate(richness_sample_avs_in_fieldblank=sum(in_ctrl_field == 'yes', na.rm = TRUE)) %>%
  mutate(richness_proportion_in_fieldblank=richness_sample_avs_in_fieldblank/sample_richness) %>%
  mutate(richness_proportion_in_labblank=richness_sample_avs_in_labblank/sample_richness) %>%
  mutate(richness_proportion_in_anyblank=richness_sample_avs_anyctrl/sample_richness) %>%
  ungroup() %>%
  group_by(region_name, year) %>%
  mutate(field_blank_regional_max= max(field_blank_richness_site)) %>%
  ungroup() %>%
  ## this carries with it an assumption of how many ASVs is reasonable to expect in a sample
  ## made histogram, assigned in first "real" gap at a binwidth = 5; between 61 and 68
  ## NOTE when added all asvs, this changed to 70 vs 80+
  ## this may need to change in future, or be reassessed
  ## also included lab here with high proportion of lab blanks
  ## and re-added high proportion of sample richness in field blanks to catch ones with < 5 asvs
  mutate(flag = case_when(sample_richness > 70 ~ "richness suspiciously high",
                          # here the bacterial diversity in field blanks can be whatever it wants if it doesn't have a fish in it
                          richness_sample_avs_in_fieldblank > 5 & ctrl_field_fish == 'yes' ~ "high sample richness in field blanks",
                          richness_proportion_in_fieldblank >= 0.5 ~ "high proportion sample richness in field blanks",
                          richness_proportion_in_labblank >= 0.5 ~ "high proportion sample richness in lab blanks",
                          richness_proportion_in_anyblank >= 0.5 ~ "high proportion sample richness across all blanks"
                          )) %>%
  #relocate(any_ctrl) %>%
  # NOW filter to fish
  filter(!(rank == 'outgroup'))
 

#summarize the richness in every sample
sample_summaries<- motumatrix_fish_qcd %>%
  select(year, 
         region_name, 
         site_code, 
         sample_id, 
         sample_richness, 
         richness_sample_avs_in_fieldblank,
         richness_sample_avs_in_labblank,
         richness_sample_avs_anyctrl,
         field_blank_richness_site, 
         lab_richness_site,
         richness_site_avs_anyctrl,
         field_blank_regional_max,
         richness_proportion_in_labblank,
         richness_proportion_in_fieldblank, 
         richness_proportion_in_anyblank,
         flag) %>%
  distinct()

ggplot(sample_summaries) +
  geom_histogram(aes(x = sample_richness),
                 binwidth = 5)


motumatrix_fish <- motumatrix_fish_qcd %>%
  filter(is.na(flag))

testfish <- motumatrix_fish_qcd %>%
  filter(!is.na(any_ctrl)) %>%
  select(year, region_name, found_taxa, common_name,
         asv) %>%
  distinct()

#motumatrix_fish_flagged <- motumatrix_fish_qcd %>%
#  filter(!is.na(flag))

# note; do we want to reduce duplicates in any way?
# our mean might be weighted towards frequent sampling?
new_range <- motumatrix_fish_qcd %>%
  # updated to be just taxa vs lat/lon, could be expanded to year?
  select(found_taxa, decimal_latitude, decimal_longitude) %>%
  distinct() %>%
  group_by(found_taxa) %>%
  summarise(taxa_max = max(decimal_latitude, na.rm = TRUE), taxa_mean = mean(decimal_latitude, na.rm = TRUE),
            taxa_median = median(decimal_latitude, na.rm = TRUE), taxa_min = min(decimal_latitude, na.rm = TRUE))

motumatrix_fish_range <- motumatrix_fish_qcd %>%
  left_join(new_range) %>%
  arrange(year, sample_id)
  
  
# to check which classes
#unique(motumatrix_metadata_nofield$class)

# check for duplicate samples from redos
# should be 0 rows!
sample_dups <- motumatrix_fish %>%
  select(sample_id, library) %>%
  distinct() %>%
  add_count(sample_id) %>%
  relocate(n) %>%
  filter(n > 1)

write.csv(motumatrix_fish_range, here(processededna, paste0(datefield, runfield, 'asvmatrix_fish-novst.csv')))
```



Might not do this for that.



# phyloseq 

Phyloseq objects contain otu table, sample_data, and taxonomy table among other things.

What is the best level to normalize our data at? ; species, motu, asv, genus?

I'm doing both MOTU and taxa, so that the output can be used for different analysis.
We can do this to other levels if we want.

library and year are correlated too much currently to be separate terms as most have only one library per year.

vst or rlog, probably vst for the time factor
assay to extract values

```{r phyloseq motus}
# import otu table
# requires numeric matrix
fishonly_otus <- motumatrix_fish_range %>%
  select(sample_id, motu, reads_raw) %>%
  pivot_wider(names_from = sample_id,
              values_from = reads_raw,
              values_fn = sum) %>%
  # NA to 0
  mutate(across(.cols = everything(), ~replace(., is.na(.), 0))) %>%
  column_to_rownames('motu') %>%
  as.matrix()

fishonly_sample <- motumatrix_fish_range %>%
  select(sample_id, year, region_code, library) %>% 
  distinct() %>%
  # make sure sample names match otu table
  column_to_rownames('sample_id') %>%
  filter(rownames(.) %in% colnames(fishonly_otus)) %>%
  # needs to be a factor
  mutate(year = as.factor(year)) %>%
  # NOTE factor CANNOT include special characters
  mutate(library = as.factor(str_remove_all(library, '-')))

# this is from the deseq function, 
# if we want to keep in the "bad" samples we need to add a pseudocount of 1 to all,
# and tbh we might need this simply because eDNA is not RNAseq 
# and every species will fail to show up in at least one sample.
# The same tweak as for edgeR to avoid NaN problems
# that cause the workflow to stall/crash.
fishonly_otus = fishonly_otus + 1

fishotu_dsdata = DESeqDataSetFromMatrix(countData = fishonly_otus, 
                             colData=fishonly_sample, 
                             design = ~ library
                             )

# note it mentions it uses local for fittype; look into this
vsd_otu <- varianceStabilizingTransformation(fishotu_dsdata)

## Taxa level

# import otu table
# requires numeric matrix
fishonly_taxa_otus <- motumatrix_fish_range %>%
  select(sample_id, found_taxa, reads_raw) %>%
  pivot_wider(names_from = sample_id,
              values_from = reads_raw,
              values_fn = sum) %>%
  # NA to 0
  mutate(across(.cols = everything(), ~replace(., is.na(.), 0))) %>%
  column_to_rownames('found_taxa') %>%
  as.matrix()

fishonly_taxa_sample <- motumatrix_fish_range %>%
  select(sample_id, library, region_code) %>%
  #mutate(treatment_deseq = case_when(is.na(treatment) ~ 'Sample',
   #                            treatment == 'Field blank' ~ 'field_blank',
    #                           TRUE ~ treatment)) %>%
  distinct() %>%
  # make sure sample names match otu table
  column_to_rownames('sample_id') %>%
  filter(rownames(.) %in% colnames(fishonly_taxa_otus)) %>%
  # NOTE factor CANNOT include special characters
  mutate(library = as.factor(str_remove_all(library, '-')))

# this is from the deseq function, 
# if we want to keep in the "bad" samples we need to add a pseudocount of 1 to all,
# and tbh we might need this simply because eDNA is not RNAseq 
# and every species will fail to show up in at least one sample.
# The same tweak as for edgeR to avoid NaN problems
# that cause the workflow to stall/crash.
fishonly_taxa_otus = fishonly_taxa_otus + 1

fishtaxa_dsdata = DESeqDataSetFromMatrix(countData = fishonly_taxa_otus, 
                             colData=fishonly_taxa_sample, 
                             design = ~ library)

vsd_taxa <- varianceStabilizingTransformation(fishtaxa_dsdata)

# plots
meanSdPlot(fishonly_otus)
meanSdPlot(assay(vsd_otu))

pheatmap(fishonly_otus, 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Un-relativized MOTUs')
pheatmap(assay(vsd_otu), 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'MOTUs with VST, DESeq2')


meanSdPlot(fishonly_taxa_otus)
meanSdPlot(assay(vsd_taxa))

pheatmap(fishonly_taxa_otus, 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Un-relativized taxa')
pheatmap(assay(vsd_taxa), 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Taxa with VST, DESeq2')

```

## attach to data
```{r}
vst_taxa_tomerge <- as.data.frame(assay(vsd_taxa)) %>%
  rownames_to_column('found_taxa') %>%
  pivot_longer(cols = !found_taxa,
               names_to = 'sample_id',
               values_to = 'relabund_vst_taxa') %>%
  mutate(relabund_vst_taxa = ifelse(relabund_vst_taxa < 0, 0, relabund_vst_taxa))

vst_otu_tomerge <- as.data.frame(assay(vsd_otu)) %>%
  rownames_to_column('motu') %>%
  pivot_longer(cols = !motu,
               names_to = 'sample_id',
               values_to = 'relabund_vst_motu') %>%
  mutate(relabund_vst_motu = ifelse(relabund_vst_motu < 0, 0, relabund_vst_motu))


fishonly_relabund <- motumatrix_fish_range %>%
  left_join(vst_taxa_tomerge) %>%
  left_join(vst_otu_tomerge)%>%
  relocate(relabund_vst_taxa, relabund_vst_motu,
           motu_pa, region_name, region_name_label, region_code, 
           region_code_label, site_code, 
           location_name, decimal_latitude, decimal_longitude,
           treatment, asv, .before = reads_raw) %>%
  relocate(ctrl_lab0, reads_adjust_field, .before = reads_adjust_fieldlab) %>%
  relocate(pcr_blank, extraction_blank, library, .before = in_ctrl_pcr) %>%
  relocate(year, rank, .before = motu) %>%
  relocate(class, order, family, genus, .before = relabund_vst_taxa) %>%
  bind_rows(motumatrix_fish_flagged) %>%
  janitor::remove_empty()

write.csv(fishonly_relabund, here(processededna, paste0(datefield, runfield, 'asvmatrix_fish.csv')))
```


## output of vst

Todo: improve output plots later
```{r}
sample_data <- fishonly_sample %>%
  rownames_to_column('sample_id')

fishonly_taxa_dist <- dist(t(fishonly_taxa_otus))
vsd_taxa_dist <- dist(t(assay(vsd_taxa)))

fishonly_taxa_dist2 <- as.matrix(fishonly_taxa_dist)
vsd_taxa_dist2 <- as.matrix(vsd_taxa_dist)


colors <- colorRampPalette( rev(RColorBrewer::brewer.pal(9, "Blues")) )(255)
prevst <- pheatmap(fishonly_taxa_dist2,
         col = colors)$gtable
ggsave(here('figures', 'controls',  paste0(datefield, 'taxa_pre-vst.png')), plot = prevst)
postvst <- pheatmap(vsd_taxa_dist2,
         col = colors)$gtable
ggsave(here('figures', 'controls',  paste0(datefield, 'taxa_post-vst.png')), plot = postvst)

fishonly_pcoa <- pcoa(fishonly_taxa_dist2)
fishonly_pcoa_plot <- as.data.frame(fishonly_pcoa$vectors) %>% 
  janitor::clean_names() %>%
  select(axis_1, axis_2) %>%
  rownames_to_column('sample_id') %>%
  left_join(sample_data)

vsd_taxa_pcoa <- pcoa(vsd_taxa_dist2)
vsd_taxa_pcoa_plot <- as.data.frame(vsd_taxa_pcoa$vectors) %>% 
  janitor::clean_names() %>%
  select(axis_1, axis_2) %>%
  rownames_to_column('sample_id') %>%
  left_join(sample_data)

  

ggplot(fishonly_pcoa_plot, aes(x = axis_1, y = axis_2)) +
  geom_point(aes(x = axis_1, y = axis_2,
                 color = as.factor(year))) +
  labs(title = 'Unadjusted, by taxa') +
  theme_bw()

ggsave(here('figures', 'controls',  paste0(datefield, 'unadjusted_pcoa.png')))
#ggsave(here('figures', 'controls',  'unadjusted_pcoa.pdf'))

ggplot(vsd_taxa_pcoa_plot, aes(x = axis_1, y = axis_2)) +
  geom_point(aes(x = axis_1, y = axis_2,
                 color = as.factor(year))) +
  labs(title = 'VST with DESeq2, by taxa') +
  theme_bw()

ggsave(here('figures', 'controls',  paste0(datefield, 'vst_pcoa.png')))
#ggsave(here('figures', 'controls',  'vst_pcoa.pdf'))

```


# Combined regions for plotting


```{r set up for plotting}
regions <- motumatrix_fish_range %>%                                                           
  #regions with site code for organisation
  select(site_code, region_code, region_name,  region_name_label, 
         region_code_label, decimal_latitude, decimal_longitude) %>%
  filter(!is.na(site_code)) %>%
  distinct(site_code, .keep_all = TRUE) %>%
  mutate(region_code = as.factor(region_code)) %>%
  mutate(region_name = as.factor(region_name)) 



# sort regions by latitude to direct color vector
region_by_lat <- regions %>%                                                   
  select(region_name,
          decimal_latitude)%>%
  distinct()%>%
   #organising the region names by latitude
  mutate(region_name = as.factor(region_name)) %>%
  #temp removing NA
  filter(!(is.na(region_name))) %>%
  filter(!(is.na(decimal_latitude))) %>%
  #requires calculating avg latitude of each 
  group_by(region_name) %>%                                                     
  summarise(avg_region_lat = mean(decimal_latitude))%>%                          
  # arranging the region names in the data 
  arrange(desc(avg_region_lat)) %>%
  ungroup() %>%
  # this gives them the numbers for the color vector
  rowid_to_column('region_lat_name')

# site + sample count
# this is by partner, can be summed for total in region_name
region_counts <- motumatrix_fish_range %>%                                                           
  #regions with site code for organisation
  select(site_code, region_code, 
         sample_id)%>%
  distinct() %>%
  filter(!(is.na(site_code))) %>%
  group_by(region_code) %>%
  add_tally(name = 'samples') %>%
  ungroup() %>% 
  select(!(sample_id)) %>%
  distinct() %>%
  group_by(region_code) %>%
  add_tally(name = 'sites') %>%
  select(!(site_code)) %>%
  distinct()

regions <- regions %>%
  left_join(region_by_lat) %>%
  left_join(region_counts) %>%
  filter(!(is.na(region_name))) %>%
  arrange(-decimal_latitude)


write.csv(regions, here(processededna, paste0(datefield,'peco_region-name-plotting.csv')))
```

# for data dictionary

write out columns of each spreadsheet

```{r}
region_cols <- regions %>%
  mutate(across(.cols = everything(), ~ as.character(.))) %>%
  # this gives them the numbers for the color vector
  rowid_to_column('temp_for_pivot') %>%
  pivot_longer(cols = !temp_for_pivot) %>%
  select(!c(value, temp_for_pivot)) %>%
  distinct() %>%
  mutate(year = 'region_cols')

motu_cols <- fishonly_relabund %>%
  mutate(across(.cols = everything(), ~ as.character(.))) %>%
  pivot_longer(cols = !year) %>%
  select(!c(value, year)) %>%
  distinct() %>%
  bind_rows(region_cols)

write.csv(motu_cols, here(processededna, paste0(datefield,'peco_cols-for-dictionary.csv')))
```

