---
title: "ch1_12s-00_blast"
date: "2023-08-23"
---

Blast modified from Evan's script to use system2

```{r}
library(tidyverse)
library(taxize)
library(here)
# this includes the run folder so change as needed
calvertfastq <- here('rawdata', '2018calvert', 'fastq', 
                  '12S_Calvert2018_Re-run20210903') 
# Destination folders
processededna <- here('processeddata', '2018calvert', 'qcstages')
datefield <- '20240317_'
datefield_out <- '20240322_'
runfield <- 'calvert2018_12s_'

```


Evan's call:

mkdir blast_96_sim_LCA_besthit
blastn -task megablast -num_threads 38 -evalue 1e-5 -max_target_seqs 10 -perc_identity 96 -qcov_hsp_perc 50 -db /data/taxonomyDBs/NCBI_NT/2023-06-08/nt -outfmt '6 qseqid stitle sacc staxid pident qcovs evalue bitscore' -query 12S.PECO_combined.ASV_sequences.fasta  -out blast_96_sim_LCA_besthit/12S_ASV_sequences.length_var.blast.out




Note that you need to have a blast database set up locally; this is not an api call

```{r run-blast}
#path.blast <- file.path(here(calvertfastq, "blast_96sim"))
#if(!dir.exists(path.blast)) dir.create(path.blast)

#mifishdb <- file.path(here('rawdata', '12s_references', '2023aug_ncbi', '20290819_12s_ncbi-db'))
#mifishdb <- file.path(here('rawdata', '12s_references', 'crabs', 'blastdb', 'crabs-mifish'))
mifishdb <- file.path("/Users/katesheridan/Documents/GitHub/taxonomy_keys/database12s/crabs/blastdb/crabs-mifish")



# fasta file to blast
# if asv
edna_to_blast <- file.path(calvertfastq, 'processed',
                           '20240315_12s_2018calvert_rerun20210903_12s-ASV-sequences.fasta')
blast_out <- file.path(calvertfastq, 'processed', paste0(datefield, runfield, 'asv_blastresults.txt'))


# check its working
system2("blastn", args = "-h")

system2("blastn", args=c("-task", "'megablast'", # which kind of blast
                         #"-num_threads", 38, # for parallelizing
                         "-evalue", 1e-5,
                         "-max_target_seqs", 50,
                         "-perc_identity", 90,
                         "-qcov_hsp_perc", 50,
                         "-db", mifishdb, # directory of db to blast against
                         "-outfmt", "'6 qseqid stitle sacc staxid pident qcovs evalue bitscore'",
                         "-query", edna_to_blast,
                         "-out", blast_out
                         ))



```





#assign taxonomy with blast NT database at 96% similarity threshold

python2 ~/programs/galaxy-tool-BLAST/blastn_add_taxonomy_lite.py -i blast_96_sim_LCA_besthit/12S_ASV_sequences.length_var.blast.out -t /data/taxonomyDBs/NCBI_taxonomy/2023-06-08/rankedlineage.dmp -m /data/taxonomyDBs/NCBI_taxonomy/2023-06-08/merged.dmp -o blast_96_sim_LCA_besthit/taxonomy
cat <(head -n 1 ~/programs/galaxy-tool-lca/example/example.tabular) taxonomy_12S_ASV_sequences.length_var.blast.out > tmp
python2 ~/programs/galaxy-tool-lca/lca.species.py -i tmp -o blast_96_sim_LCA_besthit/taxonomy_table.12S.NCBI_NT.96sim.LCA+besthit.txt -b 100 -id 96 -cov 50 -t best_hit -tid 98 -tcov 80 -fh environmental,unidentified,kingdom -flh unclassified

mkdir blast_96_sim_LCA_only
python2 ~/programs/galaxy-tool-lca/lca.species.py -i tmp -o blast_96_sim_LCA_only/taxonomy_table.12S.NCBI_NT.96sim.LCA_only.txt -b 100 -id 96 -cov 50 -t only_lca -fh environmental,unidentified -flh unclassified

#cleanup
rm blast_96_sim_LCA_besthit/12S_ASV_sequences.length_var.blast.out #remove blast output without taxonomy
rm taxonomy_12S_ASV_sequences.length_var.blast.out #remove redundant file
mv tmp blast_96_sim_LCA_besthit/12S_ASV_sequences.length_var.blast.out #replace with taxonomy added blast output


cleanup

```{r cleanup-asvs}
# top 50 list
asv_50blast <- read_delim(here(calvertfastq, 'processed', paste0(datefield, runfield, 'asv_blastresults.txt')),
                           delim = '\t', col_names = c("asv", "hit_name", "accession",
                                                       "something", "identity_percentage", "coverage",
                                                       "evalue", "bitscore")) %>%
  select(!(something)) %>%
  mutate(hit_name = trimws(hit_name)) %>%
  mutate(hit_name = str_remove_all(hit_name, "(root;)")) %>%
  separate_wider_delim(hit_name, delim = ";", names = c('kingdom', 'phylum', 'class', 'order',
                                                        'family', 'genus', 'species')) %>%
  mutate(species = str_replace_all(species, "_", " "))

# motu numbers to join with the blast results
motu_key <- read_csv(here(calvertfastq, 'processed', '20240315_calvert2018_12s_motu2asv.csv')) %>%
  select(!asv_reads)

# Give blast results MOTU#s
motu_50blast2 <- asv_50blast %>%
  left_join(motu_key) %>%
  relocate(asv, motu)

```
# assign

old notes:
taxonomy assignment by accession number
 --missing for custom names, same format as lineage input
note content from db_import may need to be updated first
 open a github issue later
crabs assign_tax --input pgaout.fasta --output taxassigned.tsv --acc2tax nucl_gb.accession2taxid --taxid nodes.dmp --name names.dmp

```{r keys}
# load-in keys
taxa_key <- read_csv(here('processeddata', 'species_lists', 'keys',
                     '20240315_12s_crabs_taxonomy-key.csv')) %>%
  select(!n_dups, hybrid) %>%
  # issues to fix in new version
  mutate(found_taxa = case_when(taxonomy_search == 'Psedoliparis' ~ 'Pseudoliparis',
                                # taxonomic revision of polyptera greeni
                                taxonomy_search == 'Polyperai' ~ 'Liparis greeni',
                                found_taxa == 'Polyptera greeni' ~ 'Liparis greeni',
                                taxonomy_search %in% c('Brachymystax tsinlingensis',
                                                       'Macquaria wujalwujalensis') ~ taxonomy_search,
                                TRUE ~ found_taxa)) %>%
  # issues to fix in new version
  mutate(genus = case_when(genus == 'Kareius' ~ 'Platichthys',
                           str_detect(found_taxa, 'Lampanyctus') & 
                             genus == 'Nannobrachium' ~ 'Lampanyctus',
                           found_taxa %in% c('Liparis greeni') ~ 'Liparis',
                           found_taxa %in% c('Pseudoliparis') ~ 'Pseudoliparis',
                           TRUE ~ genus)) %>%
  mutate(family = case_when(genus %in% c('Liparis', 'Pseudoliparis') ~ 'Liparidae',
                            TRUE ~ family)) %>%
  mutate(order = case_when(family %in% c('Liparidae') ~ 'Perciformes',
                           TRUE ~ order)) %>%
  mutate(class == case_when(order %in% c('Perciformes') ~ 'Teleostei',
                            TRUE ~ class)) %>%
  mutate(phylum = case_when(class %in% c('Teleostei') ~ 'Chordata',
                            TRUE ~ phylum)) %>%
  mutate(kingdom = case_when(found_taxa %in% c('Unknown bacteria') ~ 'Bacteria',
                             kingdom %in% c('Animalia') ~ 'Eukaryota',
                             phylum %in% c('Chordata') ~ 'Eukaryota',
                             TRUE ~ kingdom)) %>%
  mutate(rank = case_when(found_taxa %in% c('Liparis greeni') ~ 'species',
                          found_taxa %in% c('Pseudoliparis') ~ 'genus',
                          TRUE ~ rank))

common_name <- read_csv(here('processeddata', 'species_lists', 'keys',
                             '20230929_12s_common-names.csv'))[-1] %>%
  rename(common_name_key = common_name)
```

top 50 taxonomy for ASV
```{r top hits}

# strip and replace higher taxonomy
# remove out of region and hybrids
asv50blast_updatetaxa <- motu_50blast2 %>%
  # remove columns that will be duplicated
  select(!c(kingdom, phylum, class, order, family, genus, species)) %>%
  left_join(taxa_key) %>%
  # remove known incorrect hits when good hits exist (from manual check)
  # species out of region
  filter(!(c(found_taxa %in% c('Amphisticus argenteus', 'Anisotremus scapularis', 
                            'Alcichthys elongatus', 'Alburnus alburnus', 
                            'Anisotremus surinamensis', 'Aspiolucius esocinus',
                            'Acanthogobius stigmothonus', "Anisotremus virginicus",
                            'Acanthogobius hasta', 'Alosa alosa', 'Alosa fallax',
                            'Ammodytes japonicus', 'Ammodytes marinus', 'Ammodytes tobianus',
                            'Ammodytes dubius', 'Ammodytes americanus', 
                            'Argyrosomus amoyensis', 'Argyrosomus japonicus',
                            'Anchoa mitchilli', 'Atheresthes evermanni', 'Aplodinotus grunniens',
                            #'Barbatula barbatula', 'Barbatula hispanica', 
                            'Brachygenys chrysargyreum', 'Ballerus sapa',
                            'Bairdiella chrysoura',
                            'Bahaba taipingensis', 'Boesemania microlepis',
                            'Batrachocottus talievi', 'Batrachocottus nikolskii', 
                            'Batrachocottus multiradiatus', 'Brachyopsis segaliensis',
                            'Chondrostoma nasus', "Clariger papillosus",
                            'Chirolophis saitone', 'Chirolophis ascanii', 
                            'Chirolophis wui', 'Chirolophis japonicus',
                            'Cleisthenes pinetorum', 'Cleisthenes herzensteini',
                            'Collichthys lucidus',
                            'Cottus gobio', 'Comephorus dybowskii', 'Cottus bairdii',
                            'Cottus pollux', 'Cottus szanaga', 'Cottus poecilopus', 
                            'Cottus czerskii', 'Cottus perifretum', 'Cottus hangiongensis',
                            'Clupea harengus', 'Conodon nobilis', 
                            'Cottocomephorus grewingki', 'Cottocomephorus grewingkii',
                            'Cottocomephorus inermis', 'Comephorus baikalensis',
                            'Cosmocampus elucens', 'Cryptacanthodes bergi',
                            'Ctenogobius boleosoma', 'Ctenogobius saepepallens',
                            'Ctenogobius shufeldti',
                            "Cryptacanthodes maculatus", 'Cynoscion nothus',
                            'Cynoscion arenarius', 'Cynoscion nebulosus',
                            'Cynoscion regalis',
                            'Ditrema jordani', 'Dicentrarchus labrax',
                            'Dicentrarchus punctatus',
                            'Ernogrammus hexagrammus', 'Eques lanceolatus',
                            'Equetus punctatus', 'Evorthodus lyricus',
                            'Gadus morhua','Gadus finnmarchicus',
                            'Gasterosteus nipponicus', 'Gasterosteus wheatlandi',
                            'Glyptocephalus stelleri', 'Gymnogobius macrognathos',
                            'Glyptocephalus cynoglossus', 'Gasterosteus nipponicus',
                            'Hyperoplus immaculatus', 'Hippoglossus hippoglossus',
                            'Hexagrammos agrammus', 'Hexagrammos otakii', 
                            'Haemulon carbonarium', 'Haemulon flavolineatum',
                            'Haemulon macrostomum', 'Haemulon plumierii',
                            'Hypomesus japonicus',
                            'Hyperoplus lanceolatus', 'Hozukius emblemarius', 
                            'Haemulon aurolineatum', 'Hypsoblennius ionthas',
                            'Hypleurochilus geminatus', 'Hypleurochilus springeri',
                            'Kasatkia memorabilis', 'Leuciscus baicalensis',
                            'Leuciscus merzbacheri', 'Leiostomus xanthurus',
                            'Liparis fabricii',
                            'Liparis miostomus', 'Liparis tessellatus', 
                            'Liparis tanakae', 'Liparis inquilinus',
                            'Leuciscus leuciscus', 'Limanda limanda', 'Limanda sakhalinensis',
                            'Lepidopsetta mochigarei', 'Liopsetta pinnifasciata', 
                            'Leuciscus idus', 'Leuciscus oxyrrhis', 'Lutjanus griseus',
                            'Leuciscus burdigalensis', 'Lutjanus argentimaculatus',
                            'Myzopsetta ferruginea', 'Microstomus achne', 'Morone saxatilis',
                            'Myoxocephalus scorpius', 'Malacocottus gibber',
                            'Melanogrammus aeglefinus', 'Menticirrhus littoralis',
                            'Miichthys miiuy', 'Netuma thalassina',
                            # there is a gbif record in alaskan panhandle
                            #'Myoxocephalus jaok',
                            'Oncorhynchus kawamurae', 'Opisthocentrus tenuis',
                            'Oncorhynchus gilae', 'Opisthocentrus ocellatus',
                            'Oreochromis niloticus', 'Occella iburia',
                            'Orthopristis chrysoptera',
                            'Parahucho perryi', 'Procottus major', 
                            'Pareques umbrosus', 'Parablennius thysanius',
                            'Parablennius yatabei',
                            'Paralichthys olivaceus', 'Paralichthys adspersus',
                            'Platichthys flesus', 'Pholis gunnellus', 
                            'Pholis nebulosa', 'Pholis crassispina', 
                            'Pholis picta', 'Peprilus medius', 
                            'Pogonias cromis', 'Pungitius laevis',
                            'Pomadasys stridens', 'Pomadasys argenteus',
                            'Pomadasys kaakan', 'Peprilus triacanthus',
                            'Peprilus burti', 'Peprilus paru',
                            'Psenopsis anomala',
                            'Rutilus rutilus', 'Scardinius erythrophthalmus',
                            'Salvelinus elgyticus', 'Salvethymus svetovidovi',
                            'Salvelinus boganidae', 'Salvelinus kuznetzovi',
                            'Salvelinus taranetzi', 'Salvlinus albus',
                            'Salvelinus schmidti', 'Salvelinus levanidovi',
                            'Salvelinus kronocius', 'Salvelinus drjagini',
                            'Salvelinus drjagini', 'Salvelinus albus',
                            'Salvelinus leucomaenis', 'Scomber scombrus',
                            'Scartella emarginata', 'Scartella cristata',
                            'Serranus scriba', 'Serranus cabrilla',
                            'Sebastes mentella', 'Sebastes hubbsi',
                            'Sebastes inermis', 'Sebastes wakiyai',
                            'Sebastes cheni', 'Sebastes pachycephalus',
                            'Sebastes joyneri', 'Sebastes baramenuke',
                            'Sebastes viviparus', 'Sebastes norvegicus',
                            'Sebastes mentella', 'Sebastes koreanus',
                            'Sprattus sprattus', 'Squalius cephalus',
                            'Squalus acanthias', 'Squalus japonicus',
                            'Squalus mitsukurii', 'Sciaenops ocellatus',
                            'Squalius carolitertii', 'Squalius pyrenaicus',
                            'Squalius alburnoides', 'Strongylura marina',
                            'Syngnathus fuscus', 'Stellifer lanceolatus',
                            'Syngnathus louisianae', 'Syngnathus floridae',
                            'Stichaeopsis nana', 'Sebastiscus marmoratus',
                            'Scomber colias', 'Stichaeopsis nevelskoi',
                            'Strongylura timucu', 'Stromateus brasiliensis',
                            'Taurocottus bergii', 'Thunnus thynnus',
                            'Trachurus picturatus', 'Tilesina gibbosa',
                            'Ulvaria subbifurcata',
                            # extinct
                            'Salmo pallayri',
                            # deep sea
                            'Embassichthys bathybius', # synonimized to microstomus
                            # arctic canada, gulf of alaska
                            'Arctogadus glacialis', 'Boreogadus saida',
                            'Myzopsetta proboscidea', 'Bryozoichthys lysimus',
                            # Gulf of California; may be seen sampling if expanded south
                            'Gillichthys seta', 'Scomber australasicus',
                            'Totoaba macdonaldi',
                            # PECO but not chapter 1
                            'Leiocottus hirundo', 'Ruscarius creaseri',
                            'Gillichthys mirabilis', 'Amphistichus argenteus',
                            'Hypomesus transpacificus', 'Lampanyctus tenuiformis',
                            'Acanthogobius flavimanus', 'Leuresthes tenuis',
                            'Syngnathus euchrous', 'Sebastes macdonaldi',
                            'Syngnathus exilis', 'Syngnathus auliscus',
                            'Orthonopias triacis', 'Myoxocephalus jaok',
                            # monospecific genus, update key
                            # causing duplicates
                            'Aulorhynchus'
                            # freshwater?
                            #'Prosopium williamsoni', 'Rhinichthys cataractae'
                            )))) %>%
  # drop level on outgroups that dont matter
  mutate(found_taxa = case_when(genus %in% c('Bos', 'Anas', 'Pseudo-nitzschia') ~ genus,
                                TRUE ~ found_taxa)) %>%
  # filter out odd hybrid sequences; flatfish hybrids
  filter(!(accession %in% c( # flatfish hybrids
                            'NC_082285', 'OQ656297', 'OQ656298', 
                            'OQ656299','LC145951',
                            # genus level IDs
                            'MT025508', 'MW181768', 'LC578888',
                            'MW181767', 'MW181766', 'MZ597980',
                            'LC277738', 'LC741215', 'LC772864',
                            # family level IDs
                            'NC_026537'))) %>%
  # set it up exactly as needed
  select(asv, motu, identity_percentage, evalue, found_taxa, common_name, rank, 
         genus, family, order, class, phylum, kingdom, worms_sciname,
         freshwater_only, accession, coverage, bitscore, 
         status, unacceptreason, seed_asv) 


# Here we're going to assign to MOTUs from scratch
# First; assign all MOTUs with 100% one species
blast_assign <- asv50blast_updatetaxa %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, identity_percentage, .keep_all = TRUE) %>%
  group_by(asv) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
    # remove unknowns when sp % is < 100
  filter(!(sp_percent < 1 & found_taxa %in% c('Unknown bacteria', 'Unknown organism',
                                              'Unknown eukaryote', 'Unknown algae',
                                              'Unknown protist'))) %>%
  # how many blast hits
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  # remove duplicate rows by percent
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, sp_percent, .keep_all = TRUE) %>%
  # indicate keep rows with 100% same species
  mutate(one_hit = ifelse(sp_percent == 1, 'yes', 'no')) %>%
  group_by(asv, found_taxa) %>%
  ## hits with relatively rare frequency and low identity percentage
  mutate(unlikely = case_when((any(identity_percentage == 100) & any(identity_percentage <= 99.5)) |
                                (any(sp_percent <= .25) & any(identity_percentage <= 99.5)) | # check the 20% cutoff in top 10
                               (any(identity_percentage >= 99) & any(identity_percentage <= 97.5)) |
                                identity_percentage <= 96.5 ~ 'x')) %>%
  ungroup() %>%
  # remove unlikely hits
  filter(!(one_hit == 'no' & unlikely %in% c('x') & identity_percentage <= 98.5)) %>%
  # how many hits are left?
  add_count(asv, name = 'hit_remain') %>%
  # filter for anything 100%/98% that passed the original filter
  group_by(asv) %>%
  mutate(unlikely = case_when((hit_remain > 1 & any(identity_percentage == 100) & any(identity_percentage <= 99.5)) ~ 'x')) %>%
  ungroup() %>%
  # remove unlikely hits
  filter(!(unlikely %in% c('x') & identity_percentage <= 99.5)) %>%
  # assign asvs with one hit remaining after unlikely and < 100% frequency
  # # how many hits are left?
  add_count(asv, name = 'hit_remain') %>%
  filter(hit_remain == 1) %>%
  select(!(c(hit_remain, unlikely, one_hit, sp_percent, num_blast, sp_freq)))
  
```


```{r conflicts}
# identify ASVs that have conflicts in their top 50
# This removes ASVs without conflicts, leaving only the potential problems.
blast_conflict <- asv50blast_updatetaxa %>%
  filter(!(asv %in% blast_assign$asv)) %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, identity_percentage, .keep_all = TRUE) %>%
  group_by(asv) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
      # remove unknowns when sp % is < 100
  filter(!(sp_percent < 1 & found_taxa %in% c('Unknown bacteria', 'Unknown organism',
                                              'Unknown eukaryote', 'Unknown algae',
                                              'Unknown protist'))) %>%
  # how many blast hits
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(asv,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(asv, found_taxa) %>%
  mutate(max_sp_id = max(identity_percentage)) %>%
  ungroup() %>%
  mutate(max_keep = ifelse(max_sp_id == identity_percentage, 'keep', 'toss')) %>%
  filter(max_keep == 'keep') %>%
  # remove duplicate rows by percent
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(asv, found_taxa, sp_percent, .keep_all = TRUE) %>%
    group_by(asv, found_taxa) %>%
  # remove hits with less than 98.5%% when an ID in the list has 100%,
  ## hits with relatively rare frequency and low identity percentage
  mutate(unlikely = case_when((any(identity_percentage == 100) & any(identity_percentage <= 99.5)) |
                                (any(sp_percent <= .2) & any(identity_percentage <= 99.5)) | # check the 20% cutoff in top 10
                               (any(identity_percentage >= 99) & any(identity_percentage <= 97.5)) |
                                 identity_percentage <= 96.5 ~ 'x')) %>%
  ungroup() %>%
  filter(!(unlikely %in% c('x') & identity_percentage <= 98 |
             identity_percentage <= 97)) %>%
  select(!(c(unlikely))) %>%
    #remove anything that's more than .5 lower
  #distinct(asv, found_taxa, max_sp_id, .keep_all = TRUE) %>%
  group_by(asv) %>%
  # what is the highest identity percentage on that asv
  mutate(max_id = max(identity_percentage)) %>%
  ungroup() %>%
  # if the species max is .5% lower than the max
  # this seems to be about as far away as makes sense for this data
  # Note that maybe for other data it might be different.
  # May also need more solid empirical justification than 
  # my past experience handling NCBI data
  # Filter to the max
  mutate(point_diff = max_id - max_sp_id) %>%
  filter(point_diff < .5) %>%
  select(!(c(max_id, max_sp_id, point_diff))) %>%
    # recalculate species percent
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  group_by(asv, found_taxa) %>%
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(asv) %>%
    # na handling
  mutate(na_gen = ifelse(any(is.na(genus)), 'yes', 'no'),
         na_fam = ifelse(any(is.na(family)), 'yes', 'no'),
         na_ord = ifelse(any(is.na(order)), 'yes', 'no'),
         na_class = ifelse(any(is.na(class)), 'yes', 'no'),
         na_phy = ifelse(any(is.na(phylum)), 'yes', 'no'),
         na_king = ifelse(any(is.na(kingdom)), 'yes', 'no')) %>%
  mutate(updated_taxon = case_when(length(unique(found_taxa)) == 1 ~ found_taxa,
                                   length(unique(genus)) == 1 & na_gen == 'no' ~ genus,
                                   length(unique(family)) == 1 & na_fam == 'no' ~ family,
                                   length(unique(order)) == 1 & na_ord == 'no' ~ order,
                                   length(unique(class)) == 1 & na_class == 'no' ~ class,
                                   length(unique(phylum)) == 1 & na_phy == 'no' ~ phylum,
                                   length(unique(kingdom)) == 1  & na_king == 'no'~ kingdom,
                                   na_king == 'yes' | length(unique(kingdom)) > 1 ~ 'Unknown organism')) %>%
  select(!c(na_gen, na_fam, na_ord, na_class, na_phy, na_king, 
            num_blast, sp_percent, sp_freq, max_keep))


# low hits
blast_lowmatch <- asv50blast_updatetaxa %>%
  filter(!(asv %in% blast_assign$asv | asv %in% blast_conflict$asv)) %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  ## what is the highest identity percentage per species, to eliminate duplicates
  group_by(asv, found_taxa) %>%
  mutate(max_sp_id = max(identity_percentage)) %>%
  ungroup() %>%
  distinct(asv, found_taxa, max_sp_id, .keep_all = TRUE) %>%
  group_by(asv) %>%
  # what is the highest identity percentage on that asv
  mutate(max_id = max(identity_percentage)) %>%
  ungroup() %>%
  # if the species max is .5 points or about 99.5% lower than the max for the asv
  # Filter to the max
  mutate(point_diff = max_id - max_sp_id) %>%
  filter(point_diff < .5) %>%
  # recalculate species percent and remove 100s
  group_by(asv) %>%
  add_count(name = 'num_blast') %>%
  group_by(asv, found_taxa) %>%
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(asv) %>%
  mutate(low_match = ifelse(max_id < 96, 'yes', 'no')) %>%
    # if the number of flagged records always has the same genus, update. 
  # if the number of flaggd records always has the same family, update.
  # if the species percent is 100% and the species was actually not local, update
    # na handling
  mutate(na_gen = ifelse(any(is.na(genus)), 'yes', 'no'),
         na_fam = ifelse(any(is.na(family)), 'yes', 'no'),
         na_ord = ifelse(any(is.na(order)), 'yes', 'no'),
         na_class = ifelse(any(is.na(class)), 'yes', 'no'),
         na_phy = ifelse(any(is.na(phylum)), 'yes', 'no'),
         na_king = ifelse(any(is.na(kingdom)), 'yes', 'no')) %>%
  mutate(updated_taxon = case_when(length(unique(found_taxa)) == 1 & low_match == 'no' ~ found_taxa,
                                   length(unique(genus)) == 1 & na_gen == 'no' ~ genus,
                                   length(unique(family)) == 1 & na_fam == 'no' ~ family,
                                   length(unique(order)) == 1 & na_ord == 'no' ~ order,
                                   length(unique(class)) == 1 & na_class == 'no' ~ class,
                                   length(unique(phylum)) == 1 & na_phy == 'no' ~ phylum,
                                   length(unique(kingdom)) == 1  & na_king == 'no'~ kingdom,
                                   na_king == 'yes' | length(unique(kingdom)) > 1 ~ 'Unknown organism')) %>%
  ungroup() %>%
  select(!(c(max_id, max_sp_id, point_diff, 
             accession, na_gen, na_fam, na_ord, na_class, na_phy, na_king, 
             num_blast, sp_percent, sp_freq)))
```

## Update

```{r update blast}
# drop uncertanties to next taxonomic level
blast_updates <- blast_conflict %>%
      # modify dataframe to be more useful
  select(asv, motu, found_taxa, updated_taxon,
         identity_percentage, coverage, evalue, 
         bitscore, rank, kingdom, phylum, class, order, 
         family, genus, freshwater_only, seed_asv
         ) %>%
  rbind(blast_lowmatch) %>%
  distinct() %>%
  mutate(rank = case_when(rank == 'offtarget' ~ rank,
                          str_detect(updated_taxon, ' ') == TRUE ~ 'species',
                          updated_taxon == genus ~ 'genus',
                          updated_taxon == family ~ 'family',
                          updated_taxon == order ~ 'order',
                          updated_taxon == class ~ 'class',
                          updated_taxon == phylum ~ 'phylum')) %>%
  mutate(genus = case_when(rank %in% c('family', 'order')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                           TRUE ~ genus)) %>%
  mutate(family = case_when(rank %in% c('order')| 
                             str_detect(updated_taxon, 'Unknown')== TRUE ~ NA_character_,
                            TRUE ~ family)) %>%
  mutate(order = case_when(rank %in% c('class')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ order)) %>%
  mutate(class = case_when(rank %in% c('phylum')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ class)) %>%
  mutate(phylum = case_when(rank %in% c('kingdom')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ phylum)) %>%
  mutate(kingdom = case_when(str_detect(updated_taxon, 'Unknown organism') == TRUE ~ NA_character_,
                            TRUE ~ kingdom)) %>%
  distinct(asv, updated_taxon, .keep_all = TRUE) %>%
  rbind(blast_assign) %>%
  mutate(found_taxa = coalesce(updated_taxon,found_taxa)) %>%
  select(!c(updated_taxon, coverage, bitscore,
            accession, worms_sciname, status, unacceptreason)) %>%
  relocate(asv, motu, found_taxa, rank) %>%
  relocate(seed_asv, .before = low_match) %>%
  arrange(motu) %>%
  # open group somewhere?
  ungroup()


```

# assign

All our blast taxa should be updated now, we can assign

```{r assign2}
taxonomy_assigned_asv <- motu_key %>%
  full_join(blast_updates) %>%
  left_join(common_name) %>%
  mutate(common_name = coalesce(common_name_key, common_name)) %>%
  relocate(common_name, .before = rank) %>%
  mutate(blast = ifelse(is.na(found_taxa), 'No BLAST hit', 'BLAST')) %>%
  mutate(found_taxa = coalesce(found_taxa, blast)) %>%
  mutate(common_name = case_when(found_taxa %in% c('No BLAST hit', 'Unknown organism',
                                                   'Cyanobacteria') ~ found_taxa,
                                 found_taxa %in% c('Eukaryota') ~ 'Unknown eukaryote',
                                 found_taxa == 'Hirundo' ~ 'Swallow',
                                 ## FISH ##
                                 found_taxa == 'Amphistichus rhodoterus' ~ 'Redtail surfperch',
                                 found_taxa == 'Atheresthes stomias' ~ 'Arrow-tooth flounder',
                                 found_taxa == 'Cryptacanthodes giganteus' ~ 'Giant wrymouth',
                                 found_taxa == 'Glyptocephalus zachirus' ~ 'Rex sole',
                                 found_taxa == 'Hemilepidotus spinosus' ~ 'Brown Irish lord',
                                 found_taxa == 'Icelinus filamentosus' ~ 'Threadfin sculpin',
                                 found_taxa == 'Oncorhynchus clarkii' ~ 'Cutthroat trout',
                                 found_taxa == 'Oncorhynchus mykiss' ~ 'Rainbow trout',
                                 found_taxa == 'Oncorhynchus gorbuscha' ~ 'Pink salmon',
                                 found_taxa == 'Phanerodon vacca' ~ 'Pile perch',
                                 found_taxa == 'Scomber japonicus' ~ 'Chub mackerel',
                                 found_taxa == 'Syngnathus californiensis' ~ 'Bay pipefish',
                                 # genera
                                 found_taxa == 'Percis' ~ 'Percis poacher species',
                                 found_taxa == 'Platichthys' ~ 'Platichthys flounder species',
                                 TRUE ~ common_name)) %>%
  relocate(seed_asv, .before = low_match) %>%
  select(!c(low_match, blast, common_name_key))
  

```


```{r add-motus}
# motu column
by_motu <- taxonomy_assigned_asv %>%
  select(asv, seed_asv, found_taxa, common_name, rank) %>%
  filter(asv == seed_asv) %>%
  select(!asv) %>%
  distinct() %>%
  rename(found_taxa_motu = found_taxa,
         common_name_motu = common_name,
         rank_motu = rank) 

# add motu columns, relocate
taxonomy_assigned_full <- taxonomy_assigned_asv %>%
  left_join(by_motu) %>%
  relocate(found_taxa_motu, common_name_motu, rank_motu, 
           .before = identity_percentage) %>%
  rename(found_taxa_asv = found_taxa,
         rank_asv = rank,
         common_name_asv = common_name) %>%
  relocate(seed_asv, .before = freshwater_only)

write_csv(taxonomy_assigned_full, here(processededna, paste0(datefield_out, runfield,
                                                             'taxonomy_initial.csv')))
```


# not in blast

```{r not in blast}
asv_notfound <- taxonomy_assigned_full %>%
  filter(found_taxa_asv == 'No BLAST hit') %>%
  select(asv) %>%
  distinct()

write_delim(asv_notfound, here(processededna, paste0(datefield_out, runfield, 'no-blast-hit.txt')),
          col_names = FALSE)
```