---
title: "swarm 12s"
author: "Kate Sheridan"
---

This script runs SWARM on output from 00_run-processing. This requires

Brandt et al 2021 (https://onlinelibrary.wiley.com/doi/10.1111/1755-0998.13398), used swarm on asv tables from dada2.

Swarm was installed, then made available to users as explained on the github: https://github.com/torognes/swarm
the binary from the installation copied to /usr/local/bin, which is on `$PATH`

Here we will call swarm with `system2`

```{r}
library(tidyverse)
library(here)
# filepaths for use with here, basic files
calvertfastq <- here('rawdata', '2018calvert', 'fastq', 
                  '12S_Calvert2018_Re-run20210903', 'processed')
datefield <- '20240315_'
runfield <- 'calvert2018_12s_'
```


Now the formatting:
We need the fasta with ASVs to be
`>ASV##_readcount`

Run ch1_00b_seq-table2swarm.py using the asv table and fasta from 00a_run-processing.


```{r load-in}

# filepath to fasta formatted for swarm
swarmfasta <- file.path(here(calvertfastq, "20240315_calvert2018_12s_sequence-table_swarm-format.fasta"))
#generate filenames for output
swarmout <- file.path(here(calvertfastq, paste0(datefield, runfield, 'swarm_outputfile.txt')))
swarmstats <- file.path(here(calvertfastq, paste0(datefield, runfield, 'swarm_statsfile.tsv')))
swarmlog <- file.path(here(calvertfastq, paste0(datefield, runfield, 'swarm_log.txt')))
swarmreps <- file.path(here(calvertfastq, paste0(datefield, runfield, 'swarm_clusterreps.fasta')))
```


```{r swarm call}
system2("swarm", args = "-v")

system2("swarm", args = c(swarmfasta,
                          "-w", swarmreps,
                          "-o", swarmout,
                          "-s", swarmstats,
                          "-l", swarmlog,
                          "-f"))

```

# input swarm output to generate motus

should probably add 'curate with lulu' step before assigning motus; temp num to start?

```{r input swarm}

swarm_stat <- read_delim(here(calvertfastq, 
                             '20240315_calvert2018_12s_swarm_statsfile.tsv'),
                         # give columns names
                        col_names = c('n_in_cluster', 'motu_reads', 
                                      'seed_asv', 'seed_reads', 'n_1readamplicon',
                                      'max_iterations_to_limit', 'cumulative_steps')) 

swarm_stat <- swarm_stat %>%
  mutate(motu = paste0('motu', seq.int(from = 1, to = nrow(swarm_stat)))) %>%
  mutate(seed_asv_reads = paste0(seed_asv, '_', seed_reads))

swarm_out <- read_delim(here(calvertfastq, 
                             '20240315_calvert2018_12s_swarm_outputfile.txt'),
                        col_names = FALSE, delim = " ")


# format
swarm_out <- swarm_out %>%
  # identify seeds for taxonomy assignment ~TEMPORARY~
  mutate(seed_asv_reads = X1) %>%
  # add 'motu' numbers
  left_join(swarm_stat) %>%
  select(!(c(motu_reads, n_in_cluster, n_1readamplicon, seed_reads,
             max_iterations_to_limit, cumulative_steps, seed_asv_reads))) %>%
  #mutate(cluster = paste0('cluster', seq.int(from = 1, to = nrow(swarm_out)))) %>%
  # make long, no more NAs
  pivot_longer(cols = !c(motu, seed_asv)) %>%
  select(!name) %>%
  filter(!(is.na(value))) %>%
  # asv + reads each get their own column
  separate_wider_delim(value, delim = "_", names = c("asv", "asv_reads"))
  
  
write_csv(swarm_out, here(calvertfastq, paste0(datefield, runfield, 'motu2asv.csv')))

write_csv(swarm_stat, here(calvertfastq, paste0(datefield, runfield, 'swarm_statsfile-w-motunums.csv')))



# to do; replace asv names in fasta
```

Now we blast our swarm clusters.




