---
title: "ch1_coi-00c_merge-runs"
date: "2024-01-27"
---


Combining runs for taxonomy assignment. There are 4 COI runs.
Modified from Evan's combining_runs.dada2.12s.PECO.R and combining_runs.calvert-2018.R

Sequence tables
Input: sequence tables from each run
Output: sequence table per run, master sequence table from all runs

Read retention
Input: read retention log from each run
Output: read retention by run, and total

NOTE ON FILEPATHS:
datefield by run is the date the run was last processed -- not today's date! That's regular `datefield`!

```{r setup, include=FALSE}
library(tidyverse)
library(dada2)
library(seqinr)

library(here)
# here filepaths
# combined destination
fastq_combined <- here('rawdata', '2018calvert', 'fastq', 
                       'coi_kate', 'combined_runs')
datefield <- '20240205_'
runfield <- 'calvert2018_coi_'

# Genome Canada, September
coifastq_gcsept <- here('rawdata', '2018calvert', 'fastq', 
                  'coi_kate', 'CO1_Run1_GenomeQuebec_Sept232019') 
datefield_gcsept <- '20240205_'
runfield_gcsept <- 'calvert2018_coi_run1_gc20190923_'

# Genome Canada, October
coifastq_gcoct <- here('rawdata', '2018calvert', 'fastq', 
                  'coi_kate', 'CO1_Run1_GenomeQuebec_Oct212019') 
datefield_gcoct <- '20240205_'
runfield_gcoct <- 'calvert2018_coi_run1_gc20191021_'

# Quadra run1
coifastq_quad1 <- here('rawdata', '2018calvert', 'fastq', 
                  'coi_kate', 'CO1_Run1_Training_03062020-Quadra') 
datefield_quad1 <- '20240127_'
runfield_quad1 <- 'calvert2018_coi_run1_quadra20200603_'

# Quadra rerun3
coifastq_quad3 <- here('rawdata', '2018calvert', 'fastq', 
                  'coi_kate', 'CO1_ReRun3_04062020-Quadra') 
datefield_quad3 <- '20240127_'
runfield_quad3 <- 'calvert2018_coi_rerun3_quadra20200604_'

```

# Merge together

It gets mad if your sample names are duplicated, and since these are multiple runs of the same thing, that is the case here.


until the quadra runs make it through the run processing we'll have this shorter version thats just genome canada

```{r merge-tables}
####read in sequence tables####
# Genome canada september
seqtab_gcsept <- read_delim(here(coifastq_gcsept, 'processed',
                                 paste0(datefield_gcsept, runfield_gcsept, 
                                        "sequence-table_coi.txt"))) %>%
  filter(!(row_names == 'Undetermined')) %>%
  mutate(row_names = paste0(row_names, '_gcsept')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()


# genome canada october
seqtab_gcoct <- read_delim(here(coifastq_gcoct, 'processed',
                                 paste0(datefield_gcoct, runfield_gcoct, 
                                        "sequence-table_coi.txt"))) %>%
  filter(!(row_names == 'Undetermined')) %>%
  mutate(row_names = paste0(row_names, '_gcoct')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

# Quadra run 1
# genome canada october
seqtab_quad1 <- read_delim(here(coifastq_quad1, 'processed',
                                 paste0(datefield_quad1, runfield_quad1, 
                                        "sequence-table_coi.txt"))) %>%
  filter(!(row_names == 'Undetermined')) %>%
  mutate(row_names = paste0(row_names, '_quadra1')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

# Quadra rerun3
seqtab_quad3 <- read_delim(here(coifastq_quad3, 'processed',
                                 paste0(datefield_quad3, runfield_quad3, 
                                        "sequence-table_coi.txt"))) %>%
  filter(!(row_names == 'Undetermined')) %>%
  mutate(row_names = paste0(row_names, '_quadra3')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

####merge sequence tables####
#ASVs are combined on sequence alone. exact matches are simply merged
seqtab_all <- mergeSequenceTables(seqtab_gcsept, seqtab_gcoct, 
                                  seqtab_quad1, seqtab_quad3) 
```


# Combine eDNA

All sequence tables should be merged into one master table. This means that every identical sequence in every year will be given the same ASV number. This master combined table should be exported and run through seqtable2swarm.py before moving on to clustering and taxonomy assignment.

```{r write-out}

####save sequences as fasta, give sequences human-readable names, save modified sequence table####

#saving the new names and sequences as a .fasta file in your project working directory, and save a table that shows the mapping of sequences to new ASV names
asvtab_all <- t(as.data.frame(seqtab_all)) 
#transposed (OTUs are rows) data frame. 
# unclassing the otu_table() output avoids type/class errors later on

#re-save sequence table w updated names
write.table(data.frame("row_names"=rownames(seqtab_all), seqtab_all),
            here(fastq_combined,
                 paste0(datefield, runfield, "sequence-table.txt")), row.names=FALSE, quote=F, sep="\t")

ASV.seq <- as.character(unclass(row.names(asvtab_all))) #store sequences in character vector
ASV.num <- paste0("ASV", seq(ASV.seq), sep='') #create new names

asv_map <- cbind(asv = ASV.num, seq = ASV.seq) %>%
  as.data.frame()

write.table(asv_map, here(fastq_combined, 
                          paste0(datefield, runfield,
                                 "asv-sequences_name-mapping.txt")),
            sep="\t", quote=F, row.names=F, col.names=F)



write.fasta(sequences=as.list(ASV.seq), names=ASV.num, 
            here(fastq_combined, 
                 paste0(datefield, runfield, "asv-sequences.fasta"))) #save sequences with new names in fasta format






#IMPORTANT: sanity checks
colnames(seqtab_all) == ASV.seq #only proceed if this tests as true for all elements

#rename your ASVs in the taxonomy table and sequence table objects
colnames(seqtab_all) <- ASV.num

#re-save sequence table w updated names
write.table(data.frame("row_names"=rownames(seqtab_all),seqtab_all),
            here(fastq_combined,
                 paste0(datefield, runfield, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")

```

# By run

## ASV tables

These will be used for eDNA QC after taxonomy assignment.

```{r genome canada}
asvtab_gcsept <- seqtab_gcsept %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_gcsept,
            here(coifastq_gcsept, "processed",
                 paste0(datefield, runfield_gcsept, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")


asvtab_gcoct <- seqtab_gcoct %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_gcoct,
            here(coifastq_gcoct, "processed",
                 paste0(datefield, runfield_gcoct, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")

# quadra 1
asvtab_quad1 <- seqtab_quad1 %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_quad1,
            here(coifastq_quad1, "processed",
                 paste0(datefield, runfield_quad1, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")

# quadra 1
asvtab_quad3 <- seqtab_quad3 %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_quad3,
            here(coifastq_quad3, "processed",
                 paste0(datefield, runfield_quad3, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")

```

# Read retention summary statistics

These files map where the reads went throughout the entire QC process

I need to remake these because they were saved incorrectly; september and october have the same number and the filter and trim step wasn't updated.

```{r load-in read retention}
# GC September
retention_gcsept <- read_csv(here(coifastq_gcsept, 'qc_plots',
                                        paste0(datefield_gcsept, runfield_gcsept,
                                               "read-retention_coi_bysample.csv")))[-1] %>%
  janitor::clean_names() %>%
  mutate(run = 'Genome Canada September')

# GC October
retention_gcoct <- read_csv(here(coifastq_gcoct, 'qc_plots',
                                        paste0(datefield_gcoct, runfield_gcoct,
                                               "read-retention_coi_bysample.csv")))[-1] %>%
  janitor::clean_names() %>%
  mutate(run = 'Genome Canada October') 

# quadra1
retention_quad1 <- read_csv(here(coifastq_quad1, 'qc_plots',
                                        paste0(datefield_quad1, runfield_quad1,
                                               "read-retention_coi_bysample.csv")))[-1] %>%
  janitor::clean_names() %>%
  mutate(run = 'Quadra Run1')

# quadra1
retention_quad3 <- read_csv(here(coifastq_quad3, 'qc_plots',
                                        paste0(datefield_quad3, runfield_quad3,
                                               "read-retention_coi_bysample.csv")))[-1] %>%
  janitor::clean_names() %>%
  mutate(run = 'Quadra Run3')

```

```{r merge-retention}

retention_all <- retention_gcsept %>%
  bind_rows(retention_gcoct) %>%
  bind_rows(retention_quad1) %>%
  bind_rows(retention_quad3)

retention_summary <- retention_all %>%
  group_by(run) %>%
  summarize(n_samp = length(sample_id),
            raw = sum(raw_f) + sum(raw_r),
            after_cutadapt = sum(cutadapt_f) + sum(cutadapt_r),
            # get the rs for this on the summary
            after_filtertrim = sum(filtertrim_f) + sum(filtertrim_r),
            after_denoise = sum(denoised_f) + sum (denoised_r),
            merged = sum(merged),
            nosingletons = sum(nosingletons),
            nochimeras = sum(nochimeras)) %>%
  ungroup() %>%
  janitor::adorn_totals('row')


write.csv(retention_summary, here(fastq_combined, 
                                  paste0(datefield, runfield,
                                         "read_retention.csv")))
```

```{r extract read stats from merged files}
length(asvtab_all)

allsum <- seqtab_all %>%
  as.data.frame(.) %>%
  mutate(totalsample = rowSums(.))

sum(allsum$totalsample)
```



# Dereplicate FASTA test

idk what this does at the moment

Right now it looks like all these are unique?

```{r dereplicate}
library(ips)

asvtab_all <- t(as.data.frame(seqtab_all))

seq_list <- asvtab_all %>%
  as.data.frame() %>%
  rownames_to_column('asv_seq') %>%
  select(asv_seq) %>%
  mutate(temp_asv = paste0("tempasv_", seq(asv_seq))) %>%
  column_to_rownames('temp_asv')

write.fas(seq_list, here(fastq_combined, 
                         paste0(datefield, 'tempasvs.fasta')))
seq_list2 <- read.fas(here(fastq_combined, 
                           paste0(datefield, 'tempasvs.fasta')))

seq_align <- mafft(seq_list2)
write.fas(seq_align, here(fastq_combined, 
                          paste0(datefield, 'tempasvs_align.fasta')))

seq_derep <- dada2:::derepFasta(here(fastq_combined,
                                     paste0(datefield, 'tempasvs_align.fasta')))

uniq<-seq_derep$uniques %>%
  as.data.frame()

```

