---
title: "ch1_12s-00d_metadata-cleaning"
output: html_document
date: "2023-08-31"
---

All Chapter1 QC scripts are modified from pipelines developed by Anya Mueller and Ben Millard-Martin.

Originally metadata cleaning was part of 01-eDNAQC but because of the extent of cleaning required for this dataset and the fact it must be done multiple times (once for ASV and once for mOTUs) I have separated it to be its own script to produce a single 'master' 12S metadata that will simply be loaded in to 01-eDNAQC.

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(here)

#if different
datameta <- here("rawdata", "2018calvert")

cleanmeta <- here('processeddata', '2018calvert')
```


## 01 - Initial Cleaning

Initial cleaning includes: cleaning names of columns, creating a unique site ID. Modified from Anya's 2020 script

Also adding in coordinates and habitat information.

here you want to read in metadata
Coordinates may be separate from the rest of the metadata with site table(s)
Filenames if directly from Hakai will be a mix of . and _ separators, be careful entering
NOTE: watch COI vs CO1, some files my use a letter I vs a number 1.

```{r 01-clean-metadata, message=FALSE}
 edna_metadata <- read_csv(here::here('rawdata', '2018calvert', '12S',
                                    "Calvert_12S_metadata.csv")) %>% 
  clean_names()

coords1 <- read_csv(here::here('rawdata', '2018calvert', 'site_locations.csv')) %>% 
  clean_names()

coords2 <- read_csv(here::here('rawdata', 'hakai', 'site_data', 'hakai_nearshore_sites.csv')) %>% 
  clean_names()

seine_hab <- read_csv(here::here(datameta, 'seine', 'fabs_extract', '2021Nov_FABSMaster_habitat.csv')) %>%
  clean_names()


## basic metadata
edna_metadata <- edna_metadata %>%
  # only 2018 and negatives
  filter(!(year %in% c(2020,2019))) %>%
  # we don't have info about extraction blanks
  #filter(!(sample_type %in% c('extraction blank'))) %>%
  filter(!(sample_name %in% c('Negative-PCR_2019-1stRUN', "Negative-PCR_2019-RERUN",
                              "Blank001_1stRUN", "Blank002_1stRUN"))) %>%
  mutate(location = location) %>%
  mutate(site = paste(year,
                      location,
                      depth, sep = "_")) %>% # make unique site id
  relocate(site) %>%
  mutate(site=str_replace(site, " ", "-")) %>% # we had sites with spaces. not good for later.
  mutate(sample_id = tolower(sample_id),
         pcr_rep = tolower(pcr_rep),
         sample_name = tolower(sample_name),
         geo_location_name_old = tolower(geo_location_name_old))


## this dataset required merging two coordinate tables. 
## I merged them separately because they were very different

# this coords file is site_locations.csv from the Hakai data portal; 
# it was the ipad data collected with the samples. Not every sample is represented though.
coords1 <- coords1 %>%
  select(site_name, point_name, decimal_lat, decimal_long,
         substrate, biotic_subclass, biotic_community) %>%
  rename(geo_location_name = point_name,
         lat = decimal_lat,
         long = decimal_long)

# add coords
edna_metadata <- edna_metadata %>%
  full_join(coords1)

# This is from a generic hakai nearshore sites list to fill in 
# the few sites not represented in the ipad data
coords2 <- coords2 %>%
  select(!(c(starts_with("x"), established, survey_frequency))) %>%
  filter(!(is.na(site_name))) %>%
  mutate(adjacent_habitat = str_replace(adjacent_habitat, "-", "")) %>%
  mutate(site_name = tolower(site_name)) %>%
  rename(location = site_name,
         lat = decimal_lat,
         long = decimal_long) %>%
  mutate(hakai_project = ifelse(project_name == "Seagrass", "seagrass", 
                                ifelse(
    project_name == "Beach Seining", "beach_seine", 
    ifelse(project_name %in% c('Kelp - Nereo', 'Kelp - Macro', 'Kelp - Rocky Reef') ,"kelp" ,NA_character_)
  ))) %>%
  select(!(c(project_name, adjacent_habitat)))

edna_metadata <- edna_metadata %>%
  left_join(coords2, by = c("location", "hakai_project"))

# here the coordinates were added as a separate row as they are not the same for 
# sites already represented in coords1. So coalesce only fills the NA's, then we delete the columns.
edna_metadata <- edna_metadata %>%
  mutate(lat.x = coalesce(lat.x, lat.y),
         long.x = coalesce(long.x, long.y)) %>%
  select(!(c(lat.y, long.y, geo_location_name_old, site_name))) %>%
  rename(lat = lat.x,
         long = long.x)


#getting habitat data in to compare

# note this is only for seine
seine_hab <- seine_hab %>%
  select(!(c(lat, long, aspect, notes))) %>%
  rename(location = site)


edna_metadata <- left_join(x = edna_metadata, y = seine_hab, by = "location")


# lots of unnecessary columns from seine_hab
cols_to_remove <- c('primary_substrate', 'intertidal_secondary_substate', 'subtidal_secondary_substrate',
                    'intertidal_primary_substrate', 'subtidal_primary_substrate', 'biotic_subclass', 
                    'biotic_community', 'primary_habitat_class', 'primary_vegetation', 
                    'intertidal_primary_macroveg', 'intertidal_secondary_macroveg', 'subtidal_primary_macroveg',
                    'subtidal_primary_macroveg_simple', 'subtidal_secondary_macroveg', 'intertidal_primary_cover',
                    'intertidal_secondary_cover', 'subtidal_primary_cover', 'subtidal_secondary_cover', 'veg_subst',
                    'slope_intertidal', 'slope_subtidal')

# cleaning habitat info
# make combined columns, recode variables, variable with combined groups, remove bonus columns
edna_metadata <- edna_metadata %>%
  mutate(substrate = coalesce(substrate, primary_substrate),
         macroveg_group = coalesce(biotic_subclass, primary_vegetation),
         macroveg_community = coalesce(biotic_community, primary_vegetation)) %>% 
  mutate(across(macroveg_group, ~recode(., algae = 'macroalgae',
                                        `seagrass meadow` = 'seagrass',
                                        `benthic macroalgae` = 'macroalgae',
                                        sand = 'none',
                                        `no vegetation` = 'none'))) %>%
  mutate(across(macroveg_community, ~recode(., unknown = 'algae',
                                            `Zostera marina` = 'zostera',
                                            `Macrocystis pyrifera` = 'macrocystis',
                                            `Nereocystis luetkeana` = 'nereocystis',
                                            sand = 'none'))) %>%
  mutate(veg_sub = paste0(macroveg_group, "_", substrate),
         vegcom_sub = paste0(macroveg_community, "_", substrate)) %>%
  select(!(all_of(cols_to_remove))) %>%
  filter(!(is.na(site)))

# save cleaned and updated metadata
write.csv(edna_metadata, here::here(cleanmeta, "2018calvert_eDNA_metadata.csv"))
```

# table to export

```{r}
library(gt)
library(gtExtras)



meta_table <- edna_metadata %>%
  select(location, macroveg_group, bottom_depth_m, lat, long) %>%
  distinct() %>%
  filter(!(is.na(location) | is.na(macroveg_group))) %>%
  group_by(location) %>%
  mutate(max_depth_m = max(bottom_depth_m),
         Latitude = mean(lat),
         Longitude = mean(long)) %>%
  ungroup() %>%
  select(!c(bottom_depth_m, lat, long)) %>%
  mutate(Resampled = ifelse(location %in% c('wfb', 'ssp', 'ppo',
                                            'pba', 'hdo', 'chp'), 'x', ' ')) %>%
  relocate(Resampled, .before = Latitude) %>%
  distinct() %>%
  rename(Vegetation = macroveg_group,
         Depth_m = max_depth_m) %>%
  arrange(Vegetation, location) %>%
  gt(rowname_col = 'location')
  
meta_table

options(chromote.headless = "new")

gtsave(meta_table, filename = here('output', 'other', '20250320_ch1_sitetable.pdf'))
```

