---
title: "ch1eDNAQC-coi"
author: "Kate Sheridan"
date: "11/27/2021"
output: html_document
---
Version of Sunday Lab eDNA QC pipeline modified from Trawl Project 2020, scripts by Anya Mueller, Jenn Sunday, Ben Millard-Martin

Main difference with this version is that it is eDNA-only, with no cleaning of trawl related files, and Hakai samples.

August 2022 update: updated taxonomy fix to key replacement
January 2024 update: split script to be one for each run, reflecting the needs of each run.

Here I will QC on ASV and in the combining scripts filter to MOTU for the comparisons.

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(vegan)
library(here)
library(taxize)

# run output
datacoi <- here("rawdata", "2018calvert", "fastq",
                "coi_kate", "CO1_ReRun3_04062020-Quadra",
                "processed")

#if different
datameta <- here('rawdata', '2018calvert', 'COI')

# Destination folders
processededna <- here("rawdata", "2018calvert", "fastq",
                "coi_kate", "combined_runs")

cleanmeta <- here('processeddata', '2018calvert')

datefield <- '20240220_'
# note i need to update this in the past scripts to match 12s: this is the goal
runfield <- 'calvert2018_coi_quadrarerun3_'

```

## 01 - Initial Cleaning

Initial cleaning includes: cleaning names of columns, creating a unique site ID, cleaning eDNA data to be more usable. Modified from Anya's 2020 script

Also adding in coordinates and habitat information.

here you want to read in any markers and metadata
These will be sequence + ASV merged tables, saved as a .txt file, with metadata as .csv

Filenames if directly from Hakai will be a mix of . and _ separators, be careful entering
NOTE: watch COI vs CO1, some files my use a letter I vs a number 1. 

```{r 01-initialcleaning-loadin, message=FALSE}
## read in files
coi_asvmatrix <- read.table(here(datacoi,
                                 "20240205_coi_2018calvert_rerun3_quadra20200604_sequence-table_asv-names.txt"),
                            header = TRUE)

# COI metadata was separate, but we will add columns from the 12s metadata that we already cleaned for habitat etc.
## see ch1_12s-01_eDNAQC script for details
coi_metadata <- read.csv(here(datameta,
                              'eDNA_Calvert2018_metadata_RevisedMay2020.csv')) %>% 
  clean_names()

edna_metadata <- read_csv(here(cleanmeta, "2018calvert_eDNA_metadata.csv"))[-1] %>%
  select(!(c(sample_id, pcr_rep))) %>%
  distinct()
```

# Metadata

```{r 01-metadata}
# extract needed columns to merge
coi_metadata2 <- coi_metadata %>%
  select(sample_id, sample_name, pcr_rep, 
         run_id, pool) %>%
  left_join(edna_metadata)

# save cleaned and updated metadata
# export only if needed
#write.csv(edna_metadata, here::here(cleanmeta, "2018calvert_eDNA_metadata_coi.csv"))

# split into relevant run
meta_run3 <- coi_metadata2 %>%
  filter(run_id == "Run_3_redo") %>%
  # update controls with pool numbers
  mutate(sample_name = case_when(sample_name %in% c('Neg', 'POS') & 
                                   pool == 'Run3_Pool22' ~ paste0(sample_name, '22'),
                                 sample_name %in% c('Neg', 'POS') & 
                                   pool == 'Run3_Pool44' ~ paste0(sample_name, '44'),
                                 sample_name %in% c('Neg', 'POS') & 
                                   pool == 'Run3_Pool66' ~ paste0(sample_name, '66'),
                                 TRUE ~ sample_name)) %>%
  mutate(pcr_rep = ifelse(pcr_rep == '', NA_character_, pcr_rep))
```

# eDNA data

```{r 01-COI}
edna_asvmatrix_meta <-  coi_asvmatrix %>%
  mutate(sample = str_remove_all(sample, '_quadra3')) %>%
  mutate(totalread = rowSums(select(., contains("ASV")))) %>%
  # empty negative control?
  filter(!(totalread == 0)) %>%
  rename(sample_name = sample) %>%
  # pcr rep in its own column
  separate(sample_name, 
           c("sample_name", 
             "pcr_rep" 
             ),
           sep="_") %>%
  pivot_longer(cols = !c(sample_name, pcr_rep),
             names_to = 'asv',
             values_to = 'reads_raw') %>%
  filter(reads_raw > 0) %>%
  left_join(meta_run3) %>%
  mutate(run = 'Quadra Rerun3') 
```

## 03-B Dataset singletons

One method to remove putative self positives; remove all dataset-wide singletons
skipped due to pipeiline updates

```{r 03b-singletons-coi}
# whole-dataset singleton removal
# split
# 
# edna_runrep <- edna_runrep_meta %>%
#   select(sample_id, starts_with('ASV'))
# 
# # presence-absence, removal of columns that sum to 0 or 1
# edna_runrep_pa <- edna_runrep %>%
#   # must be all numbers!
#   select(!(sample_id)) %>%
#   # vegan's decostand to convert
#   decostand(method = 'pa') %>%
#   # removing control-only asvs 
#   select(where(~ sum(.) != 0)) %>%
#   #removing one observation asvs. 
#   ## not sure about this, went from 10k to 5.4k
#   select(where(~ sum(.) != 1)) %>%
#   rownames_to_column(var = 'sample_id')
# 
# # create vector of retained asvs
# edna_runrep_retainedasv <- edna_runrep_pa %>%
#   select(!(sample_id)) %>%
#   colnames() %>%
#   enframe(name = NULL, value = "asv_r")
# 
# # reduce asv matrix + matrix/metadata
# edna_runrep_2 <- edna_runrep %>%
#   select(c(sample_id, edna_runrep_retainedasv$asv_r))
# 
# edna_runrep_meta2 <- edna_runrep_meta %>%
#   select(!(starts_with("ASV"))) %>%
#   full_join(edna_runrep_2, by = 'sample_id')

# saving again if needed
#write_csv(datacoi_asvmatrix_nc2, here::here(processededna, "coi_asvmatrix_03b.csv"))

```

## 02 PCR dissimilarity

If applicable. Not all data will have pcr reps, if it doesn't skip. From Anya 2020 script

Here we are removing samples with highly dissimilar PCR replicates

```{r 02-pcr-dissimilarity}
library(usedist)

#only run rep can do this
dat <- coi_asvmatrix %>%
  mutate(sample = str_remove_all(sample, '_quadra3')) %>%
  mutate(totalread = rowSums(select(., contains("ASV")))) %>%
  # empty negative control?
  filter(!(totalread == 0)) %>%
  rename(sample_name = sample) %>%
  # pcr rep in its own column
  separate(sample_name, 
           c("sample_name", 
             "pcr_rep" 
             ),
           sep="_") %>%
  mutate(sample_id = paste0(sample_name, '_', pcr_rep))
  

#extract usefull bits: t1 <- sample names; t2 <- ASV x site
t1 <- dat %>%
  select(sample_id, sample_name, starts_with('ASV'))

#calculate bray curtis dissimilarity
t2 <- vegdist(t1[,3:ncol(t1)], method = "bray", binary = T)

# calculate distance from group centroid using Multivariate homogeneity of groups dispersions (variances)
t3 <- betadisper(t2, group = dat$sample_name, type = "centroid")

#plot with 95% CI elipses
 plot(t3, hull = F, ellipse = T, conf = 0.95, label = F, pch = 5, main = "95% CI elipse around sample replicates")

#extract distances
t4 <- cbind(dat$sample_id, dat$sample_name, as.data.frame(t3$distances))
colnames(t4) <- c("sample_id", "sample_name", "distances")

#calculate 95% CIs
t4 <- t4 %>%
  group_by(sample_name) %>%
  mutate(meandis=mean(distances), # mutate when grouped applies the summary to each row
         SDdis=sd(distances),
         CIdis = 1.96*SDdis/sqrt(3), 
         upperCIdis=meandis+CIdis) #calculate upper end of 95% CI

#check that PCR reps are within CIs
t4$pcrvariable <- ifelse(t4$upperCIdis<t4$distances, "problem", "ok")

#merge pcr info with original
edna_runrep_meta<- edna_asvmatrix_meta %>%
    left_join(t4) %>%
  relocate(c("pcrvariable"), .before = 'run_id')

```

## 03-A -- Occupancy model QC-type

We are skipping this for now, since Erikson/Mize doesn't use it. May want to re-add for other analyses?

## 04 -- remove field control reads

Not applicable here

## 05 -- Taxonomy


```{r taxonomy}

taxonomy_coi <- read_csv(here(processededna, '20240220_calvert2018_coi_assigned_asvs.csv'))

taxonomy_motus <- read_csv(here('processeddata', '2018calvert', 'qcstages',
                                '20240220_calvert2018_coi_swarm_statfile-taxonomy.csv')) %>%
  select(seed_asv, motu, found_taxa, common_name, n_in_cluster) %>%
  rename(found_taxa_motu = found_taxa,
         common_name_motu = common_name)

motu2asv <- read_csv(here('processeddata', '2018calvert', 'qcstages',
                                '20240220_calvert2018_coi_motu2asv.csv')) %>%
  select(!c(seed_reads, asv_reads)) 


asvmatrix_meta_taxonomy <- edna_runrep_meta %>%
  left_join(taxonomy_coi) %>%
  relocate(found_taxa, common_name, rank, .before = sample_id) %>%
  #add motu info
  left_join(motu2asv) %>%
  relocate(motu, .before=reads_raw) %>%
  left_join(taxonomy_motus) %>%
  relocate(found_taxa_motu, common_name_motu, .before = sample_id)




```


## Final matrices

Where we fix it up after all the control stuff is taken care of but we skip this because it's going to be taken care of in the script where we merge all runs rather htan doing it 4 times

```{r final}



#save whole list
write_csv(asvmatrix_meta_taxonomy, here(processededna,
                                          paste0(datefield, runfield,
                                                 'asvmatrix_taxonomy.csv')))


```

