---
title: "ch1_coi-01b_swarm"
output: html_document
date: "2023-08-22"
---

Here we're going to be testing swarm, according to brandt et al 2021 (https://onlinelibrary.wiley.com/doi/10.1111/1755-0998.13398), they used swarm on asv tables from dada2.
That being the case, we can start with the data from Evan for chapter 1.


Swarm was installed then made available to users as explained on the github: https://github.com/torognes/swarm
the binary from the installation copied to /usr/local/bin, which is on `$PATH`

Here we will use a similar strategy to Evan's dada2 scripts and use system2 to call swarm

```{r}
library(tidyverse)
library(here)

# filepaths for use with here, basic files
coi_in <- here("rawdata", "2018calvert", "fastq",
               'coi_kate', 'combined_runs')
# Destination folders
coi_out <- here('processeddata', '2018calvert', 'qcstages')

# datefield might be different
datefield_in <- '20240219_'
datefield_out <- '20240220_'
runfield <- 'calvert2018_coi_'
```


right now i don't have sequence tables for this data, only asv tables.
I guess step 1 is to turn the sequence table INTO an asv table?
I think this is a lot easier in python I probably shouldn't waste my time here at the moment.


Note I then need to take the sequence tables and turn them into a fasta formatted for swarm

Now that they're formatted for swarm we can run swarm??

For using system2, we need to generate filepaths for the arguments in the call.
```{r filepaths}
# filepath to fasta formatted for swarm
swarmfasta <- file.path(here(coi_in, 
                             paste0(datefield_in, runfield,
                                    "sequence-table_swarm-format.fasta")))

#generate filenames for output
swarmout <- file.path(here(coi_in, 
                           paste0(datefield_out, runfield, 
                                  'swarm_outputfile.txt')))
swarmstats <- file.path(here(coi_in, 
                             paste0(datefield_out, runfield, 
                                    'swarm_statsfile.tsv')))
swarmlog <- file.path(here(coi_in, 
                           paste0(datefield_out, runfield, 
                                  'swarm_log.txt')))
swarmreps <- file.path(here(coi_in, 
                            paste0(datefield_out, runfield, 
                                   'swarm_clusterreps.fasta')))
```


```{r swarm call}
system2("swarm", args = "-v")

system2("swarm", args = c(swarmfasta,
                          "-w", swarmreps,
                          "-o", swarmout,
                          "-s", swarmstats,
                          "-l", swarmlog,
                          "-f"))

```

# input swarm output to generate motus

should probably add 'curate with lulu' step before assigning motus; temp num to start?

```{r input swarm}

swarm_stat <- read_delim(here(coi_in, 
                             paste0(datefield_out, runfield, 
                                    'swarm_statsfile.tsv')),
                         # give columns names
                        col_names = c('n_in_cluster', 'motu_reads', 
                                      'seed_asv', 'seed_reads', 'n_1readamplicon',
                                      'max_iterations_to_limit', 'cumulative_steps')) 

swarm_stat <- swarm_stat %>%
 mutate(motu = paste0('motu', seq.int(from = 1, to = nrow(swarm_stat)))) 


swarm_out <- read_delim(here(coi_in, 
                             paste0(datefield_out, runfield, 'swarm_outputfile.txt')),
                        col_names = FALSE, delim = " ")


# format
swarm_out <- swarm_out %>%
  # identify seeds for taxonomy assignment ~TEMPORARY~
  rename(seed_asv = X1) %>%
  # make sure the seed itself is part of the key
  mutate(X3 = seed_asv) %>%
  # add 'motu' numbers
  mutate(motu = paste0('motu', seq.int(from = 1, to = nrow(swarm_out)))) %>%
  # make long, no more NAs
  pivot_longer(cols = !c(motu, seed_asv)) %>%
  select(!name) %>%
  # if its a cluster of 1, keep the value in
  mutate(value = ifelse(is.na(value), seed_asv, value)) %>%
  # separate out members of cluster
  separate_wider_delim(value, delim = " ", names_sep = "_", too_few = 'align_start') %>%
  # pair seed and each member
  pivot_longer(cols = !c(motu, seed_asv)) %>%
  # clean pivot, NA's useless here
  select(!name) %>%
  filter(!(is.na(value))) %>%
  # asv + reads each get their own column
  separate_wider_delim(value, delim = "_", names = c("asv", "asv_reads")) %>%
  # seed reads are different
  separate_wider_delim(seed_asv, delim = "_", names = c("seed_asv", "seed_reads"))
  
  
write_csv(swarm_out, here(coi_out, paste0(datefield_out, runfield, 'motu2asv.csv')))



# to do; replace asv names in fasta
```



# update eDNA

After BLAST, run top 10 for the MOTU versions
this is still the mifish version. the COI version might look quite different

top 10 taxonomy for MOTU

In this version we are going to merge all ASV annotations and motu annotations so here we just want what motu is what taxon so that we can add the column for filtering.

```{r top hits}
# load-in keys
#assigned_asvs <- read_csv(here(coi_in, paste0(datefield_out, runfield, 'assigned_asvs.csv'))) %>%
  # so that it only joins the seed asvs rather than every asv
#  rename(seed_asv = asv)

# we reblasted but didn't rerun swarm
assigned_asvs <- read_csv(here(coi_in, paste0('20240220_', runfield, 'assigned_asvs.csv'))) %>%
  # so that it only joins the seed asvs rather than every asv
  rename(seed_asv = asv)

motu_taxonomy <- swarm_stat %>%
  left_join(assigned_asvs)

#write_csv(motu_taxonomy, here(coi_out, paste0(datefield_out, runfield, 'swarm_statfile-taxonomy.csv')))
# reblasted but didn't rerun swarm
write_csv(motu_taxonomy, here(coi_out, paste0('20240220_', runfield, 'swarm_statfile-taxonomy.csv')))
```