---
title: "ch1eDNAQC-12s"
author: "Kate Sheridan"
date: "11/27/2021"
output: html_document
---

Version of Sunday Lab eDNA QC pipeline modified from Trawl Project 2020, scripts by Anya Mueller, Jenn Sunday, Ben Millard-Martin

Main difference with this version is that it is eDNA-only, with no cleaning of trawl related files, and Hakai samples. Updated August 2022 with key replacement

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
#for pcr dissimilarity
library(vegan)
library(usedist)

library(here)
# filepaths for use with here, basic files
mifish <- here("rawdata", "2018calvert", "fastq", "12S_Calvert2018_Re-run20210903", "processed")

# Destination folders
processededna <- here('processeddata', '2018calvert', 'qcstages')

datefield <- '20240323_'
runfield <- 'calvert2018_12s_'
```

## 01 - Initial Cleaning

Initial cleaning includes: cleaning names of columns, creating a unique site ID, cleaning eDNA data to be more usable. Modified from Anya's 2020 script

Also adding in coordinates and habitat information.

```{r 01-initialcleaning-loadin, message=FALSE}
mifish_asvmatrix <- read.table(here::here(mifish, 
                                           "20240315_12s_2018calvert_rerun20210903_sequence-table_12S-merged-w-ASV-names.txt"), 
                               header = TRUE)

taxonomy_12s <- read_csv(here(processededna,
                              '20240323_calvert2018_12s_taxonomy.csv'))

edna_metadata <- read_csv(here('processeddata', '2018calvert', "2018calvert_eDNA_metadata.csv"))[-1]

```

# eDNA data

First cleaning the data and generating matrix

```{r 01-join metadata}
mifish_asvmatrix_meta <-  mifish_asvmatrix %>%
  filter(!(row_names == 'Undetermined')) %>%
  mutate(totalread = rowSums(select(., contains("ASV")))) %>%
  filter(!(totalread == 0)) %>%
  # first change the columns from ASV to motu
  pivot_longer(cols = !row_names,
               names_to = 'asv',
               values_to = 'raw_reads') %>% 
  separate(row_names, 
           c("sample_name", 
             "pcr_rep" 
             ), #separate row_name label into vectors of information, 
                   #which info you need will vary based on how it was recorded and processed
           sep="-") %>%
  #make unique identifier for each row
  mutate(sample_id=paste(sample_name, pcr_rep, sep="-")) %>% 
  mutate(pcr_rep = tolower(pcr_rep),
         sample_id = tolower(sample_id),
         sample_name = tolower(sample_name)) %>%
  filter(!(asv == 'totalread')) %>%
  left_join(edna_metadata) %>%
  # this removes samples that didn't make it through run processing
  filter(!(is.na(asv))) %>%
  distinct()

```

## 02 Controls

Can't match controls with anything so currently just removing... hopefully updated eventually.
```{r 02-12s}
unique(mifish_asvmatrix_meta$sample_name)

mifish_asvmatrix_meta %>% filter(sample_name %in% c('neg1', 'neg2', 'neg3') & raw_reads >= 1)

controls <- mifish_asvmatrix_meta %>%
  filter(sample_name %in% c('neg1', 'neg2', 'neg3') & raw_reads >= 1) %>%
  select(sample_name, pcr_rep, asv, raw_reads, sample_id) %>%
  mutate(pcr_asv = paste0(pcr_rep, '_', asv)) %>%
  group_by(pcr_asv) %>%
  mutate(ctrl_mean = round(mean(raw_reads))) %>%
  ungroup() %>%
  select(pcr_rep, asv, ctrl_mean) %>%
  distinct()

# read subtract version
mifish_asvmatrix_meta_nc <- mifish_asvmatrix_meta %>%
  filter(!raw_reads == 0) %>%
  left_join(controls) %>%
  # make any NAs 0
  mutate(ctrl_mean = ifelse(is.na(ctrl_mean), 0, ctrl_mean)) %>%
  # subtract reads by pcr rep
  mutate(reads_adjust = raw_reads - ctrl_mean) %>%
  # remove named controls, edit to be what is in dataframe
  filter(!grepl("neg", sample_name)) %>%
  # make negatives 0s
  mutate(reads_adjust = ifelse(reads_adjust < 0, 0, reads_adjust))


# # remove control version
# mifish_asvmatrix_meta_nc <- mifish_asvmatrix_meta %>%
#   # remove named controls, edit to be what is in dataframe
#   filter(!grepl("neg", sample_name)) 
```

## 03-B Dataset singletons

One method to remove putative self positives; remove all dataset-wide singletons

Due to revisions of the pipeline, we'll skip this step for now.

```{r 03b-singletons-12s}
# # whole-dataset singleton removal
# 
# # presence-absence, removal of columns that sum to 0 or 1
# mifish_asvmatrix_nc_pa <- mifish_asvmatrix_nc %>%
#   # must be all numbers!
#   column_to_rownames(var = 'sample_id') %>%
#   # vegan's decostand to convert
#   decostand(method = 'pa') %>%
#   # removing control-only asvs 
#   select(where(~ sum(.) != 0)) %>%
#   #removing one observation asvs. 
#   ## not sure about this, went from 1726 to 373? match to taxon table, might be ok cause mifish?
#   select(where(~ sum(.) != 1)) %>%
#   rownames_to_column(var = 'sample_id')
# 
# # create vector of retained asvs
# mifish_retainedasv <- mifish_asvmatrix_nc_pa %>%
#   select(!(sample_id)) %>%
#   colnames() %>%
#   enframe(name = NULL, value = "asv_r")
# 
# # reduce asv matrix + matrix/metadata
# mifish_asvmatrix_nc2 <- mifish_asvmatrix_nc %>%
#   select(c(sample_id, mifish_retainedasv$asv_r))
# 
# mifish_asvmatrix_metadata_nc2 <- mifish_asvmatrix_metadata_nc %>%
#   select(!(starts_with("ASV"))) %>%
#   full_join(mifish_asvmatrix_nc2, by = 'sample_id')
# 
# # saving again if needed
# write_csv(mifish_asvmatrix_nc2, here::here(processededna, "12S_asvmatrix_03b.csv"))

```

## 02 PCR dissimilarity

If applicable. Not all data will have pcr reps, if it doesn't skip. From Anya 2020 script

Here we are removing samples with highly dissimilar PCR replicates

used reads_adjust

```{r 02-pcr-dissimilarity}
dat <- mifish_asvmatrix_meta_nc %>%
  select(sample_id, sample_name, asv, reads_adjust) %>%
  pivot_wider(names_from = asv, 
              values_from = reads_adjust) %>%
  # NA to 0
  mutate_all(~replace(., is.na(.), 0))

#extract useful bits: t1 <- sample names; t2 <- ASV x site
t1 <- dat %>%
  select(sample_id, sample_name, starts_with('asv'))

#calculate bray curtis dissimilarity
t2 <- vegdist(t1[,3:ncol(t1)], method = "bray", binary = T)

# calculate distance from group centroid using Multivariate homogeneity of groups dispersions (variances)
t3 <- betadisper(t2, group = dat$sample_name, type = "centroid")

#plot with 95% CI elipses
 plot(t3, hull = F, ellipse = T, conf = 0.95, label = F, pch = 5, main = "95% CI elipse around sample replicates")

#extract distances
t4 <- cbind(dat$sample_id, dat$sample_name, as.data.frame(t3$distances))
colnames(t4) <- c("sample_id", "sample_name", "distances")

#calculate 95% CIs
t4 <- t4 %>%
  group_by(sample_name) %>%
  mutate(meandis=mean(distances), # mutate when grouped applies the summary to each row
         SDdis=sd(distances),
         CIdis = 1.96*SDdis/sqrt(3), 
         upperCIdis=meandis+CIdis) #calculate upper end of 95% CI

#check that PCR reps are within CIs
t4$diff <- ifelse(t4$upperCIdis<t4$distances, "problem", "ok")

#list for later filtering
problemlist <- t4 %>%
  subset(diff == "problem") %>%
  distinct(sample_name, sample_id)

#now we want to remove the pcr reps identified as problems (pcr reps?)
# I am just going to remove the replicate itself
mifish_asvmatrix_nc_95ci <- mifish_asvmatrix_meta_nc %>%
  filter(!(sample_id %in% problemlist$sample_id))

#not going to keep the replicate information past here
```

## 03-A -- Occupancy model QC-type

We are skipping this for now, since Erikson/Mize doesn't use it. May want to re-add for other analyses?


## 05 -- Taxonomy

```{r 05-taxonomy-12s}
# merge taxonomy file with asv table
mifish_asvmatrix_taxonomy <- mifish_asvmatrix_nc_95ci %>%
  filter(reads_adjust > 0) %>%
  # totalread is useless
  filter(!(asv == 'totalread')) %>%
  left_join(taxonomy_12s) %>%
  select(!(c(volume_filtered, lab_notes, hakai_project,
             hakai_link_data, sample_type, veg_sub, vegcom_sub,
             ))) %>%
  relocate(asv, raw_reads, reads_adjust, found_taxa_asv, common_name_asv, 
           rank_asv, location, macroveg_group, macroveg_community,
           .before = pcr_rep)

# right now the full asv list isn't being added back in after blast, need to assess.

# sort out unidentified asvs
mifish_asvmatrix_unknown <- mifish_asvmatrix_taxonomy %>%
  filter(found_taxa_asv == 'No BLAST hit')

# write out problems
write.csv(mifish_asvmatrix_unknown, here(processededna, 
                                          paste0(datefield, runfield,
                                                 'asvmatrix_taxonomy-problems.csv')))



```



## Final matrices

```{r mifish-finalmatrix}
# species list
asv_splist <- mifish_asvmatrix_taxonomy %>%
  filter(class %in% c('Teleostei', 'Holocephali')) %>%
  select(found_taxa_asv, common_name_asv, rank_asv) %>%
  distinct()

#save whole list
write_csv(asv_splist, here('processeddata', 'species_lists', 
                           paste0(datefield, runfield,
                                          'asv_splist.csv')))


mifish_asvmatrix_final <- mifish_asvmatrix_taxonomy %>%
  group_by(sample_name, asv) %>%
  mutate(raw_reads_all = sum(raw_reads)) %>%
  mutate(reads_adjust_all = sum(reads_adjust)) %>%
  ungroup() %>%
  relocate(raw_reads_all, reads_adjust_all, .before = found_taxa_asv) %>%
  rename(raw_reads_pcr = raw_reads,
         reads_adjust_pcr = reads_adjust) %>%
  relocate(raw_reads_pcr, reads_adjust_pcr, .before = pcr_rep) %>%
  distinct() %>%
  mutate(asv_pa = 1) %>%
  relocate(asv_pa, .before = found_taxa_asv) %>%
  filter(!(is.na(class)))


#save whole list
write_csv(mifish_asvmatrix_final, here(processededna,
                                          paste0(datefield, runfield,
                                                 'asvmatrix_taxonomy.csv')))

mifish_asvmatrix_final_fish <- mifish_asvmatrix_final %>%
  filter(class %in% c('Teleostei', 'Holocephali'))

#save whole list
write_csv(mifish_asvmatrix_final, here(processededna,
                                          paste0(datefield, runfield,
                                                 'asvmatrix_taxonomy-fish.csv')))
```





