---
title: "ch1eDNAQC-12s"
author: "Kate Sheridan"
date: "11/27/2021"
output: html_document
---

Version of Sunday Lab eDNA QC pipeline modified from Trawl Project 2020, scripts by Anya Mueller, Jenn Sunday, Ben Millard-Martin

Main difference with this version is that it is eDNA-only, with no cleaning of trawl related files, and Hakai samples. Updated August 2022 with key replacement

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(vegan)
# for PCR replicates
library(usedist) 
library(here)

# filepaths for use with here, basic files
#mifish <- here("rawdata", "2018calvert", "12S")
mifish <- here("rawdata", "2018calvert", "fastq", "12S_Calvert2018_Re-run20210903", "processed")

# Destination folders
processededna <- here('processeddata', '2018calvert', 'qcstages')

datefield <- '20240323_'
runfield <- 'calvert2018_12s_'
```

## 00 - Load In

Initial cleaning includes: cleaning eDNA data to be more usable. Modified from Anya's 2020 script

here you want to read in any markers and metadata
These will be sequence + ASV merged tables, saved as a .txt file, with metadata as .csv
Filenames if directly from Hakai will be a mix of . and _ separators, be careful entering
NOTE: watch COI vs CO1, some files my use a letter I vs a number 1. 

```{r 01-initialcleaning-loadin, message=FALSE}
mifish_asvmatrix <- read.table(here::here(mifish, 
                                           "20240315_12s_2018calvert_rerun20210903_sequence-table_12S-merged-w-ASV-names.txt"), 
                               header = TRUE)

motu2asv <- read_csv(here(mifish, '20240315_calvert2018_12s_motu2asv.csv')) %>%
  select(!asv_reads)

taxonomy_12s <- read_csv(here(processededna,
                              '20240323_calvert2018_12s_taxonomy.csv'))

edna_metadata <- read_csv(here('processeddata', '2018calvert', "2018calvert_eDNA_metadata.csv"))[-1]

```


# 01 - ASV to MOTU + metadata

First cleaning the data and generating matrix

To do read subtraction at the ASV level, we will first make our matrix an ASV matrix, then rotate it to MOTU after subtracting controls.


We will match our ASV table to MOTUs to collapse and add metadata.

```{r 02 asv2motu}

mifish_asvmatrix_meta <-  mifish_asvmatrix %>%
  filter(!(row_names == 'Undetermined')) %>%
  mutate(totalread = rowSums(select(., contains("ASV")))) %>%
  filter(!(totalread == 0)) %>%
  # first change the columns from ASV to motu
  pivot_longer(cols = !row_names,
               names_to = 'asv',
               values_to = 'raw_reads') %>% 
  separate(row_names, 
           c("sample_name", 
             "pcr_rep" 
             ), #separate row_name label into vectors of information, 
                   #which info you need will vary based on how it was recorded and processed
           sep="-") %>%
  #make unique identifier for each row
  mutate(sample_id=paste(sample_name, pcr_rep, sep="-")) %>% 
  mutate(pcr_rep = tolower(pcr_rep),
         sample_id = tolower(sample_id),
         sample_name = tolower(sample_name)) %>%
  filter(!(asv == 'totalread')) %>%
  left_join(edna_metadata) %>%
  # this removes samples that didn't make it through run processing
  filter(!(is.na(asv))) %>%
  distinct() %>%
  left_join(motu2asv) %>%
  filter(!(raw_reads == 0))



mifish_motumatrix <- mifish_asvmatrix %>%
  # first change the columns from ASV to motu
  pivot_longer(cols = !row_names,
               names_to = 'asv',
               values_to = 'reads') %>%
  # use the key for which asvs collapse into an motu
  left_join(motu2asv) %>%
  select(row_names, motu, reads) %>%
  # new table; motu x sample-pcrrep
  pivot_wider(names_from = motu,
              values_from = reads, 
              values_fn = sum) %>%
  separate(row_names, 
           c("sample_name", 
             "pcr_rep" 
             ), #separate row_name label into vectors of information, 
                   #which info you need will vary based on how it was recorded and processed
           sep="-") %>%
  #make unique identifier for each row
  mutate(sample_id=paste(sample_name, pcr_rep, sep="-")) %>% 
  #bring to front
  relocate(sample_id, .before = 'motu1') %>% 
  # get rid of the NAs added from columns with no value
  mutate(sample_id=str_replace(sample_id, "-NA", "")) %>% 
  mutate(pcr_rep = tolower(pcr_rep),
         sample_id = tolower(sample_id),
         sample_name = tolower(sample_name)) %>%
  # this 'sample' doesn't seem useful?
  filter(!(sample_name == 'undetermined')) %>%
  #mutate(pcr_rep = ifelse(is.na(pcr_rep) == TRUE, 'pooled', pcr_rep)) %>% 
  mutate(totalread = rowSums(select(., contains("motu")))) %>%
  filter(!(totalread == 0)) %>%
  pivot_longer(cols = !c(sample_name, sample_id, pcr_rep),
               names_to = 'motu',
               values_to = 'raw_reads') %>%
  # join metadata
  full_join(edna_metadata) %>%
  # this removes samples that didn't make it through run processing
  filter(!(is.na(motu))) %>%
  distinct()

```

# 02 Controls

Normally we'd subtract reads or something here but we have insufficient metadata for what the controls represent, and the PCR blanks didn't make it to FASTQ files?

```{r 02-12s}
controls <- mifish_asvmatrix_meta %>%
  filter(sample_name %in% c('neg1', 'neg2', 'neg3') & raw_reads >= 1) %>%
  select(sample_name, pcr_rep, asv, raw_reads, sample_id) %>%
  mutate(pcr_asv = paste0(pcr_rep, '_', asv)) %>%
  group_by(pcr_asv) %>%
  mutate(ctrl_mean = round(mean(raw_reads))) %>%
  ungroup() %>%
  select(pcr_rep, asv, ctrl_mean) %>%
  distinct()

# read subtract version
mifish_motumatrix_nc <- mifish_asvmatrix_meta %>%
  left_join(controls) %>%
  # make any NAs 0
  mutate(ctrl_mean = ifelse(is.na(ctrl_mean), 0, ctrl_mean)) %>%
  # subtract reads by pcr rep
  mutate(reads_adjust = raw_reads - ctrl_mean) %>%
  # remove named controls, edit to be what is in dataframe
  filter(!grepl("neg", sample_name)) %>%
  # make negatives 0s
  mutate(reads_adjust = ifelse(reads_adjust < 0, 0, reads_adjust)) %>%
  # collapse and switch to motus
  group_by(motu, sample_name) %>%
  mutate(reads_adjust_sample = sum(reads_adjust)) %>%
  mutate(reads_raw_sample = sum(raw_reads)) %>%
  ungroup() %>%
  #select(!c(asv_reads))%>%
  relocate(motu, reads_adjust_sample, reads_raw_sample, .before = asv)



# 
# # remove control version
# mifish_motumatrix_nc <- mifish_motumatrix %>%
#   # remove named controls, edit to be what is in dataframe
#   filter(!grepl("neg", sample_name)) 

```

Old 03 Dataset singletons

One method to remove putative self positives; remove all dataset-wide singletons

Due to revisions of the pipeline, we'll skip this step for now.
We had already removed low-read singleton ASVs during dada2 pipeline.

```{r 03b-singletons-12s}
# # whole-dataset singleton removal
# 
# # presence-absence, removal of columns that sum to 0 or 1
# mifish_asvmatrix_nc_pa <- mifish_asvmatrix_nc %>%
#   # must be all numbers!
#   column_to_rownames(var = 'sample_id') %>%
#   # vegan's decostand to convert
#   decostand(method = 'pa') %>%
#   # removing control-only asvs 
#   select(where(~ sum(.) != 0)) %>%
#   #removing one observation asvs. 
#   ## not sure about this, went from 1726 to 373? match to taxon table, might be ok cause mifish?
#   select(where(~ sum(.) != 1)) %>%
#   rownames_to_column(var = 'sample_id')
# 
# # create vector of retained asvs
# mifish_retainedasv <- mifish_asvmatrix_nc_pa %>%
#   select(!(sample_id)) %>%
#   colnames() %>%
#   enframe(name = NULL, value = "asv_r")
# 
# # reduce asv matrix + matrix/metadata
# mifish_asvmatrix_nc2 <- mifish_asvmatrix_nc %>%
#   select(c(sample_id, mifish_retainedasv$asv_r))
# 
# mifish_asvmatrix_metadata_nc2 <- mifish_asvmatrix_metadata_nc %>%
#   select(!(starts_with("ASV"))) %>%
#   full_join(mifish_asvmatrix_nc2, by = 'sample_id')
# 
# # saving again if needed
# write_csv(mifish_asvmatrix_nc2, here::here(processededna, "12S_asvmatrix_03b.csv"))

```

## 03 PCR dissimilarity

If applicable. Not all data will have pcr reps, if it doesn't skip. From Anya 2020 script

Here we are removing samples with highly dissimilar PCR replicates

```{r 03-pcr-dissimilarity}
# Code is reused a lot here and we don't need the dat object again
# So I keep the renaming from orginal scripts.
dat <- mifish_motumatrix_nc %>%
  select(sample_id, sample_name, motu, reads_adjust_sample) %>%
  distinct() %>%
  filter(!(motu == 'totalread')) %>%
  pivot_wider(names_from = motu, 
              values_from = reads_adjust_sample) %>%
  # NA to 0
  mutate_all(~replace(., is.na(.), 0))
   

#extract useful bits: t1 <- sample names; t2 <- ASV x site
t1 <- dat %>%
  select(sample_id, sample_name, starts_with('motu')) 

#calculate bray curtis dissimilarity
t2 <- vegdist(t1[,3:ncol(t1)], method = "bray", binary = T)

# calculate distance from group centroid using Multivariate homogeneity of groups dispersions (variances)
t3 <- betadisper(t2, group = t1$sample_name, type = "centroid")

#plot with 95% CI elipses
 plot(t3, hull = F, ellipse = T, conf = 0.95, label = F, pch = 5, main = "95% CI elipse around sample replicates")

#extract distances
t4 <- cbind(dat$sample_id, dat$sample_name, as.data.frame(t3$distances)) 
colnames(t4) <- c("sample_id", "sample_name", "distances")

#calculate 95% CIs
t4 <- t4 %>%
  group_by(sample_name) %>%
  mutate(meandis = mean(distances), # mutate when grouped applies the summary to each row
         SDdis = sd(distances),
         CIdis = 1.96*SDdis/sqrt(3), 
         upperCIdis=meandis+CIdis) %>% #calculate upper end of 95% CI 
  ungroup()

#check that PCR reps are within CIs
t4$diff <- ifelse(t4$upperCIdis < t4$distances, "problem", "ok")

#list for later filtering
problemlist <- t4 %>%
  subset(diff == "problem") %>%
  distinct(sample_name, sample_id)

#now we want to remove the samples identified as problems (whole samples or just pcr reps?)
# I am just going to remove the replicate itself
mifish_motumatrix_nc_95ci <- mifish_motumatrix_nc %>%
  filter(!(sample_id %in% problemlist$sample_id))

#not going to keep the replicate information past here
```

## 03-B -- Occupancy model QC-type

We are skipping this for now, since Erikson/Mize doesn't use it. May want to re-add for other analyses?

I may want to replace the pcr step with this.


## 04 -- Taxonomy

Since we QCd the taxonomy assignment at the BLAST step, here we are simply merging the file.

```{r 04-taxonomy-12s}
# merge taxonomy file with motu table
mifish_motumatrix_taxonomy <- mifish_motumatrix_nc_95ci %>%
  filter(reads_adjust_sample> 0) %>%
  # totalread is useless
  filter(!(motu == 'totalread')) %>%
  left_join(taxonomy_12s) %>%
  select(!(c(#seed_asv_reads, seed_reads, motu_reads,
             volume_filtered, lab_notes, hakai_project,
             hakai_link_data, sample_type, veg_sub, vegcom_sub
             ))) %>%
  relocate(motu, reads_adjust_sample, found_taxa_motu, common_name_motu, 
           rank_motu, location, macroveg_group, macroveg_community,
           .before = pcr_rep)

# sort out unidentified motus
mifish_motumatrix_unknown <- mifish_motumatrix_taxonomy %>%
  filter(common_name_motu == 'No BLAST hit')

# write out problems
write.csv(mifish_motumatrix_unknown, here(processededna, 
                                          paste0(datefield, runfield,
                                                 'motumatrix_taxonomy-problems.csv')))


```


## Final matrices

remove unidentified motus, add p/a column

```{r mifish-finalmatrix}
# species list
motu_splist <- mifish_motumatrix_taxonomy %>%
  filter(class %in% c('Teleostei', 'Holocephali')) %>%
  select(found_taxa_motu, common_name_motu, rank_motu) %>%
  distinct()

#save whole list
write_csv(motu_splist, here('processeddata', 'species_lists', 
                            paste0(datefield, runfield,
                                          'motu_splist.csv')))

mifish_motumatrix_final <- mifish_motumatrix_taxonomy %>%
  filter(!(is.na(class))) %>%
  mutate(motu_pa = 1) %>%
  relocate(motu_pa, .before = found_taxa_motu)


#save whole list
write_csv(mifish_motumatrix_final, here(processededna,
                                          paste0(datefield, runfield,
                                                 'motumatrix_taxonomy.csv')))



mifish_motumatrix_final_fish <- mifish_motumatrix_final %>%
  filter(class %in% c('Teleostei', 'Holocephali'))

#save whole list
write_csv(mifish_motumatrix_final_fish, here(processededna,
                                          paste0(datefield, runfield,
                                                 'motumatrix_taxonomy-fish.csv')))

```





