---
title: "merge-years"
output: html_document
date: "2023-05-31"
---

This script combines years of data into one file of each type

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(PNWColors)

# here filepaths
edna_2021 <- here('processeddata', 'peco', '2021_edna')
edna_2022 <- here('processeddata', 'peco', '2022_edna')
edna_2023 <- here('processeddata', 'peco', '2023_edna')
edna_2024 <- here('processeddata', 'peco', '2024_edna')

edna_out <- here('processeddata', 'peco', '2021-24_edna')

datefield <- '20251107_'
runfield <- 'peco-combined_12s_'

# Palette
## match map with 14 levels if by region
## or match to 17 levels corresponding with partners: note that San Francisco Bay has 2 partners
pal <-  pnw_palette("Bay", 17, type = "continuous")                               
#make it an object to use as the colour value in the plot

pal14 <- pnw_palette("Bay", 14, type = "continuous")

pal2023 <- pnw_palette("Bay", 15, type = "continuous")
```


asv matrix combine
Blend 2021 and 2022, standardize to the same blast run

```{r combine matrices}
meta_all <- read_csv(here(edna_out, '20251107_peco-combined_metadata.csv')) 

asvs2021 <- read_csv(here(edna_2021, '20250528_peco-2021_12s_asvmatrix_complete.csv')) %>%
  mutate(year = 2021) %>%
  select(!c(dna_extraction_date))

asvs2022 <- read_csv(here(edna_2022, '20250528_peco-2022_12s_motumatrix_complete.csv')) %>%
  mutate(year = 2022)%>%
  select(!c(dna_extraction_date))

# note; update this?
asvs2023 <- read_csv(here(edna_2023, '20250528_peco2023_12s-asvmatrix-temp.csv')) %>%
  mutate(year = 2023) %>%
  mutate(sample_collection_date = as.character(sample_collection_date),
         salinity_psu = as.character(salinity_psu),
         temperature_c = as.character(temperature_c))


asvs2024 <- read_csv(here(edna_2024, '20250528_peco2024_12s-asvmatrix-temp.csv')) %>%
    mutate(sample_collection_date = as.character(sample_collection_date),
         salinity_psu = as.character(salinity_psu),
         temperature_c = as.character(temperature_c),
         notes_about_sample = as.character(notes_about_sample))

asv_all <- asvs2021 %>%
  full_join(asvs2022) %>%
  full_join(asvs2023) %>%
  full_join(asvs2024) %>%
  janitor::remove_empty() %>%
  relocate(sample_id, motu, found_taxa, common_name, rank, year) %>%
  # don't know where this error is
  mutate(region_code = ifelse(region_code == 'CA_Nca', 'CA_NCa',region_code)) %>%
  # duplicate to remove soon
  mutate(site_code = case_when(site_code == 'SD_SD' ~ 'SD_SDL',
                               site_code == 'YB_SN' ~ 'YB_SBN',
                               site_code == 'YB_SS' ~ 'YB_SBS',
                               site_code == 'OR_SBS' ~ 'YB_SBS',
                               site_code == 'WB_NA' ~ 'WB_PA',
                               site_code == 'HB_SB' ~ 'HB_SR',
                               site_code == 'SD_Coronado' ~ 'SD_CI',
                               site_code == "YB_FP" ~ "CB_FP",
                               site_code == "YB_BV" ~ "CB_BV",
                               site_code %in% c('NV_KI', 'NV_KH') ~ 'NV_KB',
                               site_code == 'NC_PS' & region_code == 'WA_PSb' ~ 'PS_NC',
                               TRUE ~ site_code))%>%
  mutate(site_code = ifelse(site_code == 'HB_SB', 'HB_SR', site_code)) %>%
    mutate(region_name = case_when(site_code %in% c("CB_FP", "CB_BV") ~ "OR_CB",
                                 site_code %in% c("NC_PM", "NC_PS",
                                                  "NC_RB", "NC_WP") ~ "CA_SF",
                                 TRUE ~ region_name)) %>%
  mutate(region_code = ifelse(region_name == 'OR_CB', 'OR_CB',region_code))

new_range <- asv_all %>%
  group_by(found_taxa) %>%
  summarise(taxa_max = max(decimal_latitude, na.rm = TRUE), taxa_mean = mean(decimal_latitude, na.rm = TRUE),
            taxa_median = median(decimal_latitude, na.rm = TRUE), taxa_min = min(decimal_latitude, na.rm = TRUE))

fishonly_range <- asv_all %>%
  left_join(new_range) %>%
  filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii'))

write_csv(fishonly_range, here(edna_out, paste0(datefield, runfield, 'asvmatrix_fish_all.csv')))
```


# controls
```{r combine controls}
ctrl2021 <- read_csv(here(edna_2021, '20250528_peco-2021_12s_controls.csv')) %>%
  mutate(year = 2021) 

ctrl2022 <- read_csv(here(edna_2022, '20250505_peco2022_12s_controls.csv')) %>%
  mutate(year = 2022) 

ctrl2023 <- read_csv(here(edna_2023, '20250528_peco2023_12s_controls.csv')) %>%
  mutate(year = 2023) %>%
  mutate(sample_collection_date = as.character(sample_collection_date),
         salinity_psu = as.character(salinity_psu),
         temperature_c = as.character(temperature_c))

ctrl2024 <- read_csv(here(edna_2024, '20250528_peco2024_12s_controls.csv')) %>%
  mutate(sample_collection_date = as.character(sample_collection_date),
         salinity_psu = as.character(salinity_psu),
         temperature_c = as.character(temperature_c))



ctrl_all <- ctrl2021 %>%
  full_join(ctrl2022) %>%
  full_join(ctrl2023) %>%
  full_join(ctrl2024)

write_csv(ctrl_all, here(edna_out, paste0(datefield, 'peco-combined_12s_controls.csv')))
```
# uniques and splist
```{r}
#uniques for quick mapping
#all
mifish_coords <- fishonly_range %>%
  distinct(found_taxa, common_name, rank, 
           decimal_latitude, decimal_longitude,
           taxa_min, taxa_max, taxa_median, taxa_mean) %>%
# there are NAs at the moment, delete later
  filter(!(is.na(decimal_latitude))) %>%
  filter(!(is.na(found_taxa))) %>%
  arrange(found_taxa)
#save if needed
write.csv(mifish_coords, here(edna_out, paste0(datefield, runfield, 'fish-coords.csv')))

# uniques for pcoa etc
mifish_coords2 <- fishonly_range %>%
  select(found_taxa, common_name, rank, 
         decimal_latitude, decimal_longitude,
         region_code, site_code, location_name) %>%
  distinct() %>%
# there are NAs at the moment, delete later
  filter(!(is.na(decimal_latitude))) %>%
  filter(!(is.na(found_taxa))) # %>%
  #left_join(region_label)
#save if needed
write.csv(mifish_coords2, here(edna_out, paste0(datefield, runfield, 'fish-coords-uniques_regions.csv')))




species_list <- fishonly_range%>%
  select(found_taxa, common_name, rank) %>%
  filter(rank == 'species') %>%
  distinct()


# species list
write.csv(species_list, here(edna_out, paste0(datefield, runfield, 'species-list.csv')))
```


# phyloseq test

Phyloseq objects contain otu table, sample_data, and taxonomy table among other things.

What is the best level to normalize our data at? ; species, motu, asv, genus?

metadata needed for blind = false
next vs miseq
year
library
field blank vs sample

vst or rlog, probably vst for the time factor
assay to extract values



```{r phyloseq motus}
library(phyloseq)
library(DESeq2)
library(pheatmap)
library(vsn)
library(ape)

# import otu table
# requires numeric matrix
fishonly_otus <- asv_all %>%
 filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  filter(treatment %in% c('Sample')) %>%
  select(sample_id, motu, reads_raw) %>%
  pivot_wider(names_from = sample_id,
              values_from = reads_raw,
              values_fn = sum) %>%
  # NA to 0
  mutate(across(.cols = everything(), ~replace(., is.na(.), 0))) %>%
  column_to_rownames('motu') %>%
  as.matrix()

# there weren't any 0 columns?
#fishonly_otus[,-(which(colSums(x)==0))] 

fishonly_sample <- asv_all %>%
  filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  select(sample_id, year, region_code) %>% 
  # a few have no treatment, look into later
  #  mutate(treatment_deseq = case_when(is.na(treatment) ~ 'Sample',
   #                            treatment == 'Field blank' ~ 'field_blank',
    #                           TRUE ~ treatment)) %>%
  distinct() %>%
  # make sure sample names match otu table
  column_to_rownames('sample_id') %>%
  filter(rownames(.) %in% colnames(fishonly_otus)) %>%
  mutate(seq_method = ifelse(year == 2023, 'nextseq', 'miseq')) %>%
  mutate(year = as.factor(year))

# this is from the deseq function, 
# if we want to keep in the "bad" samples we need to add a pseudocount of 1 to all,
# and tbh we might need this simply because eDNA is not RNAseq 
# and every species will fail to show up in at least one sample.
# The same tweak as for edgeR to avoid NaN problems
# that cause the workflow to stall/crash.
fishonly_otus = fishonly_otus + 1

fishotu_dsdata = DESeqDataSetFromMatrix(countData = fishonly_otus, 
                             colData=fishonly_sample, 
                             #design = ~ treatment_deseq + seq_method
                             design = ~ year
                             )

vsd_otu <- varianceStabilizingTransformation(fishotu_dsdata)

## Taxa level

# import otu table
# requires numeric matrix
fishonly_taxa_otus <- asv_all %>%
 filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  filter(treatment == 'Sample') %>%
  select(sample_id, found_taxa, reads_raw) %>%
  pivot_wider(names_from = sample_id,
              values_from = reads_raw,
              values_fn = sum) %>%
  # NA to 0
  mutate(across(.cols = everything(), ~replace(., is.na(.), 0))) %>%
  column_to_rownames('found_taxa') %>%
  as.matrix()

fishonly_taxa_sample <- asv_all %>%
  filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  select(sample_id, year, region_code) %>%
  #mutate(treatment_deseq = case_when(is.na(treatment) ~ 'Sample',
   #                            treatment == 'Field blank' ~ 'field_blank',
    #                           TRUE ~ treatment)) %>%
  distinct() %>%
  # make sure sample names match otu table
  column_to_rownames('sample_id') %>%
  filter(rownames(.) %in% colnames(fishonly_taxa_otus)) %>%
  mutate(seq_method = ifelse(year == 2023, 'nextseq', 'miseq')) %>%
  mutate(year = as.factor(year))

# this is from the deseq function, 
# if we want to keep in the "bad" samples we need to add a pseudocount of 1 to all,
# and tbh we might need this simply because eDNA is not RNAseq 
# and every species will fail to show up in at least one sample.
# The same tweak as for edgeR to avoid NaN problems
# that cause the workflow to stall/crash.
fishonly_taxa_otus = fishonly_taxa_otus + 1

fishtaxa_dsdata = DESeqDataSetFromMatrix(countData = fishonly_taxa_otus, 
                             colData=fishonly_taxa_sample, 
                             design = ~ year)

vsd_taxa <- varianceStabilizingTransformation(fishtaxa_dsdata)

# plots
meanSdPlot(fishonly_otus)
meanSdPlot(assay(vsd_otu))

pheatmap(fishonly_otus, 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Un-relativized MOTUs')
pheatmap(assay(vsd_otu), 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'MOTUs with VST, DESeq2')


meanSdPlot(fishonly_taxa_otus)
meanSdPlot(assay(vsd_taxa))

pheatmap(fishonly_taxa_otus, 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Un-relativized taxa')
pheatmap(assay(vsd_taxa), 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Taxa with VST, DESeq2')

```

## attach to data
```{r}
wis_taxa <- vegan::wisconsin(fishonly_taxa_otus)
wis_otu <- vegan::wisconsin(fishonly_otus)

wis_taxa_tomerge <- as.data.frame(wis_taxa) %>%
    rownames_to_column('found_taxa') %>%
  pivot_longer(cols = !found_taxa,
               names_to = 'sample_id',
               values_to = 'relabund_wis_taxa')

wis_otu_tomerge <- as.data.frame(wis_otu) %>%
    rownames_to_column('motu') %>%
  pivot_longer(cols = !motu,
               names_to = 'sample_id',
               values_to = 'relabund_wis_motu')
  


vst_taxa_tomerge <- as.data.frame(assay(vsd_taxa)) %>%
  rownames_to_column('found_taxa') %>%
  pivot_longer(cols = !found_taxa,
               names_to = 'sample_id',
               values_to = 'relabund_vst_taxa') %>%
  mutate(relabund_vst_taxa = ifelse(relabund_vst_taxa < 0, 0, relabund_vst_taxa))

vst_otu_tomerge <- as.data.frame(assay(vsd_otu)) %>%
  rownames_to_column('motu') %>%
  pivot_longer(cols = !motu,
               names_to = 'sample_id',
               values_to = 'relabund_vst_motu') %>%
  mutate(relabund_vst_motu = ifelse(relabund_vst_motu < 0, 0, relabund_vst_motu))


fishonly_relabund <- fishonly_range %>%
  filter(treatment == 'Sample') %>%
  left_join(vst_taxa_tomerge) %>%
  left_join(vst_otu_tomerge)%>%
  left_join(wis_taxa_tomerge) %>%
  left_join(wis_otu_tomerge) %>%
  relocate(relabund_vst_taxa, relabund_vst_motu, relabund_wis_taxa, relabund_wis_motu,
           motu_pa, .before = reads_raw) %>%
  select(!c(n_dups, fish_base_url, group))

write.csv(fishonly_relabund, here(edna_out, paste0(datefield, runfield, 'asvmatrix_fish.csv')))
```



## Compare to other methods


```{r}
#this function calculates the eDNA index on a taxonomy by sample matrix from raw read numbers
#Author: Anya Mueller
#adapted from (Jacobs-Palmer et al, 2020)

edna_index <- function(df, row_var) {
  #df = a dataframe with taxonomy/ASV in columns, samples in rows
  #row_var = variable from the dataframe of desired row-wise identifier
  
  ##prep data
  #make dataframe
  x <- as.data.frame(df)
  #put row_var into rownames
  rownames(x) <- x[[row_var]]
  #remove row_var column
  x[[row_var]] <- NULL
  #remove rows who sum to zero
  x <- subset(x, rowSums(x)>0) 
  
  ##Take proportions from table of raw read counts 
  #apply operation to a matrix
  x <- sweep(x, 
             #do it by row
             MARGIN = 1,
             #value used for operation (sum by sample)
             STATS = rowSums(x), 
             #divide raw read number by row sum
             FUN = "/")
  
  #remove columns who sum to zero
  x[,-(which(colSums(x)==0))] 
  
  ##Take indices using a table with proportions of reads 
  #apply operation to a matrix
  x <- sweep(x, 
             #do it by column
             MARGIN = 2, 
             #value used for operation (row max)
             STATS =
               apply(x, 
                     #do it by column
                     MARGIN = 2, 
                     #take max of column (max proportion for taxonomy/ASV)
                     max), 
             #divide proportion by column max
             FUN = "/")
  
  ##add back row variable
  x[[row_var]] <- rownames(x)
  rownames(x) <- NULL
  
  return(x)
}

mifish_asvmat_pfs <- list()
for(i in c(0, 0.05, 0.1)){
  mifish_asvmat_pfs[[paste(paste("prev",i, sep = "-"), "mifish", sep = "-")]] <- mifish_asvmatrix_taxonomy %>%
    filter(!(found_taxa %in% c("No BLAST hit")),
           asv_prevalence > i) %>%
    select(asv,
           sample_id,
           asv_read_raw) %>%
    pivot_wider(names_from = "asv",
                values_from = "asv_read_raw")
}

```


```{r}
#wis_otu <- vegan::wisconsin(fishonly_otus)
wis_taxa <- vegan::wisconsin(fishonly_taxa_otus)

#revwis_taxa <- fishonly_otus


meanSdPlot(fishonly_taxa_otus)
meanSdPlot(assay(vsd_taxa))
meanSdPlot(wis_taxa)

pheatmap(fishonly_taxa_otus, 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Un-relativized taxa')
pheatmap(assay(vsd_taxa), 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Taxa with VST, DESeq2')
pheatmap(wis_taxa, 
         show_rownames = FALSE, show_colnames = FALSE,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'wisconsin with vegan')
```


```{r save heatmaps}
pdf(file = here(edna_out, 'untransformed_large.pdf'),
    width = 50, height = 8.5)
pheatmap(fishonly_taxa_otus, 
         fontsize = 5,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Untransformed',
         na_col = 'white')
dev.off()

pdf(file = here(edna_out, 'vst_large.pdf'),
    width = 50, height = 8.5)
pheatmap(assay(vsd), 
         fontsize = 5,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'VST with DESeq2',
         na_col = 'white')
dev.off()

pdf(file = here(edna_out, 'wisconsin_large.pdf'),
    width = 50, height = 8.5)
pheatmap(wis, 
         fontsize = 5,
         cluster_cols = FALSE, cluster_rows = FALSE,
         main = 'Wisconsin with Vegan',
         na_col = 'black')
dev.off()
```




```{r pcoas}
sample_data <- fishonly_sample %>%
  rownames_to_column('sample_id')

fishonly_taxa_dist <- dist(t(fishonly_taxa_otus))
vsd_taxa_dist <- dist(t(assay(vsd_taxa)))
wis_taxa_dist <- dist(t(wis_taxa))

fishonly_taxa_dist2 <- as.matrix(fishonly_taxa_dist)
vsd_taxa_dist2 <- as.matrix(vsd_taxa_dist)
wis_taxa_dist2 <- as.matrix(wis_taxa_dist)


colors <- colorRampPalette( rev(RColorBrewer::brewer.pal(9, "Blues")) )(255)
pheatmap(fishonly_taxa_dist2,
         col = colors)
pheatmap(vsd_taxa_dist2,
         col = colors)
pheatmap(wis_taxa_dist2,
         col = colors)

fishonly_pcoa <- pcoa(fishonly_taxa_dist2)
fishonly_pcoa_plot <- as.data.frame(fishonly_pcoa$vectors) %>% 
  janitor::clean_names() %>%
  select(axis_1, axis_2) %>%
  rownames_to_column('sample_id') %>%
  left_join(sample_data)

vsd_taxa_pcoa <- pcoa(vsd_taxa_dist2)
vsd_taxa_pcoa_plot <- as.data.frame(vsd_taxa_pcoa$vectors) %>% 
  janitor::clean_names() %>%
  select(axis_1, axis_2) %>%
  rownames_to_column('sample_id') %>%
  left_join(sample_data)

wis_taxa_pcoa <- pcoa(wis_taxa_dist2)
wis_taxa_pcoa_plot <- as.data.frame(wis_taxa_pcoa$vectors) %>% 
  janitor::clean_names() %>%
  select(axis_1, axis_2) %>%
  rownames_to_column('sample_id') %>%
  left_join(sample_data)
  

ggplot(fishonly_pcoa_plot, aes(x = axis_1, y = axis_2)) +
  geom_point(aes(x = axis_1, y = axis_2,
                 color = as.factor(year))) +
  labs(title = 'Unadjusted, by taxa') +
  theme_bw()

ggsave(here(edna_out, 'qc', 'unadjusted_pcoa.png'))

ggplot(vsd_taxa_pcoa_plot, aes(x = axis_1, y = axis_2)) +
  geom_point(aes(x = axis_1, y = axis_2,
                 color = as.factor(year))) +
  labs(title = 'VST with DESeq2, by taxa') +
  theme_bw()

ggsave(here(edna_out, 'qc', 'vst_pcoa.png'))

ggplot(wis_taxa_pcoa_plot, aes(x = axis_1, y = axis_2)) +
  geom_point(aes(x = axis_1, y = axis_2,
                 color = as.factor(year))) +
  labs(title = 'Wisconsin with vegan, by taxa') +
  theme_bw()

ggsave(here(edna_out, 'qc', 'wisconsin_pcoa.png'))

plotPCA(fishonly_taxa_dist)
plotPCA(vsd_taxa, intgroup = c('year'))
```


# scratch

```{r deseq function}
deseq_varstab = function(physeq, sampleConditions=rep("A", nsamples(physeq)), ...){
	require("DESeq2")
	# Enforce orientation.
	if( !taxa_are_rows(physeq) ){
		physeq <- t(physeq)
	}
	x = as(otu_table(physeq), "matrix")
	# The same tweak as for edgeR to avoid NaN problems
	# that cause the workflow to stall/crash.
	x = x + 1
	# Create annotated data.frame with the taxonomy table, in case it is useful later
	taxADF = as(data.frame(as(tax_table(physeq), "matrix"), stringsAsFactors=FALSE), "AnnotatedDataFrame")
	cds = DESeqDataSetFromMatrix(x, sampleConditions, featureData=taxADF)
	#cds = newCountDataSet(x, sampleConditions, featureData=taxADF)
	# First estimate library size factors
	cds = estimateSizeFactors(cds)
	# Variance estimation, passing along additional options
	cds = estimateDispersions(cds, ...)
	# Determine which column(s) have the dispersion estimates
	dispcol = grep("disp\\_", colnames(fData(cds)))
	# Enforce that there are no infinite values in the dispersion estimates
	if( any(!is.finite(fData(cds)[, dispcol])) ){
		fData(cds)[which(!is.finite(fData(cds)[, dispcol])), dispcol] <- 0.0
	}
  vsmat = exprs(varianceStabilizingTransformation(cds))
  otu_table(physeq) <- otu_table(vsmat, taxa_are_rows=TRUE)
  return(physeq)
}
```

ˆ
```{r}
# import otu table
# requires numeric matrix
fishonly_taxa_otus <- asv_all %>%
 filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  select(sample_id, found_taxa, reads_raw) %>%
  pivot_wider(names_from = sample_id,
              values_from = reads_raw,
              values_fn = sum) %>%
  # NA to 0
  mutate(across(.cols = everything(), ~replace(., is.na(.), 0))) %>%
  column_to_rownames('found_taxa') %>%
  as.matrix()

fishonly_taxa_taxonomy <- asv_all %>%
  filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  select(found_taxa, found_taxa, genus, family, order, class) %>%
  distinct() %>% 
  column_to_rownames('found_taxa') %>%
  # make sure the same 'species' are in each
  filter(rownames(.) %in% rownames(fishonly_taxa_otus)) %>%
  as.matrix()

fishonly_taxa_sample <- asv_all %>%
  filter(class %in% c('Teleostei', 'Chondrostei', 'Holocephali', 'Elasmobranchii')) %>%
  select(sample_id, year, region_code, treatment) %>%
  mutate(treatment_deseq = case_when(is.na(treatment) ~ 'Sample',
                               treatment == 'Field blank' ~ 'field_blank',
                               TRUE ~ treatment)) %>%
  distinct() %>%
  # make sure sample names match otu table
  column_to_rownames('sample_id') %>%
  filter(rownames(.) %in% colnames(fishonly_taxa_otus))



fishtaxa_phylo <- merge_phyloseq(fishtaxa_otutable, fishtaxa_samples,
                                 fishtaxa_taxonomy)


fishmotu_otutable <- otu_table(fishonly_otus, taxa_are_rows = TRUE)
fishmotu_samples <- sample_data(fishonly_sample)
fishmotu_taxonomy <- tax_table(fishonly_taxonomy)

fishmotu_phylo <- merge_phyloseq(fishmotu_otutable, fishmotu_samples,
                                 fishmotu_taxonomy)


fishmotu_deseq <- deseq_varstab(fishmotu_phylo)



fishotu_deseq <- DESeq(fishotu_dsdata)

```




```{r}
exblanks_2021 <-asvs2021 %>%
  select(sample_id, library, extraction_blank, region_name) %>%
  distinct()
```

