---
title: "ch2_02a_edna-qc"
author: "Kate Sheridan"
date: "11/22/2022"
---

Updated to switch back to ASVs and with merged metadata format October 7
Updated for MOTUs and lulu September 25, 2024
Updated with new metadata October 04, 2024

```{r setup, include=FALSE}
library(tidyverse)
library(here)

# here filepaths
# In
fastq <- here('rawdata', 'peco', 'fastq', 
              '12S_PECO_2021_Redos_Run20250416',
              'Analysis', '1', 'Data', 
              'fastq', 'fixed', 'processed') 
# Out
edna_combined <- here('processeddata', 'peco', '2021-24_edna')
edna_out <- here('processeddata', 'peco', '2021_edna')
# fields
datefield_out <- '20250711_'
runfield <- 'peco-2021_12s_'
```

# Load In

In: taxonomy from BLAST, asv table, metadata

Updating sample names at the end to keep from having to rewrite the whole script

```{r load-in}
metadata <- read_csv(here(edna_combined, '20240925_peco-combined_metadata.csv')) %>%
  filter(year == 2021) %>%
  mutate(bsaredo = ifelse(str_detect(sample_id, 'bsaredo'), 'redone', NA_character_)) %>%
  mutate(sample_id = str_remove_all(sample_id, '-bsaredo'))

taxonomy <- read_csv(here(edna_combined, '20250528_peco-combined_12s_taxonomy.csv'))

motu2asvs <- read_csv(here('rawdata', 'peco', 
                           'fastq', 'allyears', 
                           '20250528_peco-combined_12s_lulu_motu2asv.csv'))

asvtable <- read_delim(here(fastq,
                            '20250528_peco-combined_12s_sequence-table_asv-names_2021.txt'))

```

```{r merge}
motumatrix_metadata <- asvtable %>%
  # pivot to long so sequence can map to ASV
  pivot_longer(cols = !sample,
               names_to = 'asv') %>%
  filter(value >= 1) %>% 
  mutate(sample = tolower(sample)) %>%
  mutate(sample_id = sample) %>%
    #update blanks
  mutate(sample_id = case_when(str_detect(sample_id, 'pecoe-') ~ 
                                 str_replace_all(sample_id, 'pecoe-', 'pecoe-0'),
                               str_detect(sample_id, 'exblank-2021-') ~ 
                                 str_replace_all(sample_id, 'exblank-2021-', 'exblank-2021-0'),
                               TRUE ~ sample_id)) %>%
  filter(sample_id %in% metadata$sample_id) %>%
  left_join(taxonomy) %>%
  mutate(motu_pa = 1) %>%
  rename(reads_raw = value) %>%
  left_join(metadata) %>%
  #filter(!(treatment == 'Calibration')) %>%
  # remove columns that don't matter this year
  janitor::remove_empty() %>%
  # check before removing
  select(!c(motu_swarm, curated, lulu_rank)) 
```

# eDNA data

clean, matrix, singletons, pcr dissimilarity

## Lab controls

Subtracting reads from pcr / extraction controls

labs; 
extraction negative -- group = extraction date
calibration -- group == library. samples that were run twice, could look at library level differences in read number etc.
pcr negative -- group == library
Field blank -- group = site
libraries: BSA_redo, Library1, Library2
samples not in OSF

bsa-redo not populated in other columns

Can't do this until I have info about which blanks are for which samples.
```{r control-subtract}
pcr_ctrl <- motumatrix_metadata %>%
  filter(treatment %in% c('PCR blank')) %>%
  select(asv, reads_raw, treatment, library, found_taxa) %>%
  # make single code that combines site and asv
  mutate(pcrasv = paste0(library, '_', asv)) %>%
  select(pcrasv, reads_raw, found_taxa) %>%
  rename(reads_control = reads_raw)


extract_ctrl <- motumatrix_metadata %>%
  filter(treatment %in% c('Extraction blank')) %>%
  select(asv, reads_raw, treatment, extraction_blank, found_taxa) %>%
  filter(!(is.na(extraction_blank))) %>%
  # make single code that combines site and motu
  mutate(extractasv = paste0(extraction_blank, '_', asv)) %>%
  select(extractasv, reads_raw, found_taxa) %>%
  rename(reads_control = reads_raw)

motumatrix_metadata_nolab <- motumatrix_metadata %>%
  # insert minus part here
  # subtract extraction controls
  # single code to combine site and motu
  mutate(extractasv = paste0(extraction_blank, '_', asv)) %>%
  # merge to have field control values for each site-motu combo
  left_join(extract_ctrl) %>%
  mutate(in_ctrl_extract = ifelse(reads_control > 0, 'yes', NA_character_)) %>%
  #subtract control reads from regular reads
  mutate(reads_adjust = reads_raw - reads_control) %>%
  mutate(reads_adjust = coalesce(reads_adjust, reads_raw)) %>%
  mutate(reads_adjust = ifelse(reads_adjust <= 0, 0, reads_adjust)) %>%
  # only samples that still have reads
  #filter(reads_adjust > 0) %>%
  select(!(c(reads_control, extractasv))) %>%
  # subtract pcr controls by library
  # single code to combine library and motu
  mutate(pcrasv = paste0(library, '_', asv)) %>%
  # merge to have field control values for each site-motu combo
  left_join(pcr_ctrl) %>%
  mutate(in_ctrl_pcr = ifelse(reads_control > 0, 'yes', NA_character_)) %>%
  #subtract control reads from regular reads
  mutate(reads_adjust_lab = reads_adjust - reads_control) %>%
  mutate(reads_adjust_lab = coalesce(reads_adjust_lab, reads_adjust)) %>%
  mutate(reads_adjust_lab= ifelse(reads_adjust_lab <= 0, 0, reads_adjust_lab)) %>%
  # only samples that still have reads
  #filter(reads_adjust_pcr > 0) %>%
  select(!(c(reads_control, reads_adjust, pcrasv))) %>%
  #rename(reads_raw = reads) %>%
  # n site code = lab control
  filter(!(is.na(site_code))) %>%
  mutate(ctrl_lab0 = ifelse(reads_adjust_lab == 0, 'yes', 'no')) %>%
  relocate(in_ctrl_pcr, in_ctrl_extract, reads_adjust_lab, ctrl_lab0, .before = kingdom) %>%
  distinct()
```




## Field controls

Here we are subtracting reads in field controls from all samples in that site.**

Adjusted to annotate ASVs in field controls and whether they'd be removed in QC with subtracting reads

```{r field-control}
field_ctrl <- motumatrix_metadata_nolab %>%
  filter(treatment == 'Field blank') %>%
  select(site_code, asv, reads_raw, found_taxa) %>%
  # make single code that combines site and motu
  mutate(siteasv = paste0(site_code, '_', asv)) %>%
  select(siteasv, reads_raw, found_taxa) %>%
  rename(reads_control = reads_raw)

motumatrix_metadata_nofield <- motumatrix_metadata_nolab %>%
  # single code to combine site and motu
  mutate(siteasv = paste0(site_code, '_', asv)) %>%
  # merge to have field control values for each site-motu combo
  left_join(field_ctrl) %>%
  #subtract control reads from regular reads
  mutate(reads_adjust = reads_adjust_lab - reads_control) %>%
  mutate(reads_adjust = coalesce(reads_adjust, reads_adjust_lab)) %>%
  mutate(reads_adjust = ifelse(reads_adjust <= 0, 0, reads_adjust)) %>%
  mutate(ctrl_fieldlab0 = ifelse(reads_adjust == 0, 'yes', 'no')) %>%
  mutate(in_ctrl_field = ifelse(reads_control > 0, 'yes', 'no')) %>%
  # only samples that still have reads
  #filter(reads_adjust > 0) %>%
  #subtract control reads from raw reads
  mutate(reads_adjust_field = reads_raw - reads_control) %>%
  mutate(reads_adjust_field = coalesce(reads_adjust_field, reads_raw)) %>%
  mutate(reads_adjust_field = ifelse(reads_adjust_field <= 0, 0, reads_adjust_field)) %>%
  select(!(c(reads_control, siteasv))) %>%
  #relocate(reads_adjust, .before = reads_raw) %>%
  mutate(ctrl_field0 = ifelse(reads_adjust_field == 0, 'yes', 'no')) %>%
  relocate(in_ctrl_field, reads_adjust_field, ctrl_field0, ctrl_fieldlab0, .before = reads_adjust_lab) %>%
  select(!c(n_dups))

write_csv(motumatrix_metadata_nofield, here(edna_out, 
                                              paste0(datefield_out, runfield,
                                                     'asvmatrix_complete.csv')))
```

# Taxonomy
Filter to fish only then send to species setup

```{r taxonomy-filter}
#check options
#unique(motumatrix_metadata_nosingles$class)

#select only fish
motumatrix_metadata_fish <- motumatrix_metadata_nofield %>%
  filter(class %in% c('Teleostei', 'Holocephali', 'Elasmobranchii'))

write_csv(motumatrix_metadata_fish, here(edna_out, 
                                         paste0(datefield_out, 
                                                runfield, 'asvmatrix_fish.csv')))

unique(motumatrix_metadata_nofield$class)
```



## collect problems

also need to put in found taxa == NA, these are likely the below-species level IDs
```{r problems}
problems <- motumatrix_metadata_nofield %>%
  distinct() %>%
  # depending on how you want to filter
  filter(!(motu %in% c(motumatrix_metadata_fish$motu))) %>%
  filter(!(class %in% c('Teleostei', 'Holocephali', 'Elasmobranchii'))) %>%
  filter(common_name %in% c('No BLAST hit', 'Unknown'))



write_csv(problems, here(edna_out, paste0(datefield_out, 
                                          runfield, 'problems-motus.csv')))

```



## save controls

```{r}
all_ctrl <- motumatrix_metadata %>%
  filter(treatment %in% c('Extraction blank', 'PCR blank', 'Field blank')) %>%
  select(sample_id, treatment, motu, reads_raw, 
         found_taxa, common_name, site_code, 
         region_name, region_code, decimal_latitude, decimal_longitude,
         genus, family, order, class)
  

write_csv(all_ctrl, here(edna_out, 
                         paste0(datefield_out, runfield, 'controls.csv')))
```

