---
title: "ch2_00d_BLAST"
output: html_document
date: "2023-10-10"
---

Changes Sept 2024:
We're greatly reducing our usage of ASVs moving forward so I've eliminated some of the unnecessary steps to add in asv conversions.

Blast modified from Evan's script to use system2

```{r}
library(tidyverse)
library(taxize)
library(here)

# here filepaths
pecofastq <- here('rawdata', 'peco', 'fastq', 'allyears') 
datefield <- '20250528_'
datefield_new <- '20250528_'
# if different
datefield_swarm <- '20250528_'
runfield <- 'peco-combined_12s_'

# Destination folders
processededna <- here('processeddata', 'peco', '2021-24_edna')
```

# blast

Evan's call:

mkdir blast_96_sim_LCA_besthit
blastn 
-task megablast 
-num_threads 38 
-evalue 1e-5 
-max_target_seqs 10 
-perc_identity 96 
-qcov_hsp_perc 50 
-db /data/taxonomyDBs/NCBI_NT/2023-06-08/nt 
-outfmt '6 qseqid stitle sacc staxid pident qcovs evalue bitscore' 
-query 12S.PECO_combined.ASV_sequences.fasta  
-out blast_96_sim_LCA_besthit/12S_ASV_sequences.length_var.blast.out


Note that you need to have a blast database set up locally; this is not an api call

```{r run-blast}
#path.blast <- file.path(here(calvertfastq, "blast_96sim"))
#if(!dir.exists(path.blast)) dir.create(path.blast)

mifishdb <- file.path("/Users/katesheridan/Documents/GitHub/taxonomy_keys/database12s/crabs/blastdb_make/crabs-mifish")


edna_to_blast <- file.path(pecofastq,
                           '20250528_peco-combined_12s_post-lulu-motus.fasta')
blast_out <- file.path(pecofastq, paste0(datefield_new, runfield, 'swarm_blastresults_motu.txt'))

# check its working
system2("blastn", args = "-h")

system2("blastn", args=c("-task", "'megablast'", # which kind of blast
                         #"-num_threads", 38, # for parallelizing
                         "-evalue", 1e-5,
                         "-max_target_seqs", 50,
                         "-perc_identity", 85,
                         "-qcov_hsp_perc", 50,
                         "-db", mifishdb, # directory of db to blast against
                         "-outfmt", "'6 qseqid pident evalue sacc stitle bitscore qcovs mismatch'",
                         "-query", edna_to_blast,
                         "-out", blast_out
                         ))



```



# cleanup

```{r cleanup-motus}
# top 10 list
motu_50blast <- read_delim(here(pecofastq, paste0(datefield_new, runfield, 'swarm_blastresults_motu.txt')),
                           delim = '\t', col_names = c("motu", 
                                                       "identity_percentage", "evalue", "accession", "hit_name",
                                                       "bitscore", "coverage", "n_mismatches")) %>%
  mutate(hit_name = trimws(hit_name)) %>%
  mutate(hit_name = str_remove_all(hit_name, "(root;)")) %>%
  separate_wider_delim(hit_name, delim = ";", names = c('kingdom', 'phylum', 'class', 'order',
                                                        'family', 'genus', 'species')) %>%
  mutate(species = str_replace_all(species, "_", " ")) %>%
  # fill in NAs
  mutate(across(kingdom:genus, ~ na_if(., "")))

# motu numbers to join with the blast results
motu_key <- read_csv(here(pecofastq, '20250528_peco-combined_12s_lulu_motu2asv.csv')) %>%
  rename(motu = motu_lulu)
```

# assignment

```{r keys}
# load-in keys
taxa_key <- read_csv(here('processeddata', 'sp_lists', 'keys',
                     '20241015_12s_crabs_taxonomy-key.csv')) %>%
  #select(!hybrid) %>%
  mutate(found_taxa = case_when(taxonomy_search %in% c('Brachymystax tsinlingensis',
                                                       'Macquaria wujalwujalensis') ~ taxonomy_search,
                                found_taxa == 'Oxyjulis californica' ~ 'Halichoeres californica',
                                TRUE ~ found_taxa)) 


common_name <- read_csv(here('processeddata', 'sp_lists', 'keys',
                             '20241001_fish_common-names.csv'))[-1] %>%
  rename(common_name_key = common_name)
```
 previous notes

 5 assign_tax
 taxonomy assignment by accession number
 --missing for custom names, same format as lineage input
 note content from db_import may need to be updated first
open a github issue later
crabs assign_tax --input pgaout.fasta --output taxassigned.tsv --acc2tax nucl_gb.accession2taxid --taxid nodes.dmp --name names.dmp


## top 50 taxonomy
```{r top hits}
# strip and replace higher taxonomy
# remove out of region and hybrids
blast50_updatetaxa <- motu_50blast %>%
    # remove columns that will be duplicated
  select(!c(kingdom, phylum, class, order, family, genus, species)) %>%
  left_join(taxa_key) %>%
  # remove known incorrect hits when good hits exist (from manual check)
  # species out of region
  filter(!(c(found_taxa %in% c('Amphisticus argenteus', 'Aspiolucius esocinus',
                            'Alcichthys elongatus', 'Alburnus alburnus',
                            'Albula pacifica',
                            'Acipenser mikadoi', 'Acipenser schrenckii',
                            'Acipenser dabryanus', 'Acantholabrus palloni',
                            'Acanthogobius stigmothonus', "Anisotremus virginicus",
                            'Anisotremus scapularis', 'Anisotremus surinamensis', 
                            'Acanthogobius hasta', 'Alosa alosa', 'Alosa agone', 'Alosa fallax',
                            'Ammodytes japonicus', 'Ammodytes marinus', 'Ammodytes tobianus',
                            'Ammodytes dubius', 'Ammodytes americanus', 
                            'Argyrosomus amoyensis', 'Argyrosomus japonicus',
                            'Anchoa mitchilli', 'Atheresthes evermanni', 
                            'Aplodinotus grunniens', 'Askoldia variegata',
                            #'Barbatula barbatula', 'Barbatula hispanica', 
                            'Brachygenys chrysargyreum', 'Ballerus sapa',
                            'Bairdiella chrysoura', 'Bero elegans',
                            'Bahaba taipingensis', 'Boesemania microlepis',
                            'Batrachocottus talievi', 'Batrachocottus nikolskii', 
                            'Batrachocottus multiradiatus', 'Brachyopsis segaliensis',
                            'Chondrostoma nasus', "Clariger papillosus",
                            'Chirolophis saitone', 'Chirolophis ascanii', 
                            'Chirolophis wui', 'Chirolophis japonicus',
                            'Cleisthenes pinetorum', 'Cleisthenes herzensteini',
                            'Collichthys lucidus',
                            'Cottus gobio', 'Comephorus dybowskii', 'Cottus bairdii',
                            'Cottus pollux', 'Cottus szanaga', 'Cottus poecilopus', 
                            'Cottus czerskii', 'Cottus perifretum', 'Cottus hangiongensis',
                            'Clupea harengus', 'Conodon nobilis', 
                            'Cottocomephorus grewingki', 'Cottocomephorus grewingkii',
                            'Cottocomephorus inermis', 'Comephorus baikalensis',
                            'Cosmocampus elucens', 'Cryptacanthodes bergi',
                            'Ctenogobius boleosoma', 'Ctenogobius saepepallens',
                            'Ctenogobius shufeldti', 'Ctenolabrus rupestris',
                            "Cryptacanthodes maculatus", 'Cynoscion nothus',
                            'Cynoscion arenarius', 'Cynoscion nebulosus',
                            'Cynoscion regalis',
                            'Ditrema jordani', 'Dicentrarchus labrax',
                            'Dicentrarchus punctatus',
                            'Ernogrammus hexagrammus', 'Eques lanceolatus',
                            'Enchelyopus cimbrius',
                            'Equetus punctatus', 'Evorthodus lyricus',
                            'Gadus morhua','Gadus finnmarchicus',
                            'Gasterosteus nipponicus', 'Gasterosteus wheatlandi',
                            'Glyptocephalus stelleri', 'Gymnogobius macrognathos',
                            'Glyptocephalus cynoglossus', 'Gasterosteus nipponicus',
                            'Hyperoplus immaculatus', 'Hippoglossus hippoglossus',
                            'Hexagrammos agrammus', 'Hexagrammos otakii', 
                            'Haemulon carbonarium', 'Haemulon flavolineatum',
                            'Haemulon macrostomum', 'Haemulon plumierii',
                            'Hypomesus japonicus', 'Haemulon sciurus',
                            'Huso dauricus',
                            'Hyperoplus lanceolatus', 'Hozukius emblemarius', 
                            'Haemulon aurolineatum', 'Hypsoblennius ionthas',
                            'Hypleurochilus geminatus', 'Hypleurochilus springeri',
                            'Kasatkia memorabilis', 'Leuciscus baicalensis',
                            'Leuciscus merzbacheri', 'Leiostomus xanthurus',
                            'Liparis fabricii',
                            'Liparis miostomus', 'Liparis tessellatus', 
                            'Liparis tanakae', 'Liparis inquilinus',
                            'Leuciscus leuciscus', 'Limanda limanda', 'Limanda sakhalinensis',
                            'Lepidopsetta mochigarei', 'Liopsetta pinnifasciata', 
                            'Leuciscus idus', 'Leuciscus oxyrrhis', 'Lutjanus griseus',
                            'Leuciscus burdigalensis', 'Lutjanus argentimaculatus',
                            'Myzopsetta ferruginea', 'Microstomus achne', 
                            # introduced on the west coast
                            #'Morone saxatilis',
                            'Myoxocephalus scorpius', 'Myoxocephalus aenaeus',
                            'Malacocottus gibber',
                            'Melanogrammus aeglefinus', 'Menticirrhus littoralis',
                            'Miichthys miiuy', 'Netuma thalassina',
                            # there is a gbif record in alaskan panhandle
                            #'Myoxocephalus jaok',
                            'Oncorhynchus kawamurae', 'Opisthocentrus tenuis',
                            'Oncorhynchus gilae', 'Opisthocentrus ocellatus',
                            'Opisthocentrus zonope',
                            'Oreochromis niloticus', 'Occella iburia',
                            'Orthopristis chrysoptera', 'Oxyurichthys cornutus',
                            'Parahucho perryi', 'Procottus major', 
                            'Pareques umbrosus', 'Parablennius thysanius',
                            'Parablennius yatabei', 'Pholidapus dybowskii',
                            'Paralichthys olivaceus', 'Paralichthys adspersus',
                            'Platichthys flesus', 'Pholis gunnellus', 
                            'Pholis nebulosa', 'Pholis crassispina', 
                            'Pholis picta', 'Peprilus medius', 
                            'Pogonias cromis', 'Pungitius laevis',
                            'Pomadasys stridens', 'Pomadasys argenteus',
                            'Pomadasys kaakan', 'Peprilus triacanthus',
                            'Peprilus burti', 'Peprilus paru',
                            'Psenopsis anomala',
                            'Rutilus rutilus', 'Scardinius erythrophthalmus',
                            'Salvelinus elgyticus', 'Salvethymus svetovidovi',
                            'Salvelinus boganidae', 'Salvelinus kuznetzovi',
                            'Salvelinus taranetzi', 'Salvlinus albus',
                            'Salvelinus schmidti', 'Salvelinus levanidovi',
                            'Salvelinus kronocius', 'Salvelinus drjagini',
                            'Salvelinus drjagini', 'Salvelinus albus',
                            'Salvelinus leucomaenis', 'Scomber scombrus',
                            'Scartella emarginata', 'Scartella cristata',
                            'Serranus scriba', 'Serranus cabrilla',
                            'Sebastes mentella', 'Sebastes hubbsi',
                            'Sebastes inermis', 'Sebastes wakiyai',
                            'Sebastes cheni', 'Sebastes pachycephalus',
                            'Sebastes joyneri', 'Sebastes baramenuke',
                            'Sebastes viviparus', 'Sebastes norvegicus',
                            'Sebastes mentella', 'Sebastes koreanus',
                            'Sprattus sprattus', #'Squalius cephalus', ?maybe?
                            'Squalus acanthias', 'Squalus japonicus',
                            'Squalus mitsukurii', 'Sciaenops ocellatus',
                            'Squalius carolitertii', 'Squalius pyrenaicus',
                            'Squalius alburnoides', 'Strongylura marina',
                            'Syngnathus fuscus', 'Stellifer lanceolatus',
                            'Syngnathus louisianae', 'Syngnathus floridae',
                            'Stichaeopsis nana', 'Sebastiscus marmoratus',
                            'Scomber colias', 'Stichaeopsis nevelskoi',
                            'Strongylura timucu', 'Stromateus brasiliensis',
                            'Taurocottus bergii', 'Thunnus thynnus',
                            'Trachurus picturatus', 'Tilesina gibbosa',
                            # biological control with atlatic salmon?
                            # could this be in runoff?
                            #'Tautogolabrus adspersus', 
                            'Tautoga onitis',
                            # invasive established in California
                            #'Tridentiger trigonocephalus',
                            'Ulvaria subbifurcata', 'Xenolumpenus longipterus',
                            # extinct
                            'Salmo pallayri',
                            # deep sea
                            'Embassichthys bathybius', # synonimized to microstomus
                            # arctic canada, gulf of alaska
                            'Arctogadus glacialis', 'Boreogadus saida',
                            'Myzopsetta proboscidea', 'Bryozoichthys lysimus',
                            # Gulf of California/Baja peninsula; may be seen sampling if expanded south
                            'Gillichthys seta', 'Scomber australasicus',
                            'Totoaba macdonaldi', 'Paralabrax auroguttatus',
                            'Haemulon flaviguttatum',
                            # freshwater?
                            'Morone chrysops', 'Morone americana'
                            #'Prosopium williamsoni', 'Rhinichthys cataractae'
                            )))) %>%
  # drop level on outgroups that dont matter
  mutate(found_taxa = case_when(genus %in% c('Bos', 'Anas', 'Pseudo-nitzschia') ~ genus,
                                TRUE ~ found_taxa)) %>%
    # filter out odd sequences
  filter(!(accession %in% c( # hybrids
                            'NC_082285', 'OQ656297', 'OQ656298', 
                            'OQ656299','LC145951', 'NC_027522',
                            'NC_034784', 'KP218514',
                            # genus level IDs
                            'MT025508', 'MW181768', 'LC578888',
                            'MW181767', 'MW181766', 'MZ597980',
                            #'LC277738', # incorrectly on the list?
                            'LC741215', 'LC772864',
                            'KF668288',
                            # family level IDs
                            'NC_026537', 'OQ846303',
                            'OQ789493'))) %>%
   # set it up exactly as needed
  select(motu, identity_percentage, evalue, found_taxa, common_name, rank, 
         genus, family, order, class, phylum, kingdom, worms_sciname,
         freshwater_only, accession, coverage, bitscore, 
         status, unacceptreason)





# the original pipeline assumed we were working from a matrix with taxonomy already assigned.
# Here we're going to assign to MOTUs from scratch
# First; assign all MOTUs with 100% one species
blast_assign <- blast50_updatetaxa %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(motu, found_taxa, identity_percentage, .keep_all = TRUE) %>%
  group_by(motu) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(motu,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
    # remove unknowns when sp % is < 100
  filter(!(sp_percent < 1 & found_taxa %in% c('Unknown bacteria', 'Unknown organism',
                                              'Unknown eukaryote', 'Unknown algae',
                                              'Unknown protist'))) %>%
  # how many blast hits
  group_by(motu) %>%
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(motu,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  # remove duplicate rows by percent
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(motu, found_taxa, sp_percent, .keep_all = TRUE) %>%
  # indicate keep rows with 100% same species
  mutate(one_hit = ifelse(sp_percent == 1, 'yes', 'no')) %>%
  group_by(motu, found_taxa) %>%
  ## hits with relatively rare frequency and low identity percentage
  mutate(unlikely = case_when((any(identity_percentage == 100) & any(identity_percentage <= 99.5)) |
                                (any(sp_percent <= .25) & any(identity_percentage <= 99.5)) | # check the 20% cutoff in top 10
                               (any(identity_percentage >= 99) & any(identity_percentage <= 97.5)) |
                                identity_percentage <= 96.5 ~ 'x')) %>%
  ungroup() %>%
  # remove unlikely hits
  filter(!(one_hit == 'no' & unlikely %in% c('x') & identity_percentage <= 98.5|
             identity_percentage <= 97)) %>%
  # how many hits are left?
  add_count(motu, name = 'hit_remain') %>%
  # filter for anything 100%/98% that passed the original filter
  group_by(motu) %>%
  mutate(unlikely = case_when((hit_remain > 1 & any(identity_percentage == 100) & any(identity_percentage <= 99.5)) ~ 'x')) %>%
  ungroup() %>%
  # remove unlikely hits
  filter(!(unlikely %in% c('x') & identity_percentage <= 99.5)) %>%
  # assign asvs with one hit remaining after unlikely and < 100% frequency
  # # how many hits are left?
  add_count(motu, name = 'hit_remain') %>%
  filter(hit_remain == 1) %>%
  select(!(c(hit_remain, unlikely, one_hit, sp_percent, num_blast, sp_freq)))
  
```

## Conflicts

If there isn't a single identification, use stricter criteria or drop to higher taxonomy
```{r conflicts}
# identify motus that have conflicts in their top 10
# This removes motus without conflicts, leaving only the potential problems.
blast_conflict <- blast50_updatetaxa %>%
  filter(!(motu %in% blast_assign$motu)) %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(motu, found_taxa, identity_percentage, .keep_all = TRUE) %>%
  group_by(motu, found_taxa) %>%
  mutate(max_sp_id = max(identity_percentage)) %>%
  ungroup() %>%
  mutate(max_keep = ifelse(max_sp_id == identity_percentage, 'keep', 'toss')) %>%
  filter(max_keep == 'keep') %>%
  group_by(motu) %>%
  # how many blast hits
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(motu,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
      # remove unknowns when sp % is < 100
  filter(!(sp_percent < 1 & found_taxa %in% c('Unknown bacteria', 'Unknown organism',
                                              'Unknown eukaryote', 'Unknown algae',
                                              'Unknown protist'))) %>%
  # how many blast hits
  group_by(motu) %>%
  add_count(name = 'num_blast') %>%
  ungroup() %>%
  group_by(motu,found_taxa) %>%
  # how often does that species come up
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  # remove duplicate rows by percent
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  distinct(motu, found_taxa, sp_percent, .keep_all = TRUE) %>%
    group_by(motu, found_taxa) %>%
  # remove hits with less than 98.5%% when an ID in the list has 100%,
  ## hits with relatively rare frequency and low identity percentage
  mutate(unlikely = case_when((any(identity_percentage == 100) & any(identity_percentage <= 99.5)) |
                                (any(sp_percent <= .2) & any(identity_percentage <= 99.5)) | # check the 20% cutoff in top 10
                               (any(identity_percentage >= 99) & any(identity_percentage <= 97.5)) |
                                 identity_percentage <= 96.5 ~ 'x')) %>%
  ungroup() %>%
  filter(!(unlikely %in% c('x') & identity_percentage <= 98 |
             identity_percentage <= 97)) %>%
  select(!(c(unlikely))) %>%
    #remove anything that's more than .5 lower
  group_by(motu) %>%
  # what is the highest identity percentage on that motu
  mutate(max_id = max(identity_percentage)) %>%
  ungroup() %>%
  # if the species max is .5% lower than the max
  # this seems to be about as far away as makes sense for this data
  # Note that maybe for other data it might be different.
  # May also need more solid empirical justification than 
  # my past experience handling NCBI data
  # Filter to the max
  mutate(point_diff = max_id - max_sp_id) %>%
  filter(point_diff < .5) %>%
  select(!(c(max_id, max_sp_id, point_diff))) %>%
    # recalculate species percent
  group_by(motu) %>%
  add_count(name = 'num_blast') %>%
  group_by(motu, found_taxa) %>%
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(motu) %>%
    # na handling
  mutate(na_gen = ifelse(any(is.na(genus)), 'yes', 'no'),
         na_fam = ifelse(any(is.na(family)), 'yes', 'no'),
         na_ord = ifelse(any(is.na(order)), 'yes', 'no'),
         na_class = ifelse(any(is.na(class)), 'yes', 'no'),
         na_phy = ifelse(any(is.na(phylum)), 'yes', 'no'),
         na_king = ifelse(any(is.na(kingdom)), 'yes', 'no')) %>%
  mutate(updated_taxon = case_when(length(unique(found_taxa)) == 1 ~ found_taxa,
                                   length(unique(genus)) == 1 & na_gen == 'no' ~ genus,
                                   length(unique(family)) == 1 & na_fam == 'no' ~ family,
                                   length(unique(order)) == 1 & na_ord == 'no' ~ order,
                                   length(unique(class)) == 1 & na_class == 'no' ~ class,
                                   length(unique(phylum)) == 1 & na_phy == 'no' ~ phylum,
                                   length(unique(kingdom)) == 1  & na_king == 'no'~ kingdom,
                                   na_king == 'yes' | length(unique(kingdom)) > 1 ~ 'Unknown organism')) %>%
  select(!c(na_gen, na_fam, na_ord, na_class, na_phy, na_king, 
            num_blast, sp_percent, sp_freq)) 

blast_lowmatch <- blast50_updatetaxa %>%
  filter(!(motu %in% blast_assign$motu | motu %in% blast_conflict$motu)) %>%
  # remove duplicate rows by freqency
  ## this will make the merge faster as it will be done on fewer unique rows with no info loss
  ## what is the highest identity percentage per species, to eliminate duplicates
  group_by(motu, found_taxa) %>%
  mutate(max_sp_id = max(identity_percentage)) %>%
  ungroup() %>%
  distinct(motu, found_taxa, max_sp_id, .keep_all = TRUE) %>%
  group_by(motu) %>%
  # what is the highest identity percentage on that motu
  mutate(max_id = max(identity_percentage)) %>%
  ungroup() %>%
  # if the species max is .5 points or about 99.5% lower than the max for the motu
  # eliminate that hit
  # Filter to the max
  mutate(point_diff = max_id - max_sp_id) %>%
  filter(point_diff < .5) %>%
  # recalculate species percent and remove 100s
  group_by(motu) %>%
  add_count(name = 'num_blast') %>%
  group_by(motu, found_taxa) %>%
  add_count(name = 'sp_freq') %>%
  ungroup() %>%
  mutate(sp_percent = sp_freq/num_blast) %>%
  group_by(motu) %>%
  mutate(low_match = ifelse(max_id < 96, 'yes', 'no')) %>%
  mutate(very_low_match = ifelse(max_id < 93, 'yes', 'no')) %>%
  mutate(extremely_low_match = ifelse(max_id < 90.99, 'yes', 'no')) %>%
    # if the number of flagged records always has the same genus, update. 
  # if the number of flaggd records always has the same family, update.
  # if the species percent is 100% and the species was actually not local, update
    # na handling
  mutate(na_gen = ifelse(any(is.na(genus)), 'yes', 'no'),
         na_fam = ifelse(any(is.na(family)), 'yes', 'no'),
         na_ord = ifelse(any(is.na(order)), 'yes', 'no'),
         na_class = ifelse(any(is.na(class)), 'yes', 'no'),
         na_phy = ifelse(any(is.na(phylum)), 'yes', 'no'),
         na_king = ifelse(any(is.na(kingdom)), 'yes', 'no')) %>%
  mutate(updated_taxon = case_when(length(unique(found_taxa)) == 1 & low_match == 'no' ~ found_taxa,
                                   length(unique(genus)) == 1 & na_gen == 'no' & 
                                     very_low_match == 'no' ~ genus,
                                   length(unique(family)) == 1 & na_fam == 'no' & 
                                     extremely_low_match == 'no' ~ family,
                                   length(unique(order)) == 1 & na_ord == 'no' ~ order,
                                   length(unique(class)) == 1 & na_class == 'no' ~ class,
                                   length(unique(phylum)) == 1 & na_phy == 'no' ~ phylum,
                                   length(unique(kingdom)) == 1  & na_king == 'no'~ kingdom,
                                   na_king == 'yes' | length(unique(kingdom)) > 1 ~ 'Unknown organism')) %>%
  ungroup() %>%
  select(!(c(max_id, max_sp_id, point_diff, 
             accession, na_gen, na_fam, na_ord, na_class, na_phy, na_king, 
             num_blast, sp_percent, sp_freq)))

```

## all nonnative hits

```{r}
hit_nohit <- motu_50blast %>%
  filter(!(motu %in% blast_conflict$motu | motu %in% blast_lowmatch$motu | motu %in% blast_assign$motu)) %>%
  group_by(motu) %>%
    mutate(updated_taxon = case_when(length(unique(genus)) == 1 ~ genus,
                                   length(unique(family)) == 1 ~ family)) %>%
  ungroup() %>%
  mutate(genus = case_when(family == updated_taxon ~ NA_character_,
                           TRUE ~ genus)) %>% 
  select(!c(accession, identity_percentage, evalue, 
            species, bitscore, coverage, n_mismatches)) %>%
  distinct()
```


## Update

Now we'll use updated taxa from the conflicts, and add a column for MOTU level ids based on seed asv
```{r update}
# drop uncertanties to next taxonomic level
blast_updates <- blast_conflict %>%
      # modify dataframe to be more useful
  select(motu, found_taxa, updated_taxon, common_name,
         identity_percentage, coverage, evalue, 
         bitscore, rank, kingdom, phylum, class, order, 
         family, genus, freshwater_only
         ) %>%
  rbind(blast_lowmatch) %>%
  rbind(hit_nohit) %>%
  distinct() %>%
  mutate(rank = case_when(rank == 'offtarget' ~ rank,
                          str_detect(updated_taxon, ' ') == TRUE ~ 'species',
                          updated_taxon == genus ~ 'genus',
                          updated_taxon == family ~ 'family',
                          updated_taxon == order ~ 'order',
                          updated_taxon == class ~ 'class',
                          updated_taxon == phylum ~ 'phylum')) %>%
  mutate(genus = case_when(rank %in% c('family', 'order')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                           TRUE ~ genus)) %>%
  mutate(family = case_when(rank %in% c('order')| 
                             str_detect(updated_taxon, 'Unknown')== TRUE ~ NA_character_,
                            TRUE ~ family)) %>%
  mutate(order = case_when(rank %in% c('class')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ order)) %>%
  mutate(class = case_when(rank %in% c('phylum')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ class)) %>%
  mutate(phylum = case_when(rank %in% c('kingdom')| 
                             str_detect(updated_taxon, 'Unknown') == TRUE ~ NA_character_,
                            TRUE ~ phylum)) %>%
  mutate(kingdom = case_when(str_detect(updated_taxon, 'Unknown organism') == TRUE ~ NA_character_,
                            TRUE ~ kingdom)) %>%
  distinct(motu, updated_taxon, .keep_all = TRUE) %>%
  rbind(blast_assign) %>%
  mutate(found_taxa = coalesce(updated_taxon,found_taxa)) %>%
  select(!c(updated_taxon, coverage, bitscore,
            accession, worms_sciname, status, unacceptreason)) %>%
  relocate(motu, found_taxa, common_name, rank) %>%
  arrange(motu) %>%
  # open group somewhere?
  ungroup()
```



# Assign

All our blast taxa should be updated now, we can assign

```{r assign new}
taxonomy_assigned_asv <- motu_key %>%
  full_join(blast_updates) %>%
  # update based on research
  mutate(found_taxa = case_when(found_taxa == 'Merluccius angustimanus' ~ 'Merluccius',
                                TRUE ~ found_taxa)) %>%
  mutate(rank = case_when(found_taxa %in% c('Merluccius') ~ 'genus',
                          TRUE ~ rank)) %>%
  left_join(common_name) %>%
  mutate(common_name = coalesce(common_name_key, common_name)) %>%
  relocate(common_name, .before = rank) %>%
  mutate(blast = ifelse(is.na(found_taxa), 'No BLAST hit', 'BLAST')) %>%
  mutate(found_taxa = coalesce(found_taxa, blast)) %>%
  mutate(common_name = case_when(found_taxa %in% c('No BLAST hit', 'Unknown organism',
                                                   'Cyanobacteria') ~ found_taxa,
                                 found_taxa %in% c('Eukaryota') ~ 'Unknown eukaryote',
                                 # algae
                                 is.na(common_name) & phylum %in% c('Rhodophyta', 'Chlorophyta', 'Bacillariophyta') |
                                 is.na(common_name) & class %in% c('Bolidophyceae', 'Cryptophyceae', 'Chrysophyceae', 
                                              'Dictyochophyceae', 'Dinophyceae',
                                              'Pelagophyceae', 'Trebouxiophyceae') ~ 'Algae',
                                 # marine mammals
                                 found_taxa %in% c('Phoca') ~ 'Seal',
                                 # leftover
                                 is.na(common_name) & kingdom == 'Bacteria' ~ kingdom,
                                 ## FISH ##
                                 # species
                                 found_taxa == 'Acipenser medirostris' ~ 'Green sturgeon',
                                 found_taxa == 'Acipenser transmontanus' ~ 'White sturgeon',
                                 found_taxa == 'Agonopsis vulsa' ~ 'Northern spearnose poacher',
                                 found_taxa == 'Amphistichus argenteus' ~ 'Barred surfperch',
                                 found_taxa == 'Alosa sapidissima' ~ 'American shad',
                                 found_taxa == 'Allosmerus elongatus' ~ 'Whitebait smelt',
                                 found_taxa == 'Anchoa delicatissima' ~ 'Slough anchovy',
                                 found_taxa == 'Anchoa compressa' ~ 'Deep body anchovy',
                                 found_taxa == 'Anisotremus davidsonii' ~ 'Xantic sargo',
                                 found_taxa == 'Atractoscion nobilis' ~ 'White seabass',
                                 found_taxa == 'Auxis rochei' ~ 'Bullet tuna',
                                 found_taxa == 'Bodianus pulcher' ~ 'California sheephead',
                                 found_taxa == 'Brachygenys californiensis' ~ 'Californian salema',
                                 found_taxa == 'Cebidichthys violaceus' ~ 'Monkeyface prickleback',
                                 found_taxa == 'Cheilotrema saturnum' ~ 'Black croaker',
                                 found_taxa == 'Clinocottus embryum' ~ 'Calico sculpin',
                                 found_taxa == 'Clinocottus analis' ~ 'Wooly sculpin',
                                 found_taxa == 'Cololabis saira' ~ 'Pacific saury',
                                 found_taxa == 'Cosmocampus arctus' ~ 'Snubnose pipefish',
                                 found_taxa == 'Cryptacanthodes aleutensis' ~ 'Dwarf wrymouth',
                                 found_taxa == 'Cryptacanthodes giganteus' ~ 'Giant wrymouth',
                                 found_taxa == 'Cynoscion parvipinnis' ~ 'Shortfin weakfish',
                                 found_taxa == 'Enophrys diceraus' ~ 'Antlered sculpin',
                                 found_taxa == 'Fundulus lima' ~ 'Baja California killifish',
                                 found_taxa == 'Fundulus parvipinnis' ~ 'California killifish',
                                 found_taxa == 'Genyonemus lineatus' ~ 'White croaker',
                                 found_taxa == 'Gillichthys mirabilis' ~ 'Longjaw mudsucker',
                                 found_taxa == 'Girella nigricans' ~ 'Rudderfish', # also called opaleye
                                 found_taxa == 'Girella simplicidens' ~ 'Gulf opaleye',
                                 found_taxa == 'Gobio gobio' ~ 'Gudgeon invasive',
                                 found_taxa == 'Gobiesox rhessodon' ~ 'California clingfish',
                                 found_taxa == 'Hemilepidotus spinosus' ~ 'Brown Irish lord',
                                 found_taxa == 'Heterostichus rostratus' ~ 'Giant kelpfish',
                                 found_taxa == 'Hexanchus griseus' ~ 'Bluntnose sixgill shark',
                                 found_taxa == 'Hypsoblennius jenkinsi' ~ 'Mussel blenny',
                                 found_taxa == 'Hypsoblennius gilberti' ~ 'Rockpool blenny',
                                 found_taxa == 'Hypsoblennius gentilis' ~ 'Bay blenny',
                                 found_taxa == 'Kyphosus azureus' ~ 'Zebra-perch sea chub',
                                 found_taxa == 'Lampetra richardsoni' ~ 'Western brook lamprey',
                                 found_taxa == 'Leuresthes tenuis' ~ 'California grunion',
                                 found_taxa == 'Lavinia exilicauda' ~ 'Hitch',
                                 found_taxa == 'Leuroglossus schmidti' ~ 'Northern smoothtongue',
                                 found_taxa == 'Liparis greeni' ~ 'Lobefin snailfish',
                                 found_taxa == 'Lycodes brevipes' ~ 'Shortfin eelpout',
                                 found_taxa == 'Lycodes diapterus' ~ 'Black eelpout',
                                 found_taxa == 'Nectoliparis pelagicus' ~ 'Tadpole snailfish',
                                 found_taxa == 'Mallotus villosus' ~ 'Capelin',
                                 found_taxa == 'Menticirrhus undulatus' ~ 'California kingcroaker',
                                 found_taxa == 'Myliobatis californica' ~ 'Bat ray',
                                 found_taxa == 'Myoxocephalus quadricornis' ~ 'Fourhorn sculpin',
                                 found_taxa == 'Notorynchus cepedianus' ~ 'Broadnose sevengill shark',
                                 found_taxa == 'Oncorhynchus clarkii' ~ 'Cutthroat trout',
                                 found_taxa == 'Oncorhynchus mykiss' ~ 'Rainbow trout',
                                 found_taxa == 'Oncorhynchus gorbuscha' ~ 'Pink salmon',
                                 found_taxa == 'Oxyjulis californica' ~ 'Senorita wrasse',
                                 found_taxa == 'Paraclinus integripinnis' ~ 'Reef finspot',
                                 found_taxa == 'Paralichthys californicus' ~ 'California halibut',
                                 found_taxa == 'Peprilus simillimus' ~ 'Pacific pompano',
                                 found_taxa == 'Phanerodon vacca' ~ 'Pile perch',
                                 found_taxa == 'Phanerodon furcatus' ~ 'White seaperch',
                                 found_taxa == 'Pleuronichthys guttulatus' ~ 'Diamond turbot',
                                 found_taxa == 'Pleuronichthys decurrens' ~ 'Curlfin sole',
                                 found_taxa == 'Pleuronichthys ritteri' ~ 'Spotted turbot',
                                 found_taxa == 'Podothecus accipenserinus' ~ 'Sturgeon poacher',
                                 found_taxa == 'Pogonichthys macrolepidotus' ~ 'Sacramento splittail',
                                 found_taxa == 'Psychrolutes paradoxus' ~ 'Tadpole sculpin',
                                 found_taxa == 'Ptychocheilus oregonensis' ~ 'Northern pikeminnow',
                                 found_taxa == 'Quietula y-cauda' ~ 'American shadow goby',
                                 found_taxa == 'Rhacochilus toxotes' ~ 'Rubberlip seaperch',
                                 found_taxa == 'Roncador stearnsii' ~ 'Spotfin croaker',
                                 found_taxa == 'Sardinops sagax' ~ 'South American pilchard',
                                 found_taxa == 'Seriola lalandi' ~ 'Yellowtail amberjack',
                                 found_taxa == 'Sebastes jordani' ~ 'Shortbelly rockfish',
                                 found_taxa == 'Sebastes gilli' ~ 'Bronzespotted rockfish',
                                 found_taxa == 'Seriphus politus' ~ 'Queenfish',
                                 found_taxa == 'Scomber japonicus' ~ 'Chub mackerel',
                                 found_taxa == 'Spirinchus thaleichthys' ~ 'Longfin smelt',
                                 found_taxa == 'Spirinchus starksi' ~ 'Night smelt',
                                 found_taxa == 'Sphyraena argentea' ~ 'Pacific barracuda',
                                 found_taxa == 'Strongylura exilis' ~ 'Californian needlefish',
                                 found_taxa == 'Syngnathus auliscus' ~ 'Barred pipefish',
                                 found_taxa == 'Syngnathus californiensis' ~ 'Bay pipefish',
                                 found_taxa == 'Synchirus gilli' ~ 'Manacled sculpin',
                                 found_taxa == 'Trachurus symmetricus' ~ 'Pacific jack mackerel',
                                 found_taxa == 'Thaleichthys pacificus' ~ 'Eulachon',
                                 found_taxa == 'Thunnus alalunga' ~ 'Albacore',
                                 found_taxa == 'Thunnus obesus' ~ 'Bigeye tuna',
                                 found_taxa == 'Typhlogobius californiensis' ~ 'Blind goby',
                                 found_taxa == 'Umbrina roncador' ~ 'Yellowfin drum',
                                 # genera
                                 found_taxa == 'Atherinopsis' ~ 'Atherinopsis silverside species',
                                 found_taxa == 'Anisotremus' ~ 'Grunt species',
                                 found_taxa == 'Albula' ~ 'Bonefish species',
                                 found_taxa == 'Apodichthys' ~ 'Apodichthys gunnel species',
                                 found_taxa == 'Artedius' ~ 'Artedius sculpin species',
                                 found_taxa == 'Cymatogaster' ~ 'Cymastogater perch species',
                                 found_taxa == 'Enophrys' ~ 'Enophrys sculpin species',
                                 found_taxa == 'Embiotoca' ~ 'Embiotoca surfperch species',
                                 found_taxa == 'Eucyclogobius' ~ 'Tidewater goby species',
                                 found_taxa == 'Fundulus' ~ 'Killifish species',
                                 found_taxa == 'Genyonemus' ~ 'Croaker species',
                                 found_taxa == 'Gibbonsia' ~ 'Gibbonsia kelpfish species',
                                 # note hypomesius transpacificus is endangered and only around Sacramento
                                 found_taxa == 'Hypomesus' ~ 'Smelt species',
                                 found_taxa == 'Halichoeres' ~ 'Halichoeres wrasse species',
                                 found_taxa == 'Jordania' ~ 'Jordania sculpin species',
                                 found_taxa == 'Leptocottus' ~ 'Leptocottus sculpin species',
                                 found_taxa == 'Merluccius' ~ 'Hake species',
                                 found_taxa == 'Oreochromis' ~ 'Oreochromis cichlid species',
                                 found_taxa == 'Phoxinus' ~ 'Minnow species',
                                 found_taxa == 'Rimicola' ~ 'Rimicola kelp clingfish species',
                                 found_taxa == 'Ruscarius' ~ 'Roughcheek sculpin species',
                                 found_taxa == 'Thunnus' ~ 'Thunnus tuna species',
                                 # families
                                 found_taxa == 'Agonidae' ~ 'Poacher family',
                                 found_taxa == 'Cyprinidae' ~ 'Carp-Minnow family',
                                 found_taxa == 'Gobiidae' ~ 'Goby family',
                                 found_taxa == 'Labridae' ~ 'Wrasse family',
                                 found_taxa == 'Liparidae' ~ 'Snailfish family',
                                 found_taxa == 'Sciaenidae' ~ 'Drum family',
                                 found_taxa == 'Zoarcidae' ~ 'Eelpout family',
                                 # orders
                                 found_taxa == 'Gobiiformes' ~ 'Goby order',
                                 found_taxa == 'Teleostei' ~ 'Unknown fish',
                                 # invasive?
                                 found_taxa == 'Barbatula barbatula' ~ 'Stone loach invasive',
                                 found_taxa == 'Alosa pseudoharengus' ~ 'Alewife invasive',
                                 found_taxa == 'Esox lucius' ~ 'Northern pike invasive',
                                 found_taxa == 'Gambusia affinis' ~ 'Mosquitofish invasive',
                                 found_taxa == 'Morone saxatilis' ~ 'Striped bass invasive',
                                 found_taxa == 'Squalius cephalus' ~ 'European chub no records',
                                 found_taxa == 'Tautogolabrus adspersus' ~ 'Cunner no records',
                                 found_taxa == 'Tridentiger trigonocephalus' ~ 'Chameleon goby invasive',
                                 TRUE ~ common_name)) %>%
  # some silliness
  mutate(found_taxa = case_when(found_taxa == 'Homo' ~ 'Homo sapiens',
                                TRUE ~ found_taxa)) %>%
  #filter(is.na(common_name)) %>%
  arrange(found_taxa) %>%
  select(!c(common_name_key, blast))
```



add a column for motu level ids, this way we can select columns based on the analysis we want to do
```{r add motus}
# add motu columns, relocate
taxonomy_assigned_full <- taxonomy_assigned_asv %>%
  relocate(found_taxa, common_name, rank, 
           .before = identity_percentage) %>%
  select(!c(low_match, very_low_match, extremely_low_match)) %>%
  distinct() %>%
  #TEMPORARY
  mutate(found_taxa = case_when(found_taxa == 'Atherinopsis' ~ 'Atherinops affinis',
                                 TRUE ~ found_taxa)) %>%
  mutate(rank = case_when(found_taxa %in% c('Atherinops affinis') ~ 'species',
                           TRUE ~ rank)) %>%
  mutate(common_name = case_when(found_taxa == 'Atherinops affinis' ~ 'Topsmelt silverside',
                                 TRUE ~ common_name))

write_csv(taxonomy_assigned_full, here(processededna, paste0(datefield_new, runfield,
                                                             'taxonomy_initial.csv')))


```

```{r not-in}
motu_reblast <- taxonomy_assigned_asv %>%
  mutate(to_reblast = case_when(found_taxa %in% c('No BLAST hit') ~ 'x',
                                extremely_low_match == 'yes' & rank != 'offtarget' ~ 'x',
                                very_low_match == 'yes' & rank != 'offtarget' ~ 'x',
                                rank %in% c('order', 'family') ~ 'x',
                                rank == 'genus' & is.na(common_name) ~ 'x')) %>%
  filter(!is.na(to_reblast)) %>%
  select(motu) %>%
  distinct()

write_delim(motu_reblast, here(processededna, paste0(datefield_new, runfield, 'to-reblast2.txt')),
          col_names = FALSE)
```

```{r}
taxonomyreduced <- taxonomy_assigned_full %>%
  select(motu, found_taxa, common_name, rank) %>%
  distinct()




# if the problems weren't problems
# 
# # 
write_csv(taxonomy_assigned_full, here(processededna, paste0(datefield_new, runfield,
                                                             'taxonomy.csv')))
```

