---
title: "ch2_00b_run-combine-2023"
output: html_document
date: "2024-06-04"
---

Note: this script adds the new sequences to the previously combined sequence_table


Combining runs for taxonomy assignment: all runs per year, all years.
Modified from Evan's combining_runs.dada2.12s.PECO.R

Sequence tables
Input: sequence tables from each run
Output: sequence table per year, master sequence table from all years

Read retention
Input: read retention log from each run
Output: read retention by year and total


NOTE ON FILEPATHS:
datefield by year is the date the run was last processed -- not today's date! That's regular `datefield`!

```{r setup, include=FALSE}
####Libraries####
library(tidyverse)
library(dada2)
library(seqinr)

library(here)
# here filepaths
# combined
fastq_combined <- here('rawdata', 'peco', 'fastq', 'allyears')
datefield <- '20241001_'
runfield <- 'peco-combined_12s_'

# 2021 + 2022
pecofastq2122 <- here('rawdata', 'peco', 'fastq', 
                  'allyears') 
datefield2122 <- '20231023_'
runfield2122 <- 'peco-combined_12s_'



# 2021
pecofastq2021 <- here('rawdata', 'peco', 'fastq', 
                  'peco2021_raw-miseq') 
datefield2021 <- '20230819_'
runfield2021 <- 'peco2021-combined_12s_'

# 2022
pecofastq2022 <- here('rawdata', 'peco', 'fastq', 
                  'peco2022_raw-miseq') 
datefield2022 <- '20231013_'
runfield2022 <- 'peco2022-combined_12s_'


# 2023
pecofastq2023 <- here('rawdata', 'peco', 'fastq', 
                  '12S_PECO2023_Run20240907', 'Analysis',
                  '1', 'Data', 'fastq', 'fixed', 'processed') 
datefield2023 <- '20241001_'
runfield2023 <- 'peco2023_12s_20240907_'

```

# eDNA data

## 2021


```{r merge2021}
####read in sequence tables####
#peco run 1
seqtab_2021_1 <- read_delim(here(pecofastq2021,
                                 '12S_PECO_Run1_20220829', 'processed',
                                 "20240611_12s_PECO_Run1_20220829_sequence-table_12S-merged.txt")) %>%
  filter(!(row_names == 'Undetermined')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

#peco run 2
seqtab_2021_2 <- read_delim(here(pecofastq2021,'12S_PECO_2021_Lib2_Run20221108', 
                                'processed', "20240611_12S_PECO_2021_Lib2_Run20221108_sequence-table_12S-merged.txt")) %>%
  filter(!(row_names == 'Undetermined')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

seqtab_2021_2redo <- read_delim(here(pecofastq2021,'12S_PECO_BSA_Redos_Run20230109', 'processed',
                                        "20240611_12s_peco_redos_run20230109_sequence-table_12S-merged.txt")) %>%
  filter(!(row_names == 'Undetermined')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

####merge sequence tables####
#ASVs are combined on sequence alone. exact matches are simply merged
seqtab_2021 <- mergeSequenceTables(seqtab_2021_1, seqtab_2021_2, seqtab_2021_2redo) 
```


## 2022

2022 had two runs on Miseq

```{r 2022}
#peco run 2
seqtab_2022_1 <- read_delim(here(pecofastq2022, 
                                 '12S_PECO2022_Run1_20230523', 'processed',
                                 '20221023_peco2022_12s_run1-20230523_sequence-table_12S-merged.txt')) %>%
  filter(!(row_names == 'Undetermined')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

#peco run 2
seqtab_2022_2 <- read_delim(here(pecofastq2022, 
                                 '12S_PECO2022_Run2_20230629', 'processed', 
                                '20221023_peco2022_12s_run2-20230629_sequence-table_12S-merged.txt')) %>%
  filter(!(row_names == 'Undetermined')) %>%
  column_to_rownames('row_names') %>%
  as.matrix()

####merge sequence tables####
#ASVs are combined on sequence alone. exact matches are simply merged
seqtab_2022 <- mergeSequenceTables(seqtab_2022_1, seqtab_2022_2) 
```


```{r merge2023}
# only one run for 2023
#peco 2023
seqtab_2023 <- read_delim(here(pecofastq2023,
                               "20241001_peco2023_12s_20240907_sequence-table_12S-merged.txt")) %>%
 filter(!(row_names == 'Undetermined')) %>%
 column_to_rownames('row_names') %>%
 as.matrix()
```

# Combine eDNA

All sequence tables should be merged into one master table. This means that every identical sequence in every year will be given the same ASV number. This master combined table should be exported and run through seqtable2swarm.py before moving on to clustering and taxonomy assignment.

```{r merge-all-tables}
####merge sequence tables####
#ASVs are combined on sequence alone. exact matches are simply merged
seqtab_all <- mergeSequenceTables(seqtab_2021, seqtab_2022, seqtab_2023) 


####save sequences as fasta, give sequences human-readable names, save modified sequence table####


#saving the new names and sequences as a .fasta file in your project working directory, and save a table that shows the mapping of sequences to new ASV names
asvtab_all <- t(as.data.frame(seqtab_all)) 
#transposed (OTUs are rows) data frame. 
# unclassing the otu_table() output avoids type/class errors later on

#re-save sequence table w updated names
write.table(data.frame("row_names"=rownames(seqtab_all), seqtab_all),
            here(fastq_combined,
                 paste0(datefield, runfield, "sequence-table.txt")), row.names=FALSE, quote=F, sep="\t")

ASV.seq <- as.character(unclass(row.names(asvtab_all))) #store sequences in character vector
ASV.num <- paste0("ASV", seq(ASV.seq), sep='') #create new names

asv_map <- cbind(asv = ASV.num, seq = ASV.seq) %>%
  as.data.frame()

write.table(asv_map, here(fastq_combined, 
                          paste0(datefield, runfield,
                                 "asv-sequences_name-mapping.txt")),
            sep="\t", quote=F, row.names=F, col.names=F)



write.fasta(sequences=as.list(ASV.seq), names=ASV.num, 
            here(fastq_combined, 
                 paste0(datefield, runfield, "asv-sequences.fasta"))) #save sequences with new names in fasta format



#IMPORTANT: sanity checks
colnames(seqtab_all) == ASV.seq #only proceed if this tests as true for all elements

#rename your ASVs in the taxonomy table and sequence table objects
colnames(seqtab_all) <- ASV.num

#re-save sequence table w updated names
write.table(data.frame("row_names"=rownames(seqtab_all),seqtab_all),
            here(fastq_combined,
                 paste0(datefield, runfield, "sequence-table_asv-names.txt")), row.names=FALSE, quote=F, sep="\t")

```

## ASV tables by year

These will be used for eDNA QC after taxonomy assignment.

```{r yearly sequence tables}
asvtab_2021 <- seqtab_2021 %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2021,
            here(pecofastq2021,
                 paste0(datefield, runfield, "sequence-table_asv-names_2021.txt")), row.names=FALSE, quote=F, sep="\t")

# 2022
asvtab_2022 <- seqtab_2022 %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2022 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2022,
            here(pecofastq2022,
                 paste0(datefield, runfield, "sequence-table_asv-names_2022.txt")), row.names=FALSE, quote=F, sep="\t")


# 2023
asvtab_2023 <- seqtab_2023 %>%
  as.data.frame() %>%
  rownames_to_column(var = 'sample') %>%
  pivot_longer(cols = !sample,
               names_to = 'seq') %>%
  left_join(asv_map) %>%
  select(!seq) %>%
  pivot_wider(names_from = asv)

# take seqtab 2021 and update to have the right names
#re-save sequence table w updated names
write.table(asvtab_2023,
            here(pecofastq2023,
                 paste0(datefield, runfield, "sequence-table_asv-names_2023.txt")), row.names=FALSE, quote=F, sep="\t")

```

# Read retention summary statistics

REDO 2023 FOR THESE FILES AND MERGE LATER


These files map where the reads went throughout the entire QC process

```{r load-in read retention}

# 2022 run 2
retention_2022_2 <- read_csv(here(pecofastq2022, 
                                   '12S_PECO2022_Run2_20230629', 'qc_plots',
                                   '20240603_peco2022_12s_run2-20230629_read-retention_12s_bysample.csv'))[-1] %>%
  janitor::clean_names() %>%
  mutate(year = 2022)



```

```{r merge-retention}

retention_all <- retention_2021_1 %>%
  bind_rows(retention_2021_2) %>%
  bind_rows(retention_2021_2redo) %>%
  bind_rows(retention_2022_1) %>%
  bind_rows(retention_2022_2) 

retention_summary <- retention_all %>%
  group_by(year) %>%
  summarize(n_samp = length(sample_id),
            raw = sum(raw_f) + sum(raw_r),
            after_cutadapt = sum(cutadapt_f) + sum(cutadapt_r),
            # get the rs for this on the summary
            after_filtertrim = sum(filtertrim_f) + sum(filtertrim_r),
            after_denoise = sum(denoised_f) + sum (denoised_r),
            merged = sum(merged),
            nosingletons = sum(nosingletons),
            nochimeras = sum(nochimeras)) %>%
  ungroup() %>%
  janitor::adorn_totals('row')


write.csv(retention_summary, here(fastq_combined, 
                                  paste0(datefield, runfield,
                                         "read_retention.csv")))
```




# Dereplicate FASTA test

Right now it looks like all these are unique?

```{r dereplicate}
library(ips)

asvtab_all <- t(as.data.frame(seqtab_all))

seq_list <- asvtab_all %>%
  as.data.frame() %>%
  rownames_to_column('asv_seq') %>%
  select(asv_seq) %>%
  mutate(temp_asv = paste0("tempasv_", seq(asv_seq))) %>%
  column_to_rownames('temp_asv')

write.fas(seq_list, here(fastq_combined, '20231023_tempasvs.fasta'))
seq_list2 <- read.fas(here(fastq_combined, '20231023_tempasvs.fasta'))

seq_align <- mafft(seq_list2)
write.fas(seq_align, here(fastq_combined, '20231023_tempasvs_align.fasta'))

seq_derep <- dada2:::derepFasta(here(fastq_combined, '20231023_tempasvs_align.fasta'))

uniq<-seq_derep$uniques %>%
  as.data.frame()

```

