---
title: "ch2-06a_get-environment"
author: "Kate Sheridan"
date: "`r Sys.Date()`"
output: html_document
---


Updating ch2-06_get-environment to go from erddap to summary stats.

This script uses ERDDAP to extract sea surface temperature statistics from the dataset:
NOAA ERD and CoastWatch West Coast Regional Node
Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present,
Monthly

https://upwell.pfeg.noaa.gov/erddap/info/jplMURSST41mday/index.html

Then matches the sites to the downloaded raster for each year

It also extracts the upwelling data using `erdBEUTIdaily`.
 TO ADD

```{r setup}
# needed to load data format
library(ncdf4)
# NOTE raster masks dplyr::select
library(raster)
# needed to parse times
library(CFtime)
library(sf)
# for rastrs
# # for rasters
library(terra)
library(tidyterra)

library(rerddap)
library(tidyverse)
library(here)

# temperature data
erd_data <- here('rawdata', 'temp', 'erddap')
erd_upwell <- here('rawdata', 'upwelling', 'erddap')
# eDNA
dataedna <- here('processeddata', 'peco', '2021-24_edna')
# for plots I don't think this makes plots? delete if not
#plot_out <- here('output', 'combined', 'ranges')

datefield = '20251212_'
```


# bounding boxes 
```{r bboxes}
bounding_boxes <- read_csv(here('rawdata', 'spatial',
                                '20250417_regionbbox.csv')) %>%
  janitor::clean_names() %>%
  filter(!is.na(box_name))

# turn bounding box into coordinates
bbox_sf <- bounding_boxes %>%
  # 4 corners of box
  pivot_longer(cols = c(min_lat, max_lat),
               values_to = 'latitude') %>%
  select(!name) %>%
  pivot_longer(cols = c(min_long, max_long),
               values_to ='longitude') %>%
  select(!name) %>%
  # corners to points
  st_as_sf(coords = c('longitude', 'latitude'),
                    crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
  group_by(box_name) %>%
  # make multipoints
  summarise(geometry = st_combine(geometry)) %>%
  ungroup() %>%
  # box around multipoints
  st_convex_hull()

```

# ERDDAP
```{r}
erd_info <- info('jplMURSST41')


# 2019
for (i in 1:nrow(bounding_boxes)) {
  print(bounding_boxes[i,])
  mursst_test <- griddap(erd_info, 
                       latitude = c(bounding_boxes[i,]$min_lat, bounding_boxes[i,]$max_lat), 
                       longitude = c(bounding_boxes[i,]$min_long, bounding_boxes[i,]$max_long),
                       time = c("2019-06-01", "2019-08-31"),
                       fields = 'analysed_sst',
                       # store the ncf
                       store = disk(path = here('rawdata', 'temp', 'erddap', 'forsitesst')),
                       # don't immediately read it into R
                       read = FALSE
                       )
}

# 2020
for (i in 1:nrow(bounding_boxes)) {
  print(bounding_boxes[i,])
  mursst_test <- griddap(erd_info, 
                       latitude = c(bounding_boxes[i,]$min_lat, bounding_boxes[i,]$max_lat), 
                       longitude = c(bounding_boxes[i,]$min_long, bounding_boxes[i,]$max_long),
                       time = c("2020-06-01", "2020-08-31"),
                       fields = 'analysed_sst',
                       # store the ncf
                       store = disk(path = here('rawdata', 'temp', 'erddap', 'forsitesst')),
                       # don't immediately read it into R
                       read = FALSE
                       )
}

# 2021
for (i in 1:nrow(bounding_boxes)) {
  print(bounding_boxes[i,])
  mursst_test <- griddap(erd_info, 
                       latitude = c(bounding_boxes[i,]$min_lat, bounding_boxes[i,]$max_lat), 
                       longitude = c(bounding_boxes[i,]$min_long, bounding_boxes[i,]$max_long),
                       time = c("2021-06-01", "2021-08-31"),
                       fields = 'analysed_sst',
                       # store the ncf
                       store = disk(path = here('rawdata', 'temp', 'erddap', 'forsitesst')),
                       # don't immediately read it into R
                       read = FALSE
                       )
}


# 2022
for (i in 1:nrow(bounding_boxes)) {
  print(bounding_boxes[i,])
  mursst_test <- griddap(erd_info, 
                       latitude = c(bounding_boxes[i,]$min_lat, bounding_boxes[i,]$max_lat), 
                       longitude = c(bounding_boxes[i,]$min_long, bounding_boxes[i,]$max_long),
                       time = c("2022-06-01", "2022-08-31"),
                       fields = 'analysed_sst',
                       # store the ncf
                       store = disk(path = here('rawdata', 'temp', 'erddap', 'forsitesst')),
                       # don't immediately read it into R
                       read = FALSE
                       )
}

# 2023
for (i in 1:nrow(bounding_boxes)) {
  print(bounding_boxes[i,])
  mursst_test <- griddap(erd_info, 
                       latitude = c(bounding_boxes[i,]$min_lat, bounding_boxes[i,]$max_lat), 
                       longitude = c(bounding_boxes[i,]$min_long, bounding_boxes[i,]$max_long),
                       time = c("2023-06-01", "2023-08-31"),
                       fields = 'analysed_sst',
                       # store the ncf
                       store = disk(path = here('rawdata', 'temp', 'erddap', 'forsitesst')),
                       # don't immediately read it into R
                       read = FALSE
                       )
}

# 2024
for (i in 1:nrow(bounding_boxes)) {
  print(bounding_boxes[i,])
  mursst_test <- griddap(erd_info, 
                       latitude = c(bounding_boxes[i,]$min_lat, bounding_boxes[i,]$max_lat), 
                       longitude = c(bounding_boxes[i,]$min_long, bounding_boxes[i,]$max_long),
                       time = c("2024-06-01", "2024-08-31"),
                       fields = 'analysed_sst',
                       # store the ncf
                       store = disk(path = here('rawdata', 'temp', 'erddap', 'forsitesst')),
                       # don't immediately read it into R
                       read = FALSE
                       )
}
# classify files
for (i in 1:length(list.files(here('rawdata', 'temp', 'erddap', 'forsitesst')))){
  ncname <- list.files(here('rawdata', 'temp', 'erddap', 'forsitesst'))[i]
  sstfile = nc_open(here('rawdata', 'temp', 
                     'erddap', 'forsitesst',
                     ncname))
  # get variables
  nclong <- ncvar_get(sstfile, 'longitude')
  nclat <- ncvar_get(sstfile, 'latitude')
  time <- ncvar_get(sstfile, 'time')
  tunits <- ncatt_get(sstfile,"time","units")
  # set CFtime to the units from the ncdf
  cfs <- CFtime(tunits$value, calendar = 'proleptic_gregorian', time)
  # extract the date of the map
  timestamps <- as_timestamp(cfs)
  
  # don't pick a corner
  nc_sf = expand_grid(long = nclong, lat = nclat)[25,] %>%
    # convert to points
    st_as_sf(., 
           coords = c('long', 'lat'),
           crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
    # add a buffer in case its on a line
    st_buffer(.,200)
  
  nc_bboxmatch <- st_intersection(nc_sf, bbox_sf) %>%
    st_drop_geometry() %>%
    mutate(date = timestamps[1]) %>%
    mutate(filename = ncname)
  
   write_csv(nc_bboxmatch, 
            here(erd_data, paste0(datefield,'site_erddap_key.csv')),
            append = TRUE)
  
  nc_close(sstfile)
}


```

# extract site values

```{r to-extract}
extract_key <- read_csv(here(erd_data, '20251212_site_erddap_key.csv'),
                            col_names = c('box_name', 'date', 'filename'))

regions <- read_csv(here(dataedna,
                         '20251113_peco_region-name-plotting.csv'))[-1] %>%
    rename(region_label = region_name_label) 


# site based lats and longs
# Note this isn't really different from 'regions' but 
# the code is set up to take this dataframe so lets just keep it
by_site <- regions %>%
  dplyr::select(!c(avg_region_lat, samples, sites)) %>%
  distinct()

# turn sites into coordinate points
site_sf <- st_as_sf(by_site, 
                    coords = c('decimal_longitude', 'decimal_latitude'),
                    crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0",
                    remove = FALSE)

# match sites and bboxes
site_bb <- st_intersection(site_sf, bbox_sf)

# to extract
extract_w_files <- extract_key %>%
  mutate(year = lubridate::year(date)) %>%
  select(!date) %>%
  # match sites with boxes for extraction
  left_join(site_bb) 

# vector of filenames
files_to_search <- extract_w_files %>%
  select(filename) %>%
  distinct() %>%
  arrange(filename)
```

## extraction loops

```{r extraction loop}
for (ncfile in files_to_search$filename){
  # lets be up to date on this...
  message(paste0('searching in ', ncfile))
  
  #filter points
  points_to_extract <- extract_w_files %>%
    filter(filename == ncfile) %>%
    distinct() %>%
    #for easy joining
    rownames_to_column('ID') %>%
    mutate(ID = as.numeric(ID))
  
  coords4extract <- points_to_extract %>%
    select(decimal_longitude, decimal_latitude) %>%
    vect(., geom=c("decimal_longitude", "decimal_latitude"), 
         crs="WGS84", keepgeom=FALSE)
    
  # add a 2 km buffer, the way ID is done in terra extract its easier to just do all of it
    # I may need to think about a better way to reduce time
  obisbuff2 <- terra::buffer(coords4extract,2000)
  
  #n_layer = unique(points_to_extract$layer_num)
  
  # open ncfile as a raster
  sstfile <- rast(here('rawdata', 'temp', 
                     'erddap', 'forsitesst', 
                     ncfile))
  sst_names <- names(sstfile)
  for (i in 1:length(sst_names)){
    # that layer
    sstlayer <- sstfile[[i]]
    # sst with buffer
    point_val = terra::extract(sstlayer, obisbuff2,
                             fun = function(x) mean(x, na.rm = T))

    point_val_full <- point_val %>%
      mutate(filename = ncfile) %>%
      mutate(date = terra::time(sstlayer)) %>%
      # one column for sst
      rename(sea_surface_temperature = paste0('analysed_sst_',as.character(i))) %>%
      left_join(points_to_extract) %>%
      filter(!(sea_surface_temperature == 'NaN')) %>%
      filter(!is.na(box_name)) %>%
      dplyr::select(!c(ID, geometry, region_code, region_code_label)) %>%
      # order columns specifically
      relocate(sea_surface_temperature, filename, date, box_name, year, site_code, 
               region_name, region_label, region_lat_name,
               decimal_latitude, decimal_longitude)
        # write directly to csv so it isn't stored in memory
        # # Note also this is not giving column names so we'll want to read it in and fix that asap
       write_csv(point_val_full, 
            here(erd_data, paste0(datefield, 'sitesst_erddap_extract.csv')),
            append = TRUE)
  }

  message('done')
}


```


# find missing

```{r}
# load in and give headers
extracted_sst <- read_csv(here(erd_data, '20251212_sitesst_erddap_extract.csv'),
                          col_names = c('sst', 'filename', 'date', 'box_name', 
                                        'year', 'site_code', 'region_name', 
                                        'region_label', 'region_lat_name',
                                        'decimal_latitude', 'decimal_longitude')) %>%
  distinct()

missing_sites <- by_site %>%
  filter(!(site_code %in% extracted_sst$site_code))

extract_w_files_missed <- extract_w_files %>%
  filter(site_code %in% missing_sites$site_code)

# vector of filenames
files_to_search_missed <- extract_w_files_missed %>%
  select(filename) %>%
  distinct() %>%
  arrange(filename)
```

## extract missing
Increase the buffer 

```{r}
for (ncfile in files_to_search_missed$filename){
  # lets be up to date on this...
  message(paste0('searching in ', ncfile))
  
  #filter points
  points_to_extract <- extract_w_files_missed %>%
    filter(filename == ncfile) %>%
    distinct() %>%
    #for easy joining
    rownames_to_column('ID') %>%
    mutate(ID = as.numeric(ID))
  
  coords4extract <- points_to_extract %>%
    select(decimal_longitude, decimal_latitude) %>%
    vect(., geom=c("decimal_longitude", "decimal_latitude"), 
         crs="WGS84", keepgeom=FALSE)
    
  # add a 4 km buffer for missedsites, the way ID is done in terra extract its easier to just do all of it
    # I may need to think about a better way to reduce time
  obisbuff2 <- terra::buffer(coords4extract,4000)
  
  #n_layer = unique(points_to_extract$layer_num)
  
  # open ncfile as a raster
  sstfile <- rast(here('rawdata', 'temp', 
                     'erddap', 'forsitesst', 
                     ncfile))
  sst_names <- names(sstfile)
  for (i in 1:length(sst_names)){
    # that layer
    sstlayer <- sstfile[[i]]
    # sst with buffer
    point_val = terra::extract(sstlayer, obisbuff2,
                             fun = function(x) mean(x, na.rm = T))

    point_val_full <- point_val %>%
      mutate(filename = ncfile) %>%
      mutate(date = terra::time(sstlayer)) %>%
      # one column for sst
      rename(sea_surface_temperature = paste0('analysed_sst_',as.character(i))) %>%
      left_join(points_to_extract) %>%
      filter(!(sea_surface_temperature == 'NaN')) %>%
      filter(!is.na(box_name)) %>%
            dplyr::select(!c(ID, geometry, region_code, region_code_label)) %>%
      # order columns specifically
      relocate(sea_surface_temperature, filename, date, box_name, year, site_code, 
               region_name, region_label, region_lat_name,
               decimal_latitude, decimal_longitude)
        # write directly to csv so it isn't stored in memory
        # # Note also this is not giving column names so we'll want to read it in and fix that asap
       write_csv(point_val_full, 
            here(erd_data, paste0(datefield, 'sitesst_erddap_extract_missed.csv')),
            append = TRUE)
  }

  message('done')
}

missing_sst <- read_csv(here(erd_data, '20251212_sitesst_erddap_extract_missed.csv'),
                          col_names = c('sst', 'filename', 'date', 'box_name', 
                                        'year', 'site_code', 'region_name', 
                                        'region_label', 'region_lat_name',
                                        'decimal_latitude', 'decimal_longitude')) 

still_missing_sites <- missing_sites %>%
  filter(!(site_code %in% missing_sst$site_code))

extract_w_files_still_missed <- extract_w_files %>%
  filter(site_code %in% still_missing_sites$site_code)

# vector of filenames
files_to_search_still_missed <- extract_w_files_still_missed %>%
  select(filename) %>%
  distinct() %>%
  arrange(filename)

# still missing try 10k buffer

for (ncfile in files_to_search_still_missed$filename){
  # lets be up to date on this...
  message(paste0('searching in ', ncfile))
  
  #filter points
  points_to_extract <- extract_w_files_still_missed %>%
    filter(filename == ncfile) %>%
    distinct() %>%
    #for easy joining
    rownames_to_column('ID') %>%
    mutate(ID = as.numeric(ID))
  
  coords4extract <- points_to_extract %>%
    select(decimal_longitude, decimal_latitude) %>%
    vect(., geom=c("decimal_longitude", "decimal_latitude"), 
         crs="WGS84", keepgeom=FALSE)
    
  # add a 10 km buffer for missed missed sites, the way ID is done in terra extract its easier to just do all of it
    # I may need to think about a better way to reduce time
  obisbuff2 <- terra::buffer(coords4extract,10000)
  
  #n_layer = unique(points_to_extract$layer_num)
  
  # open ncfile as a raster
  sstfile <- rast(here('rawdata', 'temp', 
                     'erddap', 'forsitesst', 
                     ncfile))
  sst_names <- names(sstfile)
  for (i in 1:length(sst_names)){
    # that layer
    sstlayer <- sstfile[[i]]
    # sst with buffer
    point_val = terra::extract(sstlayer, obisbuff2,
                             fun = function(x) mean(x, na.rm = T))

    point_val_full <- point_val %>%
      mutate(filename = ncfile) %>%
      mutate(date = terra::time(sstlayer)) %>%
      # one column for sst
      rename(sea_surface_temperature = paste0('analysed_sst_',as.character(i))) %>%
      left_join(points_to_extract) %>%
      filter(!(sea_surface_temperature == 'NaN')) %>%
      filter(!is.na(box_name)) %>%
      dplyr::select(!c(ID, geometry, region_code, region_code_label)) %>%
      # order columns specifically
      relocate(sea_surface_temperature, filename, date, box_name, year, site_code, 
               region_name, region_label, region_lat_name,
               decimal_latitude, decimal_longitude)
    
        # write directly to csv so it isn't stored in memory
        # # Note also this is not giving column names so we'll want to read it in and fix that asap
       write_csv(point_val_full, 
            here(erd_data, paste0(datefield, 'sitesst_erddap_extract_missed2.csv')),
            append = TRUE)
  }

  message('done')
}

missing_sst2 <- read_csv(here(erd_data, '20251212_sitesst_erddap_extract_missed2.csv'),
                          col_names = c('sst', 'filename', 'date', 'box_name', 
                                        'year', 'site_code', 'region_name', 
                                        'region_label', 'region_lat_name',
                                        'decimal_latitude', 'decimal_longitude')) 

still_missing_sites2 <- still_missing_sites %>%
  filter(!(site_code %in% missing_sst2$site_code))

# bind into original object
extracted_sst <- extracted_sst %>%
  bind_rows(missing_sst) %>%
  bind_rows(missing_sst2) %>%
  distinct()
```

```{r extractions summary}
sst_summarystats <- extracted_sst %>%
  filter(year > 2020) %>%
  group_by(site_code) %>%
  summarise(sst_max = max(sst),
            sst_min = min(sst),
            sst_median = median(sst),
            sst_mean = mean(sst),
            sst_sd = sd(sst),
            last5_temp95 = quantile(sst, probs = .95),
            last5_temp90 = quantile(sst, probs = .9),
            last5_temp75 = quantile(sst, probs = .75),
            last5_temp25 = quantile(sst, probs = .25),
            last5_temp10 = quantile(sst, probs = .1),
            last5_temp05 = quantile(sst, probs = .05)
            ) %>%
  left_join(regions) %>%
  dplyr::select(!c(samples, sites, avg_region_lat))

write_csv(sst_summarystats, here(erd_data, paste0(datefield, 'sitesst_last5_summarystats.csv')))

sst_by_regionyear <- extracted_sst %>%
  select(!c(filename, date, box_name)) %>%
  distinct() %>%
  group_by(region_name, year) %>%
  mutate(sst_max = max(sst),
         sst_min = min(sst),
            sst_median = median(sst),
            sst_mean = mean(sst),
            sst_sd = sd(sst),
            region_temp95 = quantile(sst, probs = .95),
            region_temp90 = quantile(sst, probs = .9),
            region_temp75 = quantile(sst, probs = .75),
            region_temp25 = quantile(sst, probs = .25),
            region_temp10 = quantile(sst, probs = .1),
            region_temp05 = quantile(sst, probs = .05)) %>%
  ungroup() %>%
  group_by(region_name) %>%
  mutate(avg_region_lat = mean(decimal_latitude)) %>%
  ungroup() %>%
  select(!c(site_code, sst, decimal_longitude, decimal_latitude)) %>%
  distinct() %>%
  arrange(region_lat_name)

write_csv(sst_by_regionyear, here(erd_data, paste0(datefield, 'mean-summer-sst_by-region.csv')))

sst_by_siteyear <- extracted_sst %>%
  select(!c(filename, date, box_name)) %>%
  distinct() %>%
  group_by(site_code, year) %>%
  summarise(sst_max = max(sst),
            sst_min = min(sst),
            sst_median = median(sst),
            sst_mean = mean(sst),
            sst_sd = sd(sst),
            site_temp95 = quantile(sst, probs = .95),
            site_temp90 = quantile(sst, probs = .9),
            site_temp75 = quantile(sst, probs = .75),
            site_temp25 = quantile(sst, probs = .25),
            site_temp10 = quantile(sst, probs = .1),
            site_temp05 = quantile(sst, probs = .05)
            ) %>%
  ungroup() %>%
  distinct() %>% 
  left_join(regions) %>%
  select(!c(region_code, region_code_label, samples, sites)) %>%
  arrange(region_lat_name)

write_csv(sst_by_siteyear, here(erd_data, paste0(datefield, 'summer-sst-summary-stats_by-site.csv')))
```