---
title: "ch2-getobisgbif"
author: "Kate Sheridan"
date: "`r Sys.Date()`"
output: html_document
---

# Download OBIS and GBIF datasets

NOTE; NEEDS OBJECT REMOVAL ADDED IN
NEEDS TESTING; this is pasted in from previous scripts and as of 2025-12-12 has not yet been tested in this repository

## GBIF
This script sets up a gbif download, then cleans the resulting download.
Download is based on taxon keys plus bounding box.

NOTE This is the whole script, but if you want to use the official gbif download from the thesis, skip to "THESIS PIPELINE STARTS HERE" 

```{r setup, include=FALSE}
#note: I had trouble installing rgbif because of a dependency that didn't want to compile. 
#The solution if something like this happens again is make sure the package Rcpp is up to date.
# gbif
library(rgbif)
library(taxize)
library(countrycode)
# obis
library(robis)
library(obistools)

# both
library(tidyverse)
library(sf)
library(here)

# dplyr is better for these
library(conflicted)
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')

# here filepaths
spatial = here('rawdata', 'spatial')
datefield <- '20251212_'
```

This function pulls from worms using worrms
```{r worms-function}
library(worrms)

## requires worrms to be loaded
# input is vector of species names
# use wm_records_names to extract 
# this doesn't need the + signs

# right now this breaks if not-found
search_records_worms <- function(spnames) {
  search <- tibble()
  for (i in spnames) {
    print(paste0('searching for ', i))
    record <- wm_records_names(i, marine_only = FALSE)
    message('done')
    search <- append(search, record)
  }
  names(search) <- spnames
  search_output <- map_dfr(.x = search, ~ data.frame(.x), .id = 'query') %>%
    janitor::clean_names() %>%
    select(!(c(url, taxon_rank_id, citation, lsid, modified)))
  return(search_output)
}
```

This taxa list is only fish

```{r taxa list}
marine_taxa_fishonly <- c('Actinopterygii', # actinopterygii/actinopteri not part of gbif backbone
                          'Elasmobranchii', 'Holocephali', 'Leptocardii', 
                          'Myxini', 'Sarcopterygii',
                          'Acipenseriformes', 'Albuliformes', 'Anguilliformes',
                          'Ateleopodiformes', 'Atheriniformes', 'Aulopiformes', 
                          'Batrachoidiformes', 'Beloniformes', 'Beryciformes', 
                          'Cetomimiformes', 'Characiformes', 'Clupeiformes', 
                          'Cypriniformes', 'Cyprinodontiformes', 'Elopiformes',
                          'Esociformes', 'Gadiformes', 'Gasterosteiformes', 
                          'Gobeisociformes', 'Gonorynchiformes', 'Lampriformes', 
                          'Lepidosteiformes', 'Lophiiformes', 'Mugiliformes', 
                          'Myctophiformes', 'Notacanthiformes', 'Ophiodiiformes', 
                          'Osmeriformes', 'Osteoglossiformes', 'Perciformes',
                          'Pleuronectiformes', 'Polymixiformes', 'Polypteriformes', 
                          'Saccopharyngiformes', 'Salmoniformes', 'Scorpaeniformes', 
                          'Siluriformes', 'Stephanoberyciformes', 'Stomiiformes', 
                          'Synbranchiformes', 'Syngnathiformes', 'Tetraodontiformes', 
                          'Zeiformes')

```


Taxize will get gbif keys for these
```{r}
gbif_taxon_keys <-
  marine_taxa_fishonly %>%
  get_gbifid_(method="backbone") %>% # match names to the GBIF backbone to get taxonkeys
  imap(~ .x %>% mutate(original_sciname = .y)) %>% # add original name back into data.frame
  bind_rows() %>% # combine all data.frames into one
  filter(matchtype == "EXACT" & status == "ACCEPTED") %>% # get only accepted and matched names
  filter(kingdom == "Animalia") # avoid homonyms with plants
```


Set up the download from gbif using bounding box
Note this bounding box is larger than necessary, but the output will be filtered to our study area.
```{r gbif download}
#set bounding box
#pacific coast to equator
gbif_bb <- "POLYGON((-168 69, -168.7 64.8, -166 63.5, -168.3 60, -162.2 57, -160 54.4, -148 57, -138 58, -133 52, -125.6 48, -126 42.7, -123 34, -117 30, -107.5 19.3, -94 14, -87.3 10.5, -81.5 6.6, -79.4 2.3, -82.4 -1, -80 -1, -76.8 3.7, -77.25 7.25, -78.56 9.1, -79.5 9.2, -81.3 8.1, -85.71 11.23, -87.27 13.5, -91.34 14.38, -94.4 16.6, -96.6 15.9, -105 20.2, -104.9 21.8, -112.8 31.4, -120 34.75, -121.8 37.2, -121.6 38, -122.75 38.2, -124 40.4, -123.8 41.6, -124.3 42.85, -123 46, -122.3 47, -122 47.6, -122.4 49.4, -126.6 51.5, -129.6 54.6, -130.85 56.1, -133.6 58.4, -135.6 59.5, -137.5 58.9, -139.5 60, -143.9 60.2, -146 61.2, -149.7 61.6, -151.75 61.1, -157.6 59, -161.4 59.3, -163.8 60.6, -163.8 60, -165.4 61.7, -162.2 63.3, -160.7 63.5, -160.7 65, -159.5 66.5, -165.3 68.5, -156.7 71.1, -168 69))"

# download data
user <- "YOURUSERNAME"
pwd <- "YOURPWD"
email <- "YOUREMAIL"

#specific taxa within bounding box
occ_download(pred_in("taxonKey", unique(gbif_taxon_keys$usagekey)),
             pred("hasCoordinate", TRUE), 
             pred_within(gbif_bb), 
             format = "SIMPLE_CSV", 
             user = user, pwd = pwd, email = email)

#then go to gbif to download the file or use download_get
```

Then download file from gbif, rename to proper format, move to folder.

### THESIS PIPELINE STARTS HERE
```{r load in gbif}
#occ_download_wait('0070329-250525065834625')

gbif_data <- occ_download_get('0070329-250525065834625') %>%
    occ_download_import()
```

### clean gbif
adopted from 'Big data biogeography' course from Physalia
```{r gbif cleaning}
glimpse(gbif_data)

gbif_subset <- gbif_data %>%
  # select only the columns we need
  select(verbatimScientificName, decimalLongitude, decimalLatitude, countryCode,
         stateProvince, gbifID, phylum, order, class, family, genus, species, 
         taxonKey, taxonRank, coordinateUncertaintyInMeters, year, month,
         basisOfRecord,occurrenceStatus, individualCount) %>% 
  mutate(summer = ifelse(month >=5 & month <=9, 't', 'f')) %>%
  #filter(month >=5 & month <=9)
  mutate(countryCode = countrycode(gbif_data$countryCode, origin =  'iso2c', destination = 'iso3c'))

# which do we have?
unique(gbif_subset$class)

# remove unsuitable data sources, especially fossils
table(gbif_subset$basisOfRecord)

# Age of records
#older records may be less reliable so we can delete older records to a point. cutoff may vary depending on groups, the sample is 1945. for many regions, there has been enough change in recording practices and land use since world war 2, this is a reasonable cutoff
table(gbif_subset$year)


#the conclusion is that material sample is dna, etc. worth keeping for my work, but not all
gbif_cl <- gbif_subset %>% 
  # remove high uncertainty records, but keep NAs. It isn't always provided
  filter(coordinateUncertaintyInMeters/1000 <= 100 | is.na(coordinateUncertaintyInMeters)) %>%
  # removing Fossils, Machine observations, and "occurrence"?
  filter(basisOfRecord == "HUMAN_OBSERVATION" | 
           basisOfRecord == "OBSERVATION" | 
           basisOfRecord == "PRESERVED_SPECIMEN" | 
           basisOfRecord == "MATERIAL_SAMPLE" | 
           is.na(basisOfRecord)) %>%
  filter(occurrenceStatus == "PRESENT") %>%
  filter(year > 1944) %>% # remove records from before second world war
  mutate(taxonRank = tolower(taxonRank)) %>%
  select(!(c(occurrenceStatus, coordinateUncertaintyInMeters))) %>%
  # remove ones that have NAs in higher taxonomy where it shouldn't
  filter(!(class == "" & order == "")) %>%
  janitor::clean_names() %>%
  mutate(gbif_id = as.numeric(gbif_id)) %>%
  mutate(across(.cols = c(class, order, family, genus, species),
                .fns = ~ na_if(., ""))) %>%
  distinct()

# temp intermediate file until taxize is finished. 
# Sometimes restarting the script is necessary so reload here if needed
saveRDS(gbif_cl, here('rawdata','bigdata','20250626_gbif_pretaxize.rds'))
gbif_cl <- readRDS(here('rawdata','bigdata','20250626_gbif_pretaxize.rds'))
```

### GBIF Taxonomy

Update the taxonomy from GBIF to the WoRMS backbone.
 
```{r fish only}
# fish only dataframe
gbiffish <- gbif_cl %>%
  # 'unranked' includes some messed up but otherwise perfectly usable data
  filter(taxon_rank %in% c('species', 'subspecies', 'unranked')) %>%
  # known issues that can make taxize mad
  mutate(species = case_when(species == 'Beringraja inornate' ~ 'Beringraja inornata',
                             species == 'Caranx fasciatus' ~ 'Caranx sexfasciatus',
                             species == 'Gobiomorus dormitator' ~ 'Gobiomorus dormitor',
                             # revisions
                             species == 'Artedius asperulus' ~ 'Artedius fenestralis',
                             species == 'Astyanax fasciatus' ~ 'Psaliodon fasciatus',
                             species == 'Antennarius reticularis' ~ 'Antennatus strigatus',
                             species == 'Bryconamericus peruanus' ~ 'Eretmobrycon peruanus',
                             species == 'Bathytroctes alveatus' ~ 'Narcetes erimelas',
                             species == 'Catostomus mniotiltus' ~ 'Catostomus occidentalis',
                             species == 'Centropomus grandoculatus' ~ 'Centropomus medius',
                             species == 'Dactyloscopus byersi' ~ 'Dactyloscopus byersorum',
                             species == 'Icelinus strabo' ~ 'Icelinus borealis',
                             species == 'Leptocephalus holti' ~ 'Neocyema erythrosoma',
                             species == 'Lythrulon opalescens' ~ 'Haemulon flaviguttatum',
                             species == 'Normichthys campbelli' ~ 'Mirorictus taningi',
                             species == 'Oligocottus borealis' ~ 'Oligocottus maculosus',
                             species == 'Ostracion cubicus' ~ 'Ostracion cubicum',
                             species == 'Oxycirrhites seftoni' ~ 'Oxycirrhites typus',
                             species == 'Raja inornata' ~ 'Beringraja inornata',
                             species == 'Sebastodes ayresii' ~ 'Sebastes rosaceus',
                             species == 'Sebastodes rosae' ~ 'Sebastes rosaceus',
                             species == 'Syngnathus tweedliei' ~ 'Syngnathus auliscus',
                             species == 'Synodus jenkinsi' ~ 'Synodus scituliceps',
                             # species with unresolved name conflicts
                             # or under revision and freshwater so not in WoRMS
                             species %in% c('Eretmobrycon dahli',
                                            'Eretmobrycon guaytarae', 'Hypsoblenniops rickettsi',
                                            'Idiacanthus panamensis') ~ NA_character_,
                             TRUE ~ species)) %>%
  filter(!(is.na(species)))


gbiffish_taxonomy <- gbiffish %>%
  # reduce columns to only taxonomy related ones
  select(phylum, class, order, family, genus, species, taxon_key, taxon_rank) %>%
  distinct() %>%
  # get rid of higher level unranked
  filter(!(is.na(species))) %>%
  # now we know they're all species
  mutate(taxon_rank = ifelse(taxon_rank == 'unranked', 'species', taxon_rank)) %>%
  distinct(species, .keep_all = TRUE)

# inserting +'s for the temporary search issue
worms_search_fish <- gbiffish_taxonomy %>%
  select(species, taxon_key) %>%
  mutate(query = gsub(" ", "+", species))

worms_fish <- get_wormsid_(sci_com = worms_search_fish$query, 
                           marine_only = FALSE, fuzzy = TRUE,
                           accepted = FALSE)
# get_wormsid_ gives back null and length 0 elements
## first remove with compact and discard then map_dfr
worms_fish2 <- worms_fish %>% 
  compact() %>%
  discard( ~ nrow(.x) == 0) %>%
  map_dfr( ~ data.frame(.x), .id = 'query')

# this cleaning step is relatively temporary until they fix the curl issue
worms_fish2 <- worms_fish2 %>%
  #rename(taxa = scientificname) %>%
  left_join(worms_search_fish) %>%
  rename(worms_aphiaid = AphiaID,
        worms_sciname = scientificname) %>%
  add_count(query, name = 'n_result') %>%
  filter(!(n_result >= 2 & status != 'accepted')) %>%
  distinct(query, .keep_all = TRUE) %>%
  select(!n_result) %>%
  filter(!(is.na(worms_sciname)))

# extract higher and info about freshwater
### this is my custom function!
fish_records <- search_records_worms(worms_fish2$worms_sciname)

# refine results to be useful
fish_records2 <- fish_records %>%
  # only animal hits
  filter(kingdom == 'Animalia') %>%
  # generate terrestrial-only and freshwater-only columns
  mutate(freshwater_only = ifelse((is_marine %in% c(0, NA) ==TRUE) & (is_brackish %in% c(0,NA) == TRUE) & (is_freshwater == 1),  "1", "0")) %>%
    mutate(terrestrial_only = ifelse((is_marine %in% c(0, NA) ==TRUE) & (is_brackish %in% c(0,NA) == TRUE) & (is_terrestrial == 1),  "1", "0")) %>%
  distinct() %>%
  # rename to format
  rename(worms_aphiaid = valid_aphia_id,
         original_aphiaid =aphia_id,
         worms_match = scientificname,
         worms_sciname = valid_name
         ) %>%
  # remove unneeded columns
  select(!(c(authority, parent_name_usage_id, is_freshwater, 
             is_terrestrial, is_marine, is_brackish, is_extinct, match_type, 
             valid_authority))) %>%
  relocate(c(worms_sciname, worms_aphiaid), .before = original_aphiaid) %>%
  # rank lowercase
  mutate(rank = tolower(rank)) %>%
  # remove unaccepted duplicates
  add_count(worms_match) %>%
  filter(!(n > 1 & status != "accepted")) %>%
  select(!(n)) %>%
  distinct()

gbif_fish_taxupdate <- gbiffish %>%
  select(!(c(phylum, class, order, family, genus))) %>%
  rename(gbif_species = species) %>%
  right_join(fish_records2, by = c('gbif_species' = 'query')) %>%
  select(!(c(gbif_id, individual_count, original_aphiaid,
             status, unacceptreason, verbatim_scientific_name,
             worms_match, terrestrial_only, taxon_rank))) %>%
  distinct() %>%
  mutate(scientific_name = coalesce(worms_sciname, gbif_species)) %>%
  relocate(scientific_name)


#save for now
write_csv(gbif_fish_taxupdate, here("processeddata", "bigdata",
                        paste0(datefield, "gbif-fish.csv")))

# remove gbif objects
# UPDATE
rm(gbif_cl)

```


## OBIS

We'll use Marine Ecoregions of the World to download OBIS.
This shapefile has been Simplified from the original download using QGIS using the Simplify function: method = Distance (Douglas-Peuker), Tolerance = 1
```{r set up ecoregions}
regions <- read_sf(here(spatial, "MEOW", "meow_ecos_simple.shp")) %>% 
  dplyr::select(c('ECOREGION', 'geometry')) %>%
  rename(region = ECOREGION) %>%
  mutate(type = 'meow')

#meow
regions$wkt <- st_as_text(regions$geometry) #create well known text of ecoregions

# use fishbase FAO regions to refine output
regions_nep <- regions %>% #filter for regions of interest
  filter(region %in% c("Aleutian Islands", "Gulf of Alaska", 
                       "North American Pacific Fijordland", 
                       "Puget Trough/Georgia Basin", 
                       "Oregon, Washington, Vancouver Coast and Shelf",
                        "Northern California", "Southern California Bight"))
```


```{r checklist function}
# requires robis to be loaded

# takes a dataframe that has our shapefiles loaded in with two main columns:
# wkt from st_as_txt() and region
obis_occur <- function(region_df) {
  # make blank object
  search <- tibble()
  region_geo <- region_df$wkt
  region_name <- region_df$region
  # loop by region
  for (i in 1:length(region_geo)) {
    #region_geo <- region_df$wkt
    print(paste0('searching for ', region_name[i]))
    found_sp <- occurrence(geometry = region_geo[i],
                          startdate = '1945-01-01') %>%
      # only keep animals
      filter(kingdom == 'Animalia') %>%
      # only keep things at the levels we want
      filter(taxonRank %in% c('Species', 'Subspecies',
                              'Tribe', 'Variety', 'Forma',
                              'Genus', 'Subgenus', NA))
      found_sp$region = paste(region_name[i])
    message('done')
    
    search <- bind_rows(search, found_sp)
  }
  return(search)
}
```

I had a lot of issues getting the whole download in one go, so to avoid restarting everything every time I split the regions so that they would download one or two at a time. Then saved the RDS in case the script or computer needed restarted.

```{r scrape-regions}
# use function to extract species list by region
# a few at a time
regions1 <- regions_nep[1:2,]
regions2 <- regions_nep[3,]
regions2_2 <- regions_nep[4,]
regions3 <- regions_nep[5,]
regions3_2 <- regions_nep[6,]
regions3_3 <- regions_nep[7,]
 
regions_nep_obis1 <- obis_occur(regions1)
regions_nep_obis2 <- obis_occur(regions2)
regions_nep_obis2_2 <- obis_occur(regions2_2)
regions_nep_obis3 <- obis_occur(regions3)
regions_nep_obis3_2 <- obis_occur(regions3_2)
regions_nep_obis3_3 <- obis_occur(regions3_3)

saveRDS(regions_nep_obis1, here('rawdata', 'bigdata', paste0(datefield, 'obis.RDS')))
saveRDS(regions_nep_obis2, here('rawdata', 'bigdata', paste0(datefield, 'obis2.RDS')))
saveRDS(regions_nep_obis2_2, here('rawdata', 'bigdata', paste0(datefield, 'obis2_2.RDS')))
saveRDS(regions_nep_obis3, here('rawdata', 'bigdata', paste0(datefield, 'obis3.RDS')))
saveRDS(regions_nep_obis3_2, here('rawdata', 'bigdata', paste0(datefield, 'obis3_2.RDS')))
saveRDS(regions_nep_obis3_3, here('rawdata', 'bigdata', paste0(datefield, 'obis3_3.RDS')))

# load if needed
regions_nep_obis1 <- readRDS(here('rawdata', 'bigdata', '20250619_obis.RDS'))
regions_nep_obis2 <- readRDS(here('rawdata', 'bigdata', '20250619_obis2.RDS'))
regions_nep_obis2_2 <- readRDS(here('rawdata', 'bigdata', '20250619_obis2_2.RDS'))
regions_nep_obis3 <- readRDS( here('rawdata', 'bigdata', '20250619_obis3.RDS'))
regions_nep_obis3_2 <- readRDS( here('rawdata', 'bigdata', '20250619_obis3_2.RDS'))
regions_nep_obis3_3 <- readRDS( here('rawdata', 'bigdata', '20250619_obis3_3.RDS'))


regions_nep_obis2 <- regions_nep_obis2 %>%
  bind_rows(regions_nep_obis2_2)

regions_nep_obis3 <- regions_nep_obis3 %>%
  bind_rows(regions_nep_obis3_2) %>%
  bind_rows(regions_nep_obis3_3)


```

Also, my computer got mad a few times if I tried to filter the data after combining all regions. So here we are cleaning it in chunks and combining the whole dataset as needed.

```{r reduce dataset}
# Things to remove
# might add insects and birds back in later
# make sure nothing new sneaks in with future downloads!
class_rem <- c('Insecta', 'Aves', 'Chilopoda')
order_rem <- c('Araneae', 'Pseudoscorpionida')

#smaller dataset with only name, coordinates, and a few taxonomic levels to filter
obis1_sub <- regions_nep_obis1 %>%
  filter(!is.na(eventDate)) %>%
  filter(marine == "TRUE") %>%
  filter(occurrenceStatus %in% c('Present', 'present', 'P', NA_character_)) %>%
  filter(basisOfRecord == "Human observation" | basisOfRecord == "HumanObservation" | basisOfRecord == "PreservedSpecimen" | basisOfRecord == "MaterialSample" | basisOfRecord == 'LivingSpecimen' | is.na(basisOfRecord)) %>%
  select(scientificName, decimalLatitude, decimalLongitude, date_year, eventDate, phylum, 
                class, order, family, genus, species, taxonRank, vernacularName,
         basisOfRecord, aphiaID, flags, region) %>%
  drop_na(phylum) %>%
  filter(!(class %in% class_rem)) %>%
  filter(!(order %in% order_rem)) %>%
  distinct()

obis2_sub <- regions_nep_obis2 %>%
  filter(!is.na(eventDate)) %>%
  filter(marine == "TRUE") %>%
  filter(occurrenceStatus %in% c('Present','present', 'P', NA_character_)) %>%
  filter(basisOfRecord == "Human observation" | basisOfRecord == "HumanObservation" | basisOfRecord == "PreservedSpecimen" | basisOfRecord == "MaterialSample" | basisOfRecord == 'LivingSpecimen' | is.na(basisOfRecord)) %>%
  select(scientificName, decimalLatitude, decimalLongitude, date_year, eventDate, phylum, 
                class, order, family, genus, species, taxonRank, vernacularName, 
         basisOfRecord, aphiaID, flags, region) %>%
  drop_na(phylum) %>%
  filter(!(class %in% class_rem)) %>%
  filter(!(order %in% order_rem)) %>%
  distinct()

obis3_sub <- regions_nep_obis3 %>%
  filter(!is.na(eventDate)) %>%
  filter(marine == "TRUE") %>%
  filter(occurrenceStatus %in% c('Present','present', 'P', NA_character_)) %>%
  filter(basisOfRecord == "Human observation" | basisOfRecord == "HumanObservation" | basisOfRecord == "PreservedSpecimen" | basisOfRecord == "MaterialSample" | basisOfRecord == 'LivingSpecimen' | is.na(basisOfRecord)) %>%
  select(scientificName, decimalLatitude, decimalLongitude, date_year, eventDate, phylum, 
                class, order, family, genus, species, taxonRank, vernacularName, 
         basisOfRecord, aphiaID, flags, region) %>%
  drop_na(phylum) %>%
  filter(!(class %in% class_rem)) %>%
  filter(!(order %in% order_rem)) %>%
  distinct()

# combine
obis_fish_region <- bind_rows(obis1_sub, obis2_sub, obis3_sub) %>%
  janitor::clean_names() %>%
  # add summer column
  mutate(summer = ifelse(str_detect(event_date, '-06-|-07-|-08-'), 't', 'f')) %>%
  # filter to fish only
  filter(phylum == "Chordata") %>%
  filter(class %in% c("Actinopteri", "Elasmobranchii", "Leptocardii",
                      "Holocephali", "Myxini", "Petromyzonti", "Teleostei") |
           order %in% c("Scorpaeniformes")) %>%
  filter(!(genus %in% c(NA_character_))) %>%
  distinct() %>%
  janitor::clean_names()

write.csv(obis_fish_region, here('processeddata', 'bigdata', 
                                 paste0(datefield, 'obis-fish_region.csv')))

```

# remove previous objects

to be added during testing

```{r remove}

```

# combine obis and gbif

```{r loadin}
#fish
gbif <- read_csv(here(bigdata, paste0(datefield, 'gbif-fish.csv')))

obis <- read_csv(here(bigdata, paste0(datefield, 'obis-fish_region.csv')))[-1]

# just fish, narrowed down to only coordinates and years
bigdat_fish <- full_join(gbif, obis) %>%
  # combine similar but different columns;
  # update script to make them the same to start
  mutate(year = coalesce(year, date_year),
         worms_aphiaid = coalesce(worms_aphiaid, aphia_id),
         species = coalesce(species, worms_sciname),
         rank = coalesce(rank, taxon_rank)) %>% 
  select(!(c(date_year, vernacular_name, flags,
             aphia_id, worms_sciname, taxon_rank, 
             kingdom, phylum, basis_of_record, event_date,
             taxon_key))) %>%
  mutate(rank = tolower(rank)) %>%
  mutate(rank = ifelse(is.na(species), "genus", "species")) %>%
  distinct()


write_csv(bigdat_fish, here(bigdata, datefield('obis-gbif_merge-fish.csv')))

# species list
taxa_all <- bigdat_fish %>%
  select(scientific_name, rank, class, order, family, genus, species) %>%
  distinct()

write.csv(taxa_all, here('processeddata','splists', 
                         paste0(datefield,'obis-gbif_fishlist.csv')))
```

