---
title: "start-here"
author: "Kate Sheridan"
date: "`r Sys.Date()`"
output: pdf_document
---

# Start Here

This file lists all packages and software that need to be installed for the scripts to run.

The R version used is 4.3.1.

This project uses `renv`. This means this project runs from its own environment. So, the project library in the 'renv' folder should contain all the packages needed to run the scripts but in case it doesn't, there are installation chunks below.

# Package installation
## general

Some packages are required for both chapters and 'other' scripts. 

```{r project}
# these are the packages needed to run the project and markdown files

#renv requires a restart of r
# only uncomment if necessary
#install.packages('renv')


#not reproducible but having an issue
#renv::hydrate('devtools')
#install.packages('devtools')
install.packages('remotes')
install.packages('BiocManager')


# load remotes
library(remotes)
# markdown
install_version('rmarkdown', '2.30')
install_version('knitr', '1.50')

renv::status()
renv::snapshot()
```

The code in this thesis relies heavily on tidyverse. 

```{r both chapters}
# plot related 
install.packages('tidyverse', 'RColorBrewer', 'PNWColors', 'plotly',
                 'ggrepel', 'ggstatsplot', 'ggeffects', 'ggmap', 'ggpubr', 'gt', 'pheatmap')
devtools::install_github("cmartin/ggConvexHull")
# issue with installing eulerr, using this version
devtools::install_github("jolars/eulerr")

# taxonomy-related
install.packages('taxize', 'rfishbase')

#renv::hydrate('robis')
install.packages("robis")
devtools::install_github("iobis/obistools")

# mapping, spatial
install.packages('rnaturalearth', 'geosphere', 'sf')

# for processing sequence data
BiocManager::install("dada2")
BiocManager::install("phyloseq")
BiocManager::install("DESeq2")
install.packages('seqinr', 'vsn')
devtools::install_github("tobiasgf/lulu")

renv::hydrate('ggtree')
#remotes::install_github("YuLab-SMU/ggtree")



# terra 
#install.packages('Rcpp')
#remotes::install_github("rspatial/terra")

install.packages('vegan', 'arsenal', 'report', 'recluster', 'blme',
                 'lmerTest', 'multcomp', 'optimx', 'iNEXT')

# also use here to improve ease of use with working directories
# and janitor for data cleaning
install.packages('here', 'janitor')
```





## scratch

some issues ; seems like a lot of it is my version of osx; test on another computer


```{r}

renv::hydrate('Rmpfr')
install.packages('Rmpfr')

install.packages('roxygen2')
install.packages('Rmpfr', type = "source", 
                 configure.args=c('--with-mpfr-include=/opt/homebrew/Cellar/mpfr'))



Sys.getenv('PKG_CONFIG_PATH')
Sys.getenv('PATH')
Sys.getenv()
#install.packages("usethis", verbose=TRUE)
#install.packages('withr')
withr::local_path("/opt/X11/lib/pkgconfig", Sys.getenv("PATH"), "suffix")


f <- function(x) {
  withr::with_path("/opt/local/include/harfbuzz", Sys.getenv("PATH"), "suffix")
  print(Sys.getenv("PATH"))
  install.packages(x)
}
f('devtools')

Sys.setenv(PKG_CONFIG_PATH = '/opt/local/include/harfbuzz')
```

