---
title: "ch1_coi-03-inext-readnum-v3"
output: html_document
date: "2024-04-09"
---

Here we want to test what will happen if we have different numbers of reads

We'll want to test some raw data but we'll start with the cleaned data, resampling reads to make a synthetic dataset of the size we want

This version uses resampling at multiple QC levels and generates a table, and has been updated with functions instead of copypaste.

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(iNEXT)
library(here)


# filepaths for use with here
# eDNA
dataedna <- here::here("processeddata", "2018calvert")
# save in
inextout <- here::here("output", "chapter1", "inext")
datefield <- '20251212_'
```

Note that the reads for pcrs are split between columns
```{r loadin}
coi_asv <- read_csv(here::here(dataedna, 
                                   '20240222_calvert2018_coi_asvmatrix_run-columns_inverts.csv')) %>%
  filter(pa_quadra3 == 1) %>%
  select(!c(pa_quadra1, pa_gcoct, pa_gcsept,
            reads_raw_gcsept, reads_raw_gcoct, reads_raw_quadra1,
            reads_raw_quadra3_NA,
            found_taxa_motu, common_name_motu)) %>%
  # also lets remove motu for now
  dplyr::rename(asv_pa = pa_quadra3)

# Reducing our dataset to only what we need
asv_reduced <-  coi_asv %>%
  select(sample_name, asv, asv_pa,
         reads_raw_quadra3_rep1, reads_raw_quadra3_rep2,
         reads_raw_quadra3_rep3,
         found_taxa, rank) %>%
  # combine pcr reads
  pivot_longer(cols = c(reads_raw_quadra3_rep2, reads_raw_quadra3_rep1,
                        reads_raw_quadra3_rep3),
               names_to = 'pcr',
               values_to = 'reads_raw') %>%
  select(!pcr) %>%
  filter(!is.na(reads_raw)) %>%
  # generate reads for asv per sample and species per sample
  mutate(asvsample = paste0(sample_name, '_', asv)) %>%
  group_by(asvsample) %>%
  mutate(reads_asvsample = sum(reads_raw)) %>%
  ungroup() %>%
  group_by(sample_name, found_taxa) %>%
  mutate(reads_spsample = sum(reads_raw)) %>%
  ungroup() %>%
  select(!reads_raw) %>%
  distinct()

# Also generate stats for the full dataset
full_stats <- read_csv(here('rawdata', '2018calvert',
                            '20240127_coi_2018calvert_rerun3_quadra20200604_read-retention_coi_bysample.csv')) %>%
  janitor::clean_names()

full_reduced <- full_stats %>%
  # clean sample name
  separate(sample_id, into = c('date', 'sample_pool'), extra = 'drop', sep = '_') %>%
  separate(sample_pool, into = c('pool', 'sample_name'), extra = 'drop', sep = '\\.') %>%
  select(sample_name, raw_f, raw_r, cutadapt_f, cutadapt_r, 
         denoised_f, denoised_r, merged, 
         nosingletons, nochimeras) %>%
  pivot_longer(!sample_name) %>%
  # lets put fw and r together now
  mutate(name = str_remove_all(name, '_f|_r')) %>%
  pivot_wider(values_fn = sum)

full_totals <- full_reduced %>%
  pivot_longer(!sample_name) %>%
  select(!sample_name) %>%
  pivot_wider(values_fn = sum)
```

# Simulate


## Empirical


~ 28,253,178 raw after demultiplexing
    - doesn't include unassigned to sample
~ 40,000,000 reads off the machine; forward and reverse

~ 10k difference: why?
~~ note, may need to update processing script where it makes totals

```{r empirical stats}
# fw + r reads off the machine
reads_total <- full_totals$raw
  

# 6,681,119 passed QC + on-target
reads_passedqc <- sum(asv_reduced$reads_asvsample)

# about 47.29% of total reads pass QC and on-target
# taking into consideration that raw reads == fw, r separate
postqc <- (reads_passedqc * 2)/reads_total
# saving the ratio to determine how many reads to generate from raw read number
# assuming the same ratio of QC/on-target

# how many reads per sample,
# what percent of total is in each sample
by_sample_stats <- asv_reduced %>% 
  left_join(full_reduced) %>%
  group_by(sample_name) %>% 
  mutate(reads_sample = sum(reads_asvsample)) %>%
  select(sample_name, reads_sample, raw) %>%
  ungroup() %>%
  distinct() %>%
  mutate(sample_percent = reads_sample/raw)

# 43,667.44 mean reads per sample passed QC
reads_avgsample <- mean(by_sample_stats$reads_sample)

# each read here is a combo of fw and r so 
# about .3% of total reads on target per each sample
(reads_avgsample * 2)/reads_total

# reads per asv-sample combo
# how many average 
by_asvsample_stats <- asv_reduced %>%
  left_join(full_reduced) %>%
  select(asvsample, sample_name, reads_asvsample, raw) %>%
  distinct() %>%
  mutate(asvsample_percent = reads_asvsample/raw) 
```

```{r all}
set.seed(2018)
sp_all <- coi_asv %>%
  filter(rank == 'species') %>%
  select(found_taxa, asv_pa, sample_name) %>%
  group_by(found_taxa, sample_name) %>%
  mutate(sample_read = sum(asv_pa)) %>%
  ungroup() %>%
  select(!(asv_pa)) %>%
  distinct() %>%
  pivot_wider(id_cols = c(found_taxa),
              names_from = sample_name, 
              values_from = sample_read) %>%
  column_to_rownames('found_taxa') %>%
  # make binary
  mutate(across(.cols = everything(), .fns = ~replace(., . > 0, 1))) %>%
  mutate(across(.cols = everything(), .fns = ~replace(., is.na(.), 0))) %>%
  as.incfreq()

sp_all_incfreq <- iNEXT(sp_all, q=0, datatype="incidence_freq")
```


# Simulation

## functions
```{r simulation functions}
simulate_qc <- function(target, postqc_reads, asvsample_stats){
  # empty vectors to fill
  qc = vector("character", target)
  spsamp = vector("character", target)
  #set the seed
  set.seed(2018)
  
  #simulate the population, part one
  #Assign origin
  qc = sample(c("Pass","Fail"), size = target,
                          prob=c(postqc_reads, 1 - postqc_reads),
                          replace = TRUE)
  
  #simulate the population, part two
  for (k in 1:target) {   
    if(qc[k] == "Fail") {spsamp[k] = sample(c('offtarget'), 
                                            size = 1,
                                            prob = c(1),
                                            replace = TRUE) } 
    if(qc[k] == "Pass") {spsamp[k] = sample(asvsample_stats$asvsample, 
                                           size = 1,
                                           prob=asvsample_stats$asvsample_percent,
                                           replace = TRUE) }
  }
  sim_result = as.data.frame(cbind(qc,spsamp)) %>%
  add_count(spsamp, name = 'reads_sim') %>%
  distinct() %>%
  rename(asvsample = spsamp)
  
  return(sim_result)
}

simulate_inext_sp <- function(simulated_data){
  pa_matrix <- simulated_data %>%
  filter(rank == 'species') %>%
  rename(species = found_taxa) %>%
  mutate(asv_pa = 1) %>%
  select(species, asv_pa, sample_name) %>%
  group_by(species, sample_name) %>%
  mutate(sample_read = sum(asv_pa)) %>%
  ungroup() %>%
  select(!(asv_pa)) %>%
  distinct() %>%
  pivot_wider(id_cols = c(species),
              names_from = sample_name, 
              values_from = sample_read) %>%
  column_to_rownames('species') %>%
  # make binary
  mutate(across(.cols = everything(), .fns = ~replace(., . > 0, 1))) %>%
  mutate(across(.cols = everything(), .fns = ~replace(., is.na(.), 0))) %>%
  as.incfreq()
  
  sp_incfreq <- iNEXT(pa_matrix, q=0, datatype="incidence_freq")
  
  return(sp_incfreq)
}

```



```{r simulations to target}
# set our target number of reads
target_half <- (reads_total * .5) / 2

simhalf <- simulate_qc(target_half, postqc, by_asvsample_stats) %>%
  left_join(asv_reduced)

# set our target number of reads
target_quarter <- (reads_total * .25) / 2

simquarter <- simulate_qc(target_quarter, postqc, by_asvsample_stats) %>%
  left_join(asv_reduced)

# set our target number of reads
target_tenth <- (reads_total * .1) / 2

simtenth <- simulate_qc(target_tenth, postqc, by_asvsample_stats) %>%
  left_join(asv_reduced)

# set our target number of reads
target_one <- (reads_total * .01) / 2

simone <- simulate_qc(target_one, postqc, by_asvsample_stats) %>%
  left_join(asv_reduced)

#
target_hundredth <- (reads_total * .001) / 2

simhundredth <- simulate_qc(target_hundredth, postqc, by_asvsample_stats) %>%
  left_join(asv_reduced)

target_thousandth <- (reads_total * .0001) / 2

simthousandth<- simulate_qc(target_thousandth, postqc, by_asvsample_stats) %>%
  left_join(asv_reduced)



```


```{r simulations into inext}
half_inext <- simulate_inext_sp(simhalf)
quarter_inext <- simulate_inext_sp(simquarter)
tenth_inext <- simulate_inext_sp(simtenth)
one_inext <- simulate_inext_sp(simone)
hundredth_inext <- simulate_inext_sp(simhundredth)
thousandth_inext <- simulate_inext_sp(simthousandth)

```


```{r}
inext_values <- sp_all_incfreq$iNextEst$coverage_based %>%
  mutate(sim_version = '100%') %>%
  bind_rows(half_inext$iNextEst$coverage_based) %>%
  mutate(sim_version = ifelse(is.na(sim_version), '50%', sim_version)) %>%
  # bind each dataframe
  bind_rows(quarter_inext$iNextEst$coverage_based) %>%
  mutate(sim_version = ifelse(is.na(sim_version), '25%', sim_version)) %>%
  bind_rows(tenth_inext$iNextEst$coverage_based) %>%
  mutate(sim_version = ifelse(is.na(sim_version), '10%', sim_version)) %>%
  bind_rows(one_inext$iNextEst$coverage_based) %>%
  mutate(sim_version = ifelse(is.na(sim_version), '1%', sim_version)) %>%
  bind_rows(hundredth_inext$iNextEst$coverage_based) %>%
  mutate(sim_version = ifelse(is.na(sim_version), '.1%', sim_version)) %>%
  bind_rows(thousandth_inext$iNextEst$coverage_based) %>%
  mutate(sim_version = ifelse(is.na(sim_version), '.01%', sim_version)) %>%
  select(SC, t, Method, qD, sim_version) %>%
  janitor::clean_names() %>%
  # criteria to keep
  mutate(keep = case_when(method == 'Observed' ~ 'observed',
                          #t > 200 & sc < .989999 ~ 'low',
                          t > 200 & sc > .999 ~ 'extra',
                          t < 10 & sc < .869 ~ 'extra',
                          sc < .91 & sc > .899 ~ '90%',
                          sc < .96 & sc > .9499 ~ '95%',
                          sc < .985 & sc > .9799 ~ '98%')) %>%
  filter(!(keep %in% c('extra'))) %>%
  # there has to be a better way to do this but i'm not focused enough to figure it out
  group_by(sim_version) %>%
  # all had 98%
  #mutate(has_98 = ifelse(any(keep %in% c("98%")), 'yes', 'no')) %>%
  mutate(has_95 = ifelse(any(keep %in% c("95%")), 'yes', 'no')) %>%
  mutate(has_90 = ifelse(any(keep %in% c("90%")), 'yes', 'no')) %>%
  ungroup() %>%
  mutate(keep = case_when(has_90 == 'no' & sc < .92 ~ '90%',
                          has_95 == 'no' & sc < .965 & sc > .94 ~ '95%',
                          TRUE ~ keep)) %>%
  # check again
  group_by(sim_version) %>%
  mutate(has_95 = ifelse(any(keep %in% c("95%")), 'yes', 'no')) %>%
  mutate(has_90 = ifelse(any(keep %in% c("90%")), 'yes', 'no')) %>%
  ungroup() %>%
  # all yes, remove extras
  filter(!(is.na(keep))) %>%
  select(!c(has_90, has_95)) %>%
  # smallest value within range
  group_by(sim_version, keep) %>%
  mutate(min_value = min(sc)) %>%
  ungroup() %>%
  filter(sc == min_value)

inext_values_plot <- inext_values  %>%
  # keep only 98% for now, the others might be useful later
  filter(keep %in% c('observed', '98%') | sim_version == '.01%') %>%
  add_count(sim_version) %>%
  filter(n == 1 | keep == '98%') %>%
    # remove below values for now
  filter(sim_version %in% c("100%","50%", "25%", "10%"))

```


```{r bayes1 test inextplot}
line_pal <- c("Rarefaction" = 'solid',
               "Observed" = "solid",
               "Extrapolation" = 'dashed'
              )

# blues dark to light with fewer reads
sim_pal <- c("100%" = '#2a3e57',
               "50%" = '#324a68', 
               "25%" = '#436188',  
               "10%" = '#688ab5', 
               "1%" = '#89a3c5',  
               ".1%" = '#9fb4d0',               
               ".01%" = '#aabdd5'
              )

# for pdf
ggplot() +
  # ribbon for confidence interval
  geom_ribbon(aes(ymin = sp_all_incfreq$iNextEst$coverage_based$qD.LCL,
                  ymax = sp_all_incfreq$iNextEst$coverage_based$qD.UCL,
                  x = sp_all_incfreq$iNextEst$coverage_based$t), 
              alpha = .25, colour = NA, fill = 'grey40') +
       geom_ribbon(aes(ymin = half_inext$iNextEst$coverage_based$qD.LCL,
                   ymax = half_inext$iNextEst$coverage_based$qD.UCL,
                   x = half_inext$iNextEst$coverage_based$t),
                 alpha = .25, colour = NA, fill = 'grey40') +
    geom_ribbon(aes(ymin = quarter_inext$iNextEst$coverage_based$qD.LCL,
                  ymax = quarter_inext$iNextEst$coverage_based$qD.UCL,
                  x = quarter_inext$iNextEst$coverage_based$t),
              alpha = .25, colour = NA, fill = 'grey40') +
    geom_ribbon(aes(ymin = tenth_inext$iNextEst$coverage_based$qD.LCL,
                  ymax = tenth_inext$iNextEst$coverage_based$qD.UCL,
                  x = tenth_inext$iNextEst$coverage_based$t), 
              alpha = .25, colour = NA, fill = 'grey40') +
      geom_ribbon(aes(ymin = hundredth_inext$iNextEst$coverage_based$qD.LCL,
                  ymax =  hundredth_inext$iNextEst$coverage_based$qD.UCL,
                  x = hundredth_inext$iNextEst$coverage_based$t), 
              alpha = .25, colour = NA, fill = 'grey40') +
      geom_ribbon(aes(ymin = thousandth_inext$iNextEst$coverage_based$qD.LCL,
                  ymax = thousandth_inext$iNextEst$coverage_based$qD.UCL,
                  x = thousandth_inext$iNextEst$coverage_based$t), 
              alpha = .25, colour = NA, fill = 'grey40') +
  geom_vline(aes(xintercept = inext_values_plot$t, colour = inext_values_plot$sim_version),
             linetype = 'dashed') +
  scale_color_manual(values = sim_pal) +
  # point at t(samples) and S.obs (species observed)
  geom_point(aes(sp_all_incfreq$DataInfo$T,
                 sp_all_incfreq$DataInfo$S.obs), 
             color = '#2a3e57', size = 4) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = sp_all_incfreq$iNextEst$coverage_based$t, 
                y = sp_all_incfreq$iNextEst$coverage_based$qD, 
                linetype = sp_all_incfreq$iNextEst$coverage_based$Method), 
            color = '#2a3e57', size = 1) +
  scale_linetype_manual(values = line_pal) +

  # # point at t(samples) and S.obs (species observed)
  geom_point(aes(half_inext$DataInfo$T,
                  half_inext$DataInfo$S.obs), 
             color = '#324a68', 
             size = 4,
             alpha = .7) +
   # line for qD, different linetypes for interpolation and rarefaction
   geom_line(aes(x = half_inext$iNextEst$coverage_based$t, 
                 y = half_inext$iNextEst$coverage_based$qD, 
                 linetype = half_inext$iNextEst$coverage_based$Method), 
             color = '#324a68', 
             size = 1,
             alpha = .7) +
  # point at t(samples) and S.obs (species observed)
  geom_point(aes(quarter_inext$DataInfo$T,
                 quarter_inext$DataInfo$S.obs),
             color = '#436188', 
             size = 4,
             alpha = .7) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = quarter_inext$iNextEst$coverage_based$t,
                y = quarter_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = '#436188', 
            size = 1,
            alpha = .7) +
    # point at t(samples) and S.obs (species observed)
  geom_point(aes(tenth_inext$DataInfo$T,
                 tenth_inext$DataInfo$S.obs),
             color = '#688ab5', 
             size = 4,
             alpha = .7) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = tenth_inext$iNextEst$coverage_based$t,
                y = tenth_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = '#688ab5', 
            size = 1,
            alpha = .7) +
      # point at t(samples) and S.obs (species observed)
  geom_point(aes(one_inext$DataInfo$T,
                 one_inext$DataInfo$S.obs),
             color = '#89a3c5', 
             size = 4,
             alpha = .7) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = one_inext$iNextEst$coverage_based$t,
                y = one_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = "#89a3c5", 
            size = 1,
            alpha = .7) +
    geom_point(aes(hundredth_inext$DataInfo$T,
                 hundredth_inext$DataInfo$S.obs),
             color = '#9fb4d0',
             size = 4) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = hundredth_inext$iNextEst$coverage_based$t,
                y = hundredth_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = '#9fb4d0', size = 1) +
      geom_point(aes(thousandth_inext$DataInfo$T,
                 thousandth_inext$DataInfo$S.obs),
             color = '#aabdd5', size = 4) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = thousandth_inext$iNextEst$coverage_based$t,
                y = thousandth_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = '#aabdd5', size = 1) +
  coord_cartesian(ylim = c(0,175), xlim = c(0,210))+
  theme_classic() +
    labs(x = 'Samples', y =  'Alpha diversity metric') +
  theme(plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.position = 'none') 

ggsave(here(inextout, paste0(datefield, "coi_readdepth-sim_b1-nolines.pdf")),
       device = 'pdf',
       height = 7,
       width = 8,
       units = 'in')

```


```{r}
line_pal <- c("Rarefaction" = 'solid',
               "Observed" = "solid",
               "Extrapolation" = 'dashed'
              )

# blues dark to light with fewer reads
sim_pal <- c("100%" = '#25140b',
               "50%" = '#5c321d', 
               "25%" = '#7C4428',  
               "10%" = '#d29374'
              )

# for pdf
ggplot() +
      geom_point(aes(x = inext_values_plot$t,
                 y = 4,
                 colour = inext_values_plot$sim_version),
             shape = 17,
             size = 5) +
  # ribbon for confidence interval
  geom_ribbon(aes(ymin = sp_all_incfreq$iNextEst$coverage_based$qD.LCL,
                  ymax = sp_all_incfreq$iNextEst$coverage_based$qD.UCL,
                  x = sp_all_incfreq$iNextEst$coverage_based$t), 
              alpha = .2, colour = NA, fill = 'grey40') + 
  geom_ribbon(aes(ymin = half_inext$iNextEst$coverage_based$qD.LCL,
                  ymax = half_inext$iNextEst$coverage_based$qD.UCL,
                  x = half_inext$iNextEst$coverage_based$t),
              alpha = .2, colour = NA, fill = 'grey40') +
    geom_ribbon(aes(ymin = quarter_inext$iNextEst$coverage_based$qD.LCL,
                  ymax = quarter_inext$iNextEst$coverage_based$qD.UCL,
                  x = quarter_inext$iNextEst$coverage_based$t),
              alpha = .2, colour = NA, fill = 'grey40') +
    geom_ribbon(aes(ymin = tenth_inext$iNextEst$coverage_based$qD.LCL,
                  ymax = tenth_inext$iNextEst$coverage_based$qD.UCL,
                  x = tenth_inext$iNextEst$coverage_based$t), 
              alpha = .2, colour = NA, fill = 'grey40') +
  scale_color_manual(values = sim_pal) +
  # point at t(samples) and S.obs (species observed)
  geom_point(aes(sp_all_incfreq$DataInfo$T,
                 sp_all_incfreq$DataInfo$S.obs), 
             color = '#25140b', size = 4) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = sp_all_incfreq$iNextEst$coverage_based$t, 
                y = sp_all_incfreq$iNextEst$coverage_based$qD, 
                linetype = sp_all_incfreq$iNextEst$coverage_based$Method), 
            color = '#25140b', size = 1) +
  scale_linetype_manual(values = line_pal) +
  # # point at t(samples) and S.obs (species observed)
  geom_point(aes(half_inext$DataInfo$T,
                  half_inext$DataInfo$S.obs), 
             color = '#5c321d', 
             size = 4,
             alpha = .7) +
   # line for qD, different linetypes for interpolation and rarefaction
   geom_line(aes(x = half_inext$iNextEst$coverage_based$t, 
                 y = half_inext$iNextEst$coverage_based$qD, 
                 linetype = half_inext$iNextEst$coverage_based$Method), 
             color = '#5c321d', 
             size = 1,
             alpha = .7) +
  # point at t(samples) and S.obs (species observed)
  geom_point(aes(quarter_inext$DataInfo$T,
                 quarter_inext$DataInfo$S.obs),
             color = '#7C4428', 
             size = 4,
             alpha = .7) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = quarter_inext$iNextEst$coverage_based$t,
                y = quarter_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = '#7C4428', 
            size = 1,
            alpha = .7) +
    # point at t(samples) and S.obs (species observed)
  geom_point(aes(tenth_inext$DataInfo$T,
                 tenth_inext$DataInfo$S.obs),
             color = '#d29374', 
             size = 4,
             alpha = .7) +
  # line for qD, different linetypes for interpolation and rarefaction
  geom_line(aes(x = tenth_inext$iNextEst$coverage_based$t,
                y = tenth_inext$iNextEst$coverage_based$qD,
                linetype = quarter_inext$iNextEst$coverage_based$Method),
            color = '#d29374', 
            size = 1,
            alpha = .7) +

  coord_cartesian(ylim = c(0,175), xlim = c(0,225),
                  expand = FALSE)+
  theme_classic() +
    labs(x = 'Samples', y =  'Alpha diversity metric') +
  theme(plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10),
        legend.position = 'none') 

ggsave(here(inextout, paste0(datefield, "coi_readdepth-sim_b1-nolines.pdf")),
       device = 'pdf',
       height = 7,
       width = 8,
       units = 'in')
```



# table


```{r}
library(gt)

sim_table <- inext_values %>%
  rename(samples = t,
         completeness = keep,
         species = q_d,
         percent_reads = sim_version) %>%
  select(percent_reads, samples, completeness) %>%
  gt() %>%
  tab_header(title = 'COI', subtitle = 'simulated reads')

sim_table

options(chromote.headless = "new")

gtsave(sim_table, here(inextout, paste0(datefield, "coi_readdepth-sim_table.png")))
gtsave(sim_table, here(inextout, paste0(datefield, "coi_readdepth-sim_table.pdf")))
```




